---
title: "Supplementary material of the paper: ``A new parametric spatial scan statistic for functional data: application to climate change data''"
author: "Zaineb Smida and Thibault Laurent"
date: "Last update: `r Sys.Date()`"
header-includes:
- \usepackage{booktabs}
- \usepackage{makecell}
output:
  github_document:   
    number_sections: yes
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(fig.align = "center",
               cache = F,
               cache.lazy = FALSE
)
options(kableExtra.latex.load_packages = FALSE)
```

This document presents the **R** codes used to obtain the computational results included in the paper "A new parametric spatial scan statistic for functional data: application to climate change data". To cite this work, please use:

Zaineb Smida and Thibault Laurent (2024). [A Hotelling spatial scan statistic for functional data: application to economic and climate data](), *WP*.

Packages needed:

```{r, message = F, warning = F}
library(mapsf) # cartography
library(maptiles) # import spatial contours
library(sf) # spatial data analysis
library(tidyverse) # tidyverse
library(latex2exp) # add LaTeX
library(ggh4x) # customize ggplot graphic
library(progress) # progress bar
library(rARPACK) # compute only the d first eigen values/eigen vectors
library(parallel) # parallel computing
```

The document is divided into three sections:

-   the first part presents the functions created for this work,
-   the second part allows to reproduce results presented in the section `Simulations Study`, 
-   the last part allows to reproduce the results presented in the section `Application to real data`.


Data can be provided upon request.

# Functions created

## Simulation of functional data

The function `simulvec()` allows to simulate functional data as presented in section `Simulation Study` in the article. It takes two arguments:

-   `npoints` the number of measurement,
-   `shape`: the law of the random variable $Z$ (`"gauss"`, `"student"`, `"chisq"` or `"exp"`).

```{r class.source = 'fold-hide'}
simulvec <- function(npoints, shape = "gauss") {
  # initi
  vect <- (0:npoints)/npoints
  k <- 1
  sigma <- 1/((k - 0.5) * pi)
  
  vecyphi <- sqrt(2) * sin(vect/sigma)
  
  if(shape == "gauss") {
    y <- rnorm(1, mean = 0, sd = sigma)
  } else {
    if(shape == "student") { 
      y <- sigma * rt(1, 4)
    } else  {
      if(shape == "chisq") { 
        y <- sigma * rchisq(1, 4)
      } else {
        y <- sigma * rexp(1, 0.5)
      }
    }
  }
  vecY <- y * vecyphi
  exvecY <- vecY
  
  flag <- TRUE
  k <- 2
  while (flag) {
    
    sigma <- 1/((k - 0.5) * pi)
    if(shape == "gauss") {
      y <- rnorm(1, mean = 0, sd = sigma)
    } else {
      if(shape == "student") { 
        y <- sigma * rt(1, 4)
      } else  {
        if(shape == "chisq") { 
        y <- sigma * rchisq(1, 4)
        } else {
          y <- sigma * rexp(1, 4)  
        }
      }
    }
    
    vecyphi <- sqrt(2) * sin(vect/sigma)
    
    exvecY <- vecY
    vecY <- vecY + y * vecyphi
    flag <- (sum((vecY - exvecY) ^ 2) / sum((vecY) ^ 2) > 0.001)
    k <- k + 1
  }
  vecY
}
```

**Example**: we simulate a sample of 50 functions measured at 100 equidistant points. 

```{r}
nobs <- 50
npoints <- 75
X <- matrix(0, npoints + 25, nobs)
set.seed(777)
for (k in 1:nobs) {
  X[, k] <- simulvec(npoints + 24, shape = "gauss") 
}
```

**Remark:** it is usual to leave aside the first simulated data. Here, we plot the functional data (figure on the left) and we only keep the last 75 values (figure on the right)

```{r, fig.width = 10, fig.height = 4.5}
par(oma = c(0, 0, 0, 0), mar = c(3, 3, 1, 1), las = 1, mfrow = c(1, 2))
matplot(X, type = "l", lty = 1, col ="grey")
abline(v = 25, lty = 2)
X <- X[26:100, ]
matplot(X, type = "l", lty = 1, col ="grey")
```



## Detect all potential clusters

The function `find_all_cluster()` takes as argument the matrix of Cartesian coordinates and returns all potential clusters. It returns a list of size 2. The first element is a list with all the potential clusters and the second element is a list with the complement of the potential clusters. 

If the geographical coordinates are given in Longitude/Latitude, we recommend the user to transform them in an appropriate Coordinate Reference System (see, for instance, https://epsg.io), using package **sf**.  
 
```{r}
find_all_cluster <- function (Matcoord) {
  n <- nrow(Matcoord)
  Matdist <- as.matrix(dist(Matcoord, upper = TRUE))
  vecord_list <- vector("list", n)
  for (k in 1:n) {
    vecord_list[[k]] <- order(Matdist[, k])
  }
  res_cluster_g1 <- vector("list", 0)
  res_cluster_g2 <- vector("list", 0)
  
  matrix_g1 <- vector("list", n)

  for(k in 1:n) {
    matrix_g1[[k]] <- rep(0, n)
  }
  
  nb_combi <- 0
  for (k in 1:(n-1)) {
    for (j in 1:n) {
      temp_1 <- vecord_list[[j]][1:k]
      temp_2 <- vecord_list[[j]][(k+1):n]

      my_vec <- my_vec_2 <-numeric(n)
      my_vec[temp_1] <- 1
      my_vec_2[temp_2] <- 1
      # my_length <- k
      cond_1 <-  any(matrix_g1[[k]] %*% my_vec == k)
      cond_2 <-  any(matrix_g1[[n-k]] %*% my_vec_2 == n-k)
        
      if (!(any(cond_1) | any(cond_2))) {
        nb_combi <- nb_combi + 1
        res_cluster_g1[[nb_combi]] <- temp_1
        res_cluster_g2[[nb_combi]] <- temp_2
        matrix_g1[[k]] <- rbind(matrix_g1[[k]], my_vec)
    #    matrix_g2[[n-k]] <- rbind(matrix_g2[[n-k]], rep(1, n) - my_vec)
      }
    }
  }
  cat("Number of unique combination: ", nb_combi, "\n")
  return(list(vec_g1 = res_cluster_g1,
              vec_g2 = res_cluster_g2))
}
```

**Example:** we consider a random spatial point process with 50 observations.

```{r, fig.width = 5, fig.height = 5}
set.seed(1)
matCoord <- cbind(runif(nobs), runif(nobs))
plot(matCoord, xlab = "x", ylab = "y", asp = 1)
```

We compute all potentially spatial clusters:

```{r}
my_pairs_ex <- find_all_cluster(matCoord)
```

**Remark**: once the potential clusters have been identified, this allows us to test 1644 combinations instead of $50\times 49=2450$.


## Representation of a circle in a map

The function `draw.circle()` returns the coordinates of a circle of radius `radius` and centered around coordinates `x` and `y`:

```{r}
draw.circle <- function (x, y, radius, nv = 100) {
  ymult <- 1
  angle.inc <- 2 * pi/nv
  angles <- seq(0, 2 * pi - angle.inc, by = angle.inc)
  for (circle in 1:length(radius)) {
    xv <- cos(angles) * radius[circle] + x
    yv <- sin(angles) * radius[circle] * ymult + y
  }
  invisible(list(x = xv, y = yv))
}
```

**Example**: we consider a fictitious cluster $C$ of size eight in the data created previously. The cluster is centered around the observation 50 and contains the eight closest observations to observation 50. The radius of the cluster is the distance between the observation 50 and the last observation in the cluster (observation 13).  

```{r}
my_cluster <- c(50, 9, 15, 8, 43, 17, 32, 13)
my_dist <- dist(matCoord[c(50, 13), ])
```


```{r, fig.width = 5, fig.height = 5}
plot(matCoord, xlab = "x", ylab = "y", asp = 1)
points(matCoord[my_cluster, ], pch = 16, col = "red")
temp_plot <- draw.circle(matCoord[50, 1], matCoord[50, 2], 
                         my_dist, nv = 100)
lines(temp_plot, col = "red")
```

For the rest of this section, we modify the functional data of the fictitious cluster. We apply a shift $\Delta_2(t)=ct(1-t)$, with $c=2$. We aim to detect the cluster by using several methods. 

```{r}
t.disc <- (1:75) / (75)
for(k in 1:8)
  X[, my_cluster[k]] <- X[, k] + 2 * t.disc
```

```{r, fig.width = 10, fig.height = 4.5}
par(oma = c(0, 0, 0, 0), mar = c(3, 3, 1, 1), las = 1, mfrow = c(1, 2))
# The function
matplot(X, type = "l", lty = 1, col ="grey")
matplot(X[, my_cluster], type = "l", lty = 1, col ="red", add = T)
# The map
plot(matCoord, xlab = "x", ylab = "y", asp = 1)
points(matCoord[my_cluster, ], pch = 16, col = "red")
lines(temp_plot, col = "red")
```


## Package HDSpatialScan

The function `SpatialScan()` from package `HDSpatialScan` (Frévent et al, 2021) can be used to compute various methods. It takes as main arguments the name(s) of the method, the spatial coordinates and the functional data of the observations. Additional parameters may be used, like the minimum/maximum size of the cluster to be detected, the number of replications to be used to compute the significance, etc. 

For instance, to compute the methods ("DFFSS", "PFSS", "NPFSS"), the function `SpatialScan()` can be used like this:  
 
```{r, eval = F}
fss_result <- HDSpatialScan::SpatialScan(c("NPFSS", "PFSS", "DFFSS"),
        t(X), sites_coord = matCoord, mini = 1, maxi = 49,
        system = "Euclidean", MC=99, typeI = 0.25)
```

For practical reasons, we implement our own functions and use them in our simulation framework.

## Functions for Scan Statistic 

We implement the functions `compute_np()`, `compute_p()`, `compute_dffss()` and 
`compute_h()`, that correspond to the four methods "NPFSS", "PFSS", "DFFSS", "HFSS". Each function takes as argument the list of the possible cluster `c1` (resp. the complement of the possible cluster `c2`) and the functional data `my_mat`. They return the value of the test statistic that was the largest among all the possible combination. They also return the associated cluster. The functions used are available in the file [functions_to_cluster.R](functions_to_cluster.R)  

```{r}
source("codes/functions_to_cluster.R")
```


### Non Parametric NPFSS method

To compute the result of the NPFSS method, we use the function `compute_np()`. We obtain the following value of statistic and associated cluster. 

```{r}
res_np <- compute_np(my_pairs_ex[[1]], my_pairs_ex[[2]], X)
res_np
```

The detected cluster contains `r length(res_np$vec)` observations. To compute the significance, we make $B$ permutations on the data and compute the number of times the scan statistic is lower than the observed one: 

```{r, eval = F}
p_value_np <- 0
B <- 99
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(X))
  MatXsim <- X[, perm]
  temp <- compute_np(my_pairs_ex[[1]], my_pairs_ex[[2]], MatXsim)
  p_value_np <- p_value_np + (res_np$stat < temp$stat)
}
cat("p-value: ", p_value_np / 100)
```

```{r, echo = F}
cat("p-value:  0.33")
```


In this example, the cluster detected by the method NPFSS is not significant.


### Parametric PFSS

To compute the result of the PFSS method, we use the function `compute_p()`. 

```{r}
res_p <- compute_p(my_pairs_ex[[1]], my_pairs_ex[[2]], X)
res_p
```

The cluster detected contains `r length(res_p$vec)` observations. To compute the significance, we make $B$ permutations on the data and compute the number of times the scan statistic is lower than the observed one: 

```{r, eval = F}
p_value_p <- 0
B <- 99
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(X))
  MatXsim <- X[, perm]
  temp <- compute_p(my_pairs_ex[[1]], my_pairs_ex[[2]], MatXsim)
  p_value_p <- p_value_p + (res_p$stat < temp$stat)
}
p_value_p / 100
```

```{r, echo = F}
cat("p-value:  0.03")
```

The detected cluster is significant. Among the `r length(res_p$vec)` observations, `r length(intersect(res_p$vec, my_cluster))` belongs to the real cluster. 

### Method DFFSS

To compute the result of the DFFSS method, we use the function `compute_dffss()`. 

```{r}
res_dffss <- compute_dffss(my_pairs_ex[[1]], my_pairs_ex[[2]], X)
res_dffss
```

The detected cluster contains `r length(res_dffss$vec)` observations. To compute the significance, we make $B$ permutations on the data and compute the number of times the scan statistic is lower than the observed one: 

```{r, eval = F}
p_value_dffss <- 0
B <- 99
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(X))
  MatXsim <- X[, perm]
  temp <- compute_dffss(my_pairs_ex[[1]], my_pairs_ex[[2]], MatXsim)
  p_value_dffss <- p_value_dffss + (res_dffss$stat < temp$stat)
}
p_value_dffss / 100
```


```{r, echo = F}
cat("p-value:  0.03")
```

The detected cluster is significant. Among the `r length(res_dffss$vec)` observations, `r length(intersect(res_dffss$vec, my_cluster))` belongs to the real cluster. 

### Method Horvath/Hotelling

To compute the result of the Horvath/Hotelling method, we use the function `compute_h()`. An additional argument allows to select the value of $d$, corresponding to the number of eigen vectors in the formula $\displaystyle\sum_{k=1}^d\frac{a_k^2}{\lambda_k}$. 

The choice of $d$ has a strong role on the result. For instance, if $d$ equals the number of measurements (i.e. 75 in our example), we remark that the statistic test is abnormally huge and the detected cluster contains only one observation  (which is a False Positive). 

This is due to numerical precision errors obtained for the small eigen values. Indeed, when $k$ becomes large, we divide a term very small (and sometimes even negative like in the order of $-10^{-16}$) by another term also very small. In that case, the numerical precision errors explain why we obtain a result that can tend to +/- infinity.    


One solution proposed for choosing the value $d$ is to use the cumulative percentage of total variance (CPV) criteria (see for instance, Joseph, Galeano and Lilo, 2015. The CPV is defined as follow : $CPV(k)=\frac{\displaystyle\sum_{j=1}^k\lambda_j}{\displaystyle\sum_{j=1}^{npoints}\lambda_j}$. Then, we select the value of $d$ as the value of $k$ from which the function CPV grows very slowly to 1.

The argument `plot_eigen` allows to plot CPV of the function `compute_h()`. Here, we recommend choosing $d=5$ which leads to explain more than $99.8\%$. 

**Example** when $d=npoints$:
```{r, fig.width = 6, fig.height = 3}
res_h <- compute_h(my_pairs_ex[[1]], my_pairs_ex[[2]], X, d = npoints,
                           plot_eigen = T)
res_h
```

**Example** when $d=5$:
```{r}
res_h <- compute_h(my_pairs_ex[[1]], my_pairs_ex[[2]], X, d = 5)
res_h
```

To compute the significance, we make $B$ permutations on the data and compute the number of times the scan statistic is lower than the observed one: 

```{r, eval = F}
p_value_h <- 0
B <- 99
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(X))
  MatXsim <- X[, perm]
  temp <- compute_h(my_pairs_ex[[1]], my_pairs_ex[[2]], MatXsim,
                               k = 5)
  p_value_h <- p_value_h + (res_h$stat < temp$stat)
}
p_value_h / 100
```

```{r, echo = F}
cat("p-value:  0.02")
```

The detected cluster is significant. Among the `r length(res_h$vec)` observations, `r length(intersect(res_h$vec, my_cluster))` belongs to the real cluster.


**Improving computational time**: if `plot_eigen=T`, the spectral decomposition of the covariance matrix is done with the function `eigen()`. It means that all the eigen values/eigen vectors are computed. If `plot_eigen=F`, we use the function `eigs_sym()` from package **rARPACK** that allows to compute only the $d$ first eigen values/eigen vectors. It improves the computation time by a factor 3. 



# Simulation part

## The spatial data

We import first the contours of the French departments:

```{r, warning = F, message = F}
dep <- read_sf("data/departements.geojson")
dep <- dep[!dep$code %in% c("2A", "2B"), ]
dep <- dep[order(dep$code), ]
my_region <- st_union(dep)
nc_osm <- get_tiles(dep, provider = "Esri.WorldShadedRelief", 
                      zoom = 7, crop = T)
```

Then, we compute the Cartesian coordinates of the centroids of the departments in the official French Coordinates Reference System (CRS 2154), such that the distances between locations are computed in meters.

```{r, warning = F, message = F}
my_proj <- 2154
dep_proj <- st_transform(dep, 2154)
Matcoord <- st_coordinates(st_centroid(dep_proj))
dist_proj <- as(dist(Matcoord), "matrix")
```

We show the departments located around the city of Paris (74, 92, 91, 93, 77, 90, 94, 76) that will be simulated differently from the rest of the departments, such that they represent the cluster we would like to detect.

```{r, warning = F, message = F}
cols = c("#D35C79", "#009593")
# id of the cluster
vecclus <- c(74, 92, 91, 93, 77, 90, 94, 76)
# graphical parameters
col_geo <- rep(rgb(0.9, 0.9, 0.9, alpha = 0.1), nrow(Matcoord))
col_geo[vecclus] <- alpha(cols[1], 0.8)
```

```{r, warning = F, message = F}
#pdf("figures/french_cluster.pdf", width = 7, height = 7)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
plot_tiles(nc_osm, adjust = T)
mf_shadow(st_geometry(st_union(dep[vecclus, ])), add = T, cex = 0.5)
plot(st_geometry(dep), border = "white", col = col_geo,  
     add = T, lwd = 0.1)
plot(st_geometry(my_region), add = T, lwd = 0.5)
#text(par()$usr[1] + 0.03 * (par()$usr[2] - par()$usr[1]), 
#     par()$usr[4] - 0.07 * (par()$usr[4] - par()$usr[3]), 
#     labels = "A)", pos = 4, cex = 2)
#dev.off()
```

## The different shifts/probabilistic models

We consider three types of shift:

* $\Delta_1(t)=ct$,
* $\Delta_2(t)=ct(1-t)$,
* $\Delta_3(t)=c\exp(-100(t-0.5)^2)/3$,

and four different probabilistic models $Z_{i,k}/\sigma_k$:

* a Gaussian $N(0,1)$,
* a Student $t(4)$
* a Chi $\chi^2(4)$
* an Exponential $e(4)$

with $i=1,\ldots,94$ and $t=1,\ldots,200$. The first 100 simulations are not kept to give time for the process to converge. For each combination shift/probabilistic, we consider different values of $c$. 

```{r}
nobs <- nrow(Matcoord)
npoints <- 100
ndrop <- 100
t.disc <- (1:(npoints)) / (npoints) 
veccluster <- rep(0, nobs)
veccluster[vecclus] <- 1 
```

  
For $c=3$, we plot an example of simulations for each of the combination shift/probabilistic model:   

```{r}
alpha <- 10
X_aggregated <- data.frame(
  x = integer(),
  y = numeric(), 
  shift = character(),
  proba = character()
) 
for(type_shift in 2:4) {
  for(shape in c("gauss", "student", "chisq", "exp")) {
    X <- matrix(0, npoints+ndrop, nobs)
    for (k in 1:nobs) {
      X[, k] <- simulvec(npoints+(ndrop-1), shape = shape) 
      if(type_shift == 1) {
        X[(ndrop+1):nrow(X), k] <- X[(ndrop+1):nrow(X), k] + alpha * (veccluster[k] == 1)
        } else {
          if (type_shift == 2) {
            alpha <- 3
            X[(ndrop+1):nrow(X), k] <- X[(ndrop+1):nrow(X), k] + alpha * t.disc * (veccluster[k] == 1)
            } else {
              if (type_shift == 3) {
                alpha <- 10
                X[(ndrop+1):nrow(X), k] <- X[(ndrop+1):nrow(X), k] + alpha * t.disc * (1 - t.disc) * (veccluster[k] == 1)
                } else {
                  alpha <- 10
                  X[(ndrop+1):nrow(X), k] <- X[(ndrop+1):nrow(X), k] + alpha * exp(-100 * (t.disc - 0.5) ^ 2) / 3 * (veccluster[k] == 1)
                }
            }
        }
    }
    # drop first observation 
    X <- X[-(1:ndrop), ]
    # aggregate data 
    X_aggregated <- rbind(X_aggregated,
                          data.frame(
                            x = seq(0, 1, length.out = 100),
                            y = as.vector(X), 
                            shift = type_shift,
                            proba = shape,
                            id = rep(1:94, each = 100)
                          ))
  }
}
X_aggregated$shift <- factor(X_aggregated$shift) 
levels(X_aggregated$shift) = c(`2` = TeX("$\\Delta_1(t)$"), 
                               `3` = TeX("$\\Delta_2(t)$"), 
                               `4` = TeX("$\\Delta_3(t)$"))
X_aggregated$Cluster <- factor(ifelse(X_aggregated$id %in% vecclus, "Yes", "No"),
                               levels = c("No", "Yes"))
X_aggregated$proba <- factor(X_aggregated$proba, levels = c("gauss", "student", "chisq", "exp"))
levels(X_aggregated$proba) <- c(`gauss` = TeX("$N(0,1)$"), 
                                `student` = TeX("$t(4)$"), 
                                `exp` = TeX("$Exp(4)$"),
                                `chisq` = TeX("$\\chi^2(4)$"))
```

* $\Delta_1(t)=ct$,
* $\Delta_2(t)=ct(1-t)$,
* $\Delta_3(t)=c\exp(-100(t-0.5)^2)/3$,

```{r}
temp <- X_aggregated[order(as.numeric(X_aggregated$Cluster), X_aggregated$id),]
X_aggregated %>%
  ggplot(aes(x = x, y = y, color = Cluster)) +
  geom_line(aes(group = id)) +
  theme_bw() +
  theme(strip.background = element_rect(color = "black", fill = alpha("#EE9E94", 0.1))) +
  facet_grid(rows=vars(proba),
  cols=vars(shift),
  labeller=label_parsed,
  scales = "free") +
      scale_colour_manual(values = c("grey", "tomato2")) +
  xlab(TeX("$t$")) +
  ylab(TeX("$X_i(t)$"))
ggsave("figures/simu.pdf", width = 10, height = 8)
```

## Results

We used a server with 94 cores and launched on each core a unique parameter set shift/probabilistic model/value of $c$. The computation time was around 7 days. We repeated the following procedure on each core:

```{r}
parms_df <- rbind(
  # line 1 
  data.frame(
    shape = "gauss", type_shift = 2, alpha = seq(0, 3, 0.5), sizeclust = 8),
  data.frame(
    shape = "gauss", type_shift = 3, alpha = seq(0, 10.5, 1.5), sizeclust = 8),
  data.frame(
    shape = "gauss", type_shift = 4, alpha = seq(0, 12, 2), sizeclust = 8),  
  # line 2
  data.frame(
    shape = "student", type_shift = 2, alpha = seq(0, 7, 1), sizeclust = 8),
  data.frame(
    shape = "student", type_shift = 3, alpha = seq(0, 14, 2), sizeclust = 8),
  data.frame(
    shape = "student", type_shift = 4, alpha = seq(0, 14, 2), sizeclust = 8),  
  # line 3
  data.frame(
    shape = "chisq", type_shift = 2, alpha = seq(0, 14, 2), sizeclust = 8),
  data.frame(
    shape = "chisq", type_shift = 3, alpha = seq(0, 14, 2), sizeclust = 8),
  data.frame(
    shape = "chisq", type_shift = 4, alpha = seq(0, 14, 2), sizeclust = 8),  
  # line 4
  data.frame(
    shape = "exp", type_shift = 2, alpha = seq(0, 7, 1), sizeclust = 8),
  data.frame(
    shape = "exp", type_shift = 3, alpha = seq(0, 14, 2), sizeclust = 8),
  data.frame(
    shape = "exp", type_shift = 4, alpha = seq(0, 14, 2), sizeclust = 8)
)
```

Repeat 200 times:

* simulate a set of functional data (with a shift on the Paris region),
* compute the scan statistic for the following methods: "HFSS", "PFSS", "NPFSS", "DFFSS", 
* draw 199 permutation samples, and compute the scan statistics on each of them to obtain significance.

From the significance and the detected clusters, we compute:

* the power,
* the percentage of True Positive,
* the percentage of False Negative.

For the Horvath/Hotelling method, for reasons of simplification, we fix the value of $d=5$ which leads to explain more or less $99\%$ of the variance. 

The codes used are given in the files [batch_cluster_size_8.R](codes/batch_cluster_size_8.R), by using the functions in [functions_to_cluster.R](codes/functions_to_cluster.R).


```{r, echo = F}
# we load the results obtained from the servor 
load("results/my_res_new_1.RData")
load("results/my_res_new_2.RData")
load("results/my_res_new_3.RData")
load("results/my_res_new_4.RData")
```

We present the results in a format that can be easily represented in a graph:

```{r}
power_to_plot <- data.frame(
  shape = character(0),
  type_shift = integer(0),
  alpha = numeric(0)
)
nb_est <- 10
for(k in 1:nrow(parms_df)) {
  power_to_plot <- rbind(
    power_to_plot,
    parms_df[rep(k, nb_est), 1:3])
}
power_to_plot$method <- c("DFFSS", "PFSS", "NPFSS", "hotelling_1", "hotelling_2", 
                          "hotelling_3", "hotelling_4", "HFSS", "hotelling_10", 
                          "hotelling_15")
FP_to_plot <- TP_to_plot <- power_to_plot

for(k in 1:length(res_par_1)) {
  power_to_plot$value[(1:nb_est)+(k-1)*nb_est] <- (res_par_1[[k]]$power + res_par_2[[k]]$power + 
                                                     res_par_3[[k]]$power + res_par_4[[k]]$power) / 200 # res.final$power / 100 # 
  TP_to_plot$value[(1:nb_est)+(k-1)*nb_est] <- (res_par_1[[k]]$nTP + res_par_2[[k]]$nTP + 
                                                  res_par_3[[k]]$nTP + res_par_4[[k]]$nTP) /  
    (res_par_1[[k]]$power + res_par_2[[k]]$power + res_par_3[[k]]$power + res_par_4[[k]]$power) / 8 # res.final$nTP / res.final$power / 8 # 
  FP_to_plot$value[(1:nb_est)+(k-1)*nb_est] <- (res_par_1[[k]]$nFP + res_par_2[[k]]$nFP + 
                                                  res_par_3[[k]]$nFP + res_par_4[[k]]$nFP) /  
    (res_par_1[[k]]$power + res_par_2[[k]]$power + res_par_3[[k]]$power + res_par_4[[k]]$power) / 86 #res.final$nFP / res.final$power / 8 # res_par[[k]]$nFP / res_par[[k]]$power / 8
}
power_to_plot$criteria <- "power"
TP_to_plot$criteria <- "TP"
FP_to_plot$criteria <- "FP"
to_plot <- rbind(power_to_plot, TP_to_plot, FP_to_plot)
# we select the most interseting points
parms_df_select <- rbind(
  # line 1 
  data.frame(
    shape = "gauss", type_shift = 2, alpha = seq(0, 3, 0.5)[-2], sizeclust = 8),
  data.frame(
    shape = "gauss", type_shift = 3, alpha = seq(0, 10.5, 1.5)[-c(2, 8)], sizeclust = 8),
  data.frame(
    shape = "gauss", type_shift = 4, alpha = seq(0, 12, 2)[-2], sizeclust = 8),  
  # line 2
  data.frame(
    shape = "student", type_shift = 2, alpha = seq(0, 7, 1)[-c(2, 8)], sizeclust = 8),
  data.frame(
    shape = "student", type_shift = 3, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 8),
  data.frame(
    shape = "student", type_shift = 4, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 8),  
  # line 3
  data.frame(
    shape = "chisq", type_shift = 2, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 8),
  data.frame(
    shape = "chisq", type_shift = 3, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 8),
  data.frame(
    shape = "chisq", type_shift = 4, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 8),  
  # line 4
  data.frame(
    shape = "exp", type_shift = 2, alpha = seq(0, 7, 1)[-c(2, 8)], sizeclust = 8),
  data.frame(
    shape = "exp", type_shift = 3, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 8),
  data.frame(
    shape = "exp", type_shift = 4, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 8)
)
to_plot <- merge(parms_df_select, to_plot, by = c("shape", "type_shift", "alpha"))
# we give labels
to_plot$type_shift <- factor(to_plot$type_shift) 
levels(to_plot$type_shift) = c(`2` = TeX("$\\Delta_1(t)$"), # TeX("$\\Delta_1(t)=ct$"), 
                         `3` = TeX("$\\Delta_2(t)$"),  # TeX("$\\Delta_2(t)=ct(1-t)$"), 
                         `4` = TeX("$\\Delta_3(t)$")) #  TeX("$\\Delta_3(t)=\\frac{c}{3}\\exp(-100(t-0.5)^2)$"))
to_plot$shape <- factor(to_plot$shape, levels = c("gauss", "student", "exp", "chisq"))
levels(to_plot$shape) <- c(`gauss` = TeX("$N(0,1)$"), 
                                `student` = TeX("$t(4)$"), 
                                `exp` = TeX("$Exp(4)$"),
                                `chisq` = TeX("$\\chi^2(4)$"))
to_plot$method <- factor(to_plot$method, c("HFSS", "DFFSS", "PFSS", "NPFSS"))
```

### Choice of $K$

```{r}
shift_levels <- levels(X_aggregated$shift)
proba_levels <- levels(X_aggregated$proba)
par(oma = c(0, 0, 0, 0), mar = c(3.5, 3, 1, 1),  mfrow = c(4, 3),
        mgp = c(2, 1, 0), las = 1)
for (i in 1:length(shift_levels)) {
  for (j in 1:length(proba_levels)) {
temp <- X_aggregated %>%
  filter(shift == shift_levels[i], proba == proba_levels[j]) %>%
  select(y, id) %>% 
  mutate(id = paste0("obs_", id)) %>%
  pivot_wider(names_from = 2, values_from = 1, values_fn = list) |> 
              unnest(cols = everything() ) %>%
  as.matrix()
compute_h(my_pairs[[1]], my_pairs[[2]], temp, d = 3, plot_eigen = T)
abline(h = 0.995, lty = 2, col = "red")
  }
}
```



### Power 

```{r}
to_plot %>%
  filter(criteria == "power") %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot %>% 
              filter(criteria == "power") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")), 
            aes(group = method)) +
  geom_point(size = 0.8, pch = 15) +
  ggh4x::facet_grid2(rows=vars(shape),
  cols=vars(type_shift),
  labeller=label_parsed, scales = "free_x", independent = "x")  +
  theme_bw()
# ggsave("figures/simu_8_power.pdf", width = 9, height = 8)
```

When the probabilistic model is from the Exponential distribution, the method "HFSS" dramatically improves the power for any shift. When the probabilistic distribution belongs to Normal, Student or $\chi^2$, the method "HFSS" behaves slightly better for $\Delta_1$ and much better for $\Delta_2$ and $\Delta_3$. 

### True Positive 

```{r}
to_plot %>%
  filter(criteria == "TP") %>%
  filter(alpha != 0) %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot %>% 
              filter(criteria == "TP") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
              filter(alpha != 0), 
            aes(group = method)) +
  geom_point(size = 1, pch = 15, alpha = 0) +
  geom_point(data = to_plot %>% 
              filter(criteria == "TP") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
              filter(alpha != 0),
             size = 0.8, pch = 15) +
  ggh4x::facet_grid2(rows=vars(shape),
  cols=vars(type_shift),
  labeller=label_parsed, scales = "free_x", independent = "x") +
  theme_bw()
#ggsave("figures/simu_8_TP.pdf", width = 9, height = 8)
```

The percentage of True positive with the method "HFSS", tends to be better for any situations, excepted for $\Delta_1$, when the probabilistic model is Gaussian or Student. In that case, the percentage of TP is slightly for small values of $c$.


### False Positive 

```{r}
to_plot %>%
  filter(criteria == "FP") %>%
  filter(alpha != 0) %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot %>% 
              filter(criteria == "FP") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
              filter(alpha != 0), 
            aes(group = method)) +
  geom_point(size = 1, pch = 15, alpha = 0) +
  geom_point(data = to_plot %>% 
              filter(criteria == "FP") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
              filter(alpha != 0),
             size = 0.8, pch = 15) +
  ggh4x::facet_grid2(rows=vars(shape),
  cols=vars(type_shift),
  labeller=label_parsed, scales = "free_x", independent = "x")  +
  theme_bw()
#ggsave("figures/simu_8_FP.pdf", width = 9, height = 8)
```


The percentage of False positive with the method "HFSS", tends to be better for any shift/probabilistic model. 



### Delta 1 

```{r}
threshold <- data.frame(
  yintercept = c(0.05, NA, NA), 
  criteria = factor(c("power", "TP", "FP"), levels = c("power", "TP", "FP")))
to_plot$criteria <- factor(to_plot$criteria, levels = c("power", "TP", "FP"))
to_plot2 <- to_plot[-which(to_plot$alpha == 0 & to_plot$criteria == "TP"), ]
to_plot2 <- to_plot2[-which(to_plot2$alpha == 0 & to_plot2$criteria == "FP"), ]
to_plot2 %>%
  filter(type_shift == "Delta[1](t)") %>%
  filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot2 %>% 
              filter(type_shift == "Delta[1](t)") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")), 
            aes(group = method)) +
  geom_point(size = 0.8, pch = 15) +
  geom_hline(data = to_plot2 %>% filter(criteria == "power"), 
               aes(yintercept = 0.05, 
                 linetype = "0.05")) +
  ggh4x::facet_grid2(rows=vars(criteria),
                     cols=vars(shape),
                     labeller=label_parsed,
                     scales = "free", 
                     axes = "margins")  +
  theme_bw() +
  xlab(TeX("$\\alpha$")) +
  ylab("") +
  ggtitle(TeX("$\\Delta_1$")) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_linetype_manual(name = "threshold", values = 2) +
  theme(strip.background = element_rect(colour = "black", 
        fill = alpha("#EE9E94", 0.1)))
# ggsave("figures/simu_8_delta_1.pdf", width = 8, height = 6.5)
```


### Delta 2 

```{r}
threshold <- data.frame(
  yintercept = c(0.05, NA, NA), 
  criteria = factor(c("power", "TP", "FP"), levels = c("power", "TP", "FP")))
to_plot$criteria <- factor(to_plot$criteria, levels = c("power", "TP", "FP"))
to_plot2 <- to_plot[-which(to_plot$alpha == 0 & to_plot$criteria == "TP"), ]
to_plot2 <- to_plot2[-which(to_plot2$alpha == 0 & to_plot2$criteria == "FP"), ]
to_plot2 %>%
  filter(type_shift == "Delta[2](t)") %>%
   # filter(alpha != 0 & criteria == "TP") %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot2 %>% 
              filter(type_shift == "Delta[2](t)") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")), 
            aes(group = method)) +
  geom_point(size = 0.8, pch = 15) +
#  geom_hline(aes(yintercept = yintercept, 
#                 linetype = criteria), data = threshold,
#             linetype = 2) +
    geom_hline(data = to_plot2 %>% filter(criteria == "power"), 
               aes(yintercept = 0.05, 
                 linetype = "0.05")) +
  ggh4x::facet_grid2(rows=vars(criteria),
                     cols=vars(shape),
                     labeller=label_parsed,
                     scales = "free", 
                     axes = "margins")  +
  theme_bw() +
  xlab(TeX("$\\alpha$")) +
  ylab("") +
  ggtitle(TeX("$\\Delta_2$")) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_linetype_manual(name = "threshold", values = 2) +
  theme(strip.background = element_rect(colour = "black", 
        fill = alpha("#EE9E94", 0.1)))
# ggsave("figures/simu_8_delta_2.pdf", width = 8, height = 6.5)
```


### Delta 3

```{r}
threshold <- data.frame(
  yintercept = c(0.05, NA, NA), 
  criteria = factor(c("power", "TP", "FP"), levels = c("power", "TP", "FP")))
to_plot$criteria <- factor(to_plot$criteria, levels = c("power", "TP", "FP"))
to_plot2 <- to_plot[-which(to_plot$alpha == 0 & to_plot$criteria == "TP"), ]
to_plot2 <- to_plot2[-which(to_plot2$alpha == 0 & to_plot2$criteria == "FP"), ]
to_plot2 %>%
  filter(type_shift == "Delta[3](t)") %>%
   # filter(alpha != 0 & criteria == "TP") %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot2 %>% 
              filter(type_shift == "Delta[3](t)") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")), 
            aes(group = method)) +
  geom_point(size = 0.8, pch = 15) +
#  geom_hline(aes(yintercept = yintercept, 
#                 linetype = criteria), data = threshold,
#             linetype = 2) +
    geom_hline(data = to_plot2 %>% filter(criteria == "power"), 
               aes(yintercept = 0.05, 
                 linetype = "0.05")) +
  ggh4x::facet_grid2(rows=vars(criteria),
                     cols=vars(shape),
                     labeller=label_parsed,
                     scales = "free", 
                     axes = "margins")  +
  theme_bw() +
  xlab(TeX("$\\alpha$")) +
  ylab("") +
  ggtitle(TeX("$\\Delta_3$")) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_linetype_manual(name = "threshold", values = 2) +
  theme(strip.background = element_rect(colour = "black", 
        fill = alpha("#EE9E94", 0.1)))
# ggsave("figures/simu_8_delta_3.pdf", width = 8, height = 6.5)
```


## Checking Robustess

We apply the same procedure by varying the size of the cluster. We consider a cluster of size 10. We added two more departments to the previous list:  59, 27


```{r, warning = F, message = F}
# id of the cluster
vecclus <- c(74, 92, 91, 93, 77, 90, 94, 76, 59, 27)
# graphical parameters
col_geo <- rep(rgb(0.9, 0.9, 0.9, alpha = 0.1), nrow(Matcoord))
col_geo[vecclus] <- alpha("#D35C79", 0.8)
```

```{r, warning = F, message = F}
#pdf("figures/french_cluster_10.pdf", width = 7, height = 7)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
plot_tiles(nc_osm, adjust = T)
mf_shadow(st_geometry(st_union(dep[vecclus, ])), add = T, cex = 0.5)
plot(st_geometry(dep), border = "white", col = col_geo,  
     add = T, lwd = 0.1)
plot(st_geometry(my_region), add = T, lwd = 0.5)
#text(par()$usr[1] + 0.03 * (par()$usr[2] - par()$usr[1]), 
#     par()$usr[4] - 0.07 * (par()$usr[4] - par()$usr[3]), 
#     labels = "B)", pos = 4, cex = 2)
#dev.off()
```

The interpretations are the same as in the previous section.

```{r, echo = F}
# we load the results obtained from the servor 
load("results/my_res_new_10_1.RData")
load("results/my_res_new_10_2.RData")
load("results/my_res_new_10_3.RData")
load("results/my_res_new_10_4.RData")
```

```{r}
power_to_plot <- data.frame(
  shape = character(0),
  type_shift = integer(0),
  alpha = numeric(0)
)
nb_est <- 10
for(k in 1:nrow(parms_df)) {
  power_to_plot <- rbind(
    power_to_plot,
    parms_df[rep(k, nb_est), 1:3])
}
power_to_plot$method <- c("DFFSS", "PFSS", "NPFSS", "hotelling_1", "hotelling_2", 
                          "hotelling_3", "hotelling_4", "HFSS", "hotelling_10", 
                          "hotelling_15")
FP_to_plot <- TP_to_plot <- power_to_plot

for(k in 1:length(res_par_1)) {
  power_to_plot$value[(1:nb_est)+(k-1)*nb_est] <- (
    res_par_1[[k]]$power + res_par_2[[k]]$power +  
    res_par_3[[k]]$power  + res_par_4[[k]]$power) / 200 
  TP_to_plot$value[(1:nb_est)+(k-1)*nb_est] <- (
    res_par_1[[k]]$nTP + res_par_2[[k]]$nTP + 
      res_par_3[[k]]$nTP + res_par_4[[k]]$nTP) /  
    (res_par_1[[k]]$power + res_par_2[[k]]$power + 
       res_par_3[[k]]$power + res_par_4[[k]]$power) / 8 
  FP_to_plot$value[(1:nb_est)+(k-1)*nb_est] <- (
    res_par_1[[k]]$nFP + res_par_2[[k]]$nFP + 
      res_par_3[[k]]$nFP + res_par_4[[k]]$nFP) /  
    (res_par_1[[k]]$power + res_par_2[[k]]$power + 
       res_par_3[[k]]$power + res_par_4[[k]]$power) / 86 
}
power_to_plot$criteria <- "power"
TP_to_plot$criteria <- "TP"
FP_to_plot$criteria <- "FP"
to_plot <- rbind(power_to_plot, TP_to_plot, FP_to_plot)
# we select the most interseting points
parms_df_select <- rbind(
  # line 1
  data.frame(
    shape = "gauss", type_shift = 2, alpha = seq(0, 3, 0.5)[-c(2)], sizeclust = 10),
  data.frame(
    shape = "gauss", type_shift = 3, alpha = seq(0, 10.5, 1.5)[-c(2, 8)], sizeclust = 10),
  data.frame(
    shape = "gauss", type_shift = 4, alpha = seq(0, 12, 2)[-c(2, 8)], sizeclust = 10),
  # line 2
  data.frame(
    shape = "student", type_shift = 2, alpha = seq(0, 7, 1)[-c(2, 8)], sizeclust = 10),
  data.frame(
    shape = "student", type_shift = 3, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 10),
  data.frame(
    shape = "student", type_shift = 4, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 10),
  # line 3
  data.frame(
    shape = "chisq", type_shift = 2, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 10),
  data.frame(
    shape = "chisq", type_shift = 3, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 10),
  data.frame(
    shape = "chisq", type_shift = 4, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 10),
  # line 4
  data.frame(
    shape = "exp", type_shift = 2, alpha = seq(0, 7, 1)[-c(2, 8)], sizeclust = 10),
  data.frame(
    shape = "exp", type_shift = 3, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 10),
  data.frame(
    shape = "exp", type_shift = 4, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 10)
)
to_plot <- merge(parms_df_select, to_plot, by = c("shape", "type_shift", "alpha"))
# we give labels
to_plot$type_shift <- factor(to_plot$type_shift) 
levels(to_plot$type_shift) = c(`2` = TeX("$\\Delta_1(t)$"), # TeX("$\\Delta_1(t)=ct$"), 
                         `3` = TeX("$\\Delta_2(t)$"),  # TeX("$\\Delta_2(t)=ct(1-t)$"), 
                         `4` = TeX("$\\Delta_3(t)$")) #  TeX("$\\Delta_3(t)=\\frac{c}{3}\\exp(-100(t-0.5)^2)$"))
to_plot$shape <- factor(to_plot$shape, levels = c("gauss", "student", "exp", "chisq"))
levels(to_plot$shape) <- c(`gauss` = TeX("$N(0,1)$"), 
                                `student` = TeX("$t(4)$"), 
                                `exp` = TeX("$Exp(4)$"),
                                `chisq` = TeX("$\\chi^2(4)$"))
to_plot$method <- factor(to_plot$method, c("HFSS", "DFFSS", "PFSS", "NPFSS"))
```

### Power 

```{r}
to_plot %>%
  filter(criteria == "power") %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot %>% 
              filter(criteria == "power") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")), 
            aes(group = method)) +
  geom_point(size = 0.8, pch = 15) +
  ggh4x::facet_grid2(rows=vars(shape),
  cols=vars(type_shift),
  labeller=label_parsed, scales = "free_x", independent = "x")
#ggsave("figures/simu_10_power.pdf", width = 9, height = 8)
```

### True Positive 

```{r}
to_plot %>%
  filter(criteria == "TP") %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot %>% 
              filter(criteria == "TP") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
              filter(alpha != 0), 
            aes(group = method)) +
  geom_point(size = 1, pch = 15, alpha = 0) +
  geom_point(data = to_plot %>% 
              filter(criteria == "TP") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
              filter(alpha != 0),
             size = 0.8, pch = 15) +
  ggh4x::facet_grid2(rows=vars(shape),
  cols=vars(type_shift),
  labeller=label_parsed, scales = "free_x", independent = "x")
#ggsave("figures/simu_10_TP.pdf", width = 9, height = 8)
```

### False Positive 

```{r}
to_plot %>%
  filter(criteria == "FP") %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot %>% 
              filter(criteria == "FP") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
              filter(alpha != 0), 
            aes(group = method)) +
  geom_point(size = 1, pch = 15, alpha = 0) +
  geom_point(data = to_plot %>% 
              filter(criteria == "FP") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
              filter(alpha != 0),
             size = 0.8, pch = 15) +
  ggh4x::facet_grid2(rows=vars(shape),
  cols=vars(type_shift),
  labeller=label_parsed, scales = "free_x", independent = "x")
#ggsave("figures/simu_10_FP.pdf", width = 9, height = 8)
```

### Delta 1 

```{r}
threshold <- data.frame(
  yintercept = c(0.05, NA, NA), 
  criteria = factor(c("power", "TP", "FP"), levels = c("power", "TP", "FP")))
to_plot$criteria <- factor(to_plot$criteria, levels = c("power", "TP", "FP"))
to_plot2 <- to_plot[-which(to_plot$alpha == 0 & to_plot$criteria == "TP"), ]
to_plot2 <- to_plot2[-which(to_plot2$alpha == 0 & to_plot2$criteria == "FP"), ]
to_plot2 %>%
  filter(type_shift == "Delta[1](t)") %>%
   # filter(alpha != 0 & criteria == "TP") %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot2 %>% 
              filter(type_shift == "Delta[1](t)") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")), 
            aes(group = method)) +
  geom_point(size = 0.8, pch = 15) +
#  geom_hline(aes(yintercept = yintercept, 
#                 linetype = criteria), data = threshold,
#             linetype = 2) +
    geom_hline(data = to_plot2 %>% filter(criteria == "power"), 
               aes(yintercept = 0.05, 
                 linetype = "0.05")) +
  ggh4x::facet_grid2(rows=vars(criteria),
                     cols=vars(shape),
                     labeller=label_parsed,
                     scales = "free", 
                     axes = "margins")  +
  theme_bw() +
  xlab(TeX("$\\alpha$")) +
  ylab("") +
  ggtitle(TeX("$\\Delta_1$")) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_linetype_manual(name = "threshold", values = 2) +
  theme(strip.background = element_rect(colour = "black", 
        fill = alpha("#EE9E94", 0.1)))
#ggsave("figures/simu_10_delta_1.pdf", width = 8, height = 6.5)
```


### Delta 2 

```{r}
threshold <- data.frame(
  yintercept = c(0.05, NA, NA), 
  criteria = factor(c("power", "TP", "FP"), levels = c("power", "TP", "FP")))
to_plot$criteria <- factor(to_plot$criteria, levels = c("power", "TP", "FP"))
to_plot2 <- to_plot[-which(to_plot$alpha == 0 & to_plot$criteria == "TP"), ]
to_plot2 <- to_plot2[-which(to_plot2$alpha == 0 & to_plot2$criteria == "FP"), ]
to_plot2 %>%
  filter(type_shift == "Delta[2](t)") %>%
   # filter(alpha != 0 & criteria == "TP") %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot2 %>% 
              filter(type_shift == "Delta[2](t)") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")), 
            aes(group = method)) +
  geom_point(size = 0.8, pch = 15) +
#  geom_hline(aes(yintercept = yintercept, 
#                 linetype = criteria), data = threshold,
#             linetype = 2) +
    geom_hline(data = to_plot2 %>% filter(criteria == "power"), 
               aes(yintercept = 0.05, 
                 linetype = "0.05")) +
  ggh4x::facet_grid2(rows=vars(criteria),
                     cols=vars(shape),
                     labeller=label_parsed,
                     scales = "free", 
                     axes = "margins")  +
  theme_bw() +
  xlab(TeX("$\\alpha$")) +
  ylab("") +
  ggtitle(TeX("$\\Delta_2$")) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_linetype_manual(name = "threshold", values = 2) +
  theme(strip.background = element_rect(colour = "black", 
        fill = alpha("#EE9E94", 0.1)))
#ggsave("figures/simu_10_delta_2.pdf", width = 8, height = 6.5)
```


### Delta 3

```{r}
threshold <- data.frame(
  yintercept = c(0.05, NA, NA), 
  criteria = factor(c("power", "TP", "FP"), levels = c("power", "TP", "FP")))
to_plot$criteria <- factor(to_plot$criteria, levels = c("power", "TP", "FP"))
to_plot2 <- to_plot[-which(to_plot$alpha == 0 & to_plot$criteria == "TP"), ]
to_plot2 <- to_plot2[-which(to_plot2$alpha == 0 & to_plot2$criteria == "FP"), ]
to_plot2 %>%
  filter(type_shift == "Delta[3](t)") %>%
   # filter(alpha != 0 & criteria == "TP") %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot2 %>% 
              filter(type_shift == "Delta[3](t)") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")), 
            aes(group = method)) +
  geom_point(size = 0.8, pch = 15) +
#  geom_hline(aes(yintercept = yintercept, 
#                 linetype = criteria), data = threshold,
#             linetype = 2) +
    geom_hline(data = to_plot2 %>% filter(criteria == "power"), 
               aes(yintercept = 0.05, 
                 linetype = "0.05")) +
  ggh4x::facet_grid2(rows=vars(criteria),
                     cols=vars(shape),
                     labeller=label_parsed,
                     scales = "free", 
                     axes = "margins")  +
  theme_bw() +
  xlab(TeX("$\\alpha$")) +
  ylab("") +
  ggtitle(TeX("$\\Delta_3$")) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_linetype_manual(name = "threshold", values = 2) +
  theme(strip.background = element_rect(colour = "black", 
        fill = alpha("#EE9E94", 0.1)))
#ggsave("figures/simu_10_delta_3.pdf", width = 8, height = 6.5)
```


# Empirical part

## Spanish region 

The file [spain_unemp.RData](spain_unemp.RData) contains three objects:

* `Matcoordalpha`, the coordinates of the centroid of the Spanish regions (expressed in Cartesian coordinates),
* `MatX`, the unemployment evolution of the Spanish regions across 80 quarters from 2002 to 2022,
* `region_spain`, the spatial contours of the Spanish regions.

```{r}
load("data/spain_unemp.RData")
dates <- seq(2002, 2021.75, by = 0.25)
y_lim <- range(MatX)
# import the OSM map
nc_osm <- get_tiles(region_spain, 
                      provider = "Esri.WorldShadedRelief", 
                      zoom = 7, crop = T)
# compute the distance between points
dist_proj <- as(dist(Matcoordalpha), "matrix")
# cartography
my_proj <- st_crs(region_spain)
spain <- st_union(region_spain)
```

We first plot the data:

```{r, fig.width = 10, height = 4}
#pdf(file = "figures/spain_data.pdf", width = 10, height = 4.5)
par(mfrow = c(1, 2), mar = c(3.7, 3, 1, 1), oma = c(0, 0, 0, 0),
    las = 1, mgp = c(2.15, 0.75, 0))
# map
plot_tiles(nc_osm)
mf_shadow(spain, add = T, cex = 0.8)
plot(st_geometry(spain), border = rgb(0.5, 0.5, 0.5), lwd = 0.4, add = T,
     col = rgb(0.82, 0.82, 0.82))
plot(st_geometry(region_spain), border = rgb(1, 1, 1), lwd = 0.4, add = T,
     col = rgb(0.82, 0.82, 0.82))
# data
plot(dates, MatX[, 1], ylim = y_lim, xlab = 'Quarters of the period 2002-2022',
       ylab = 'Unemployment rate', xaxt = "n", 
     col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
     type = "l")
# abline(v = seq(2002, 2022, by = 4), lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
abline(h = seq(0, 40, by = 10), lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))

axis(1, at = seq(2002, 2022, by = 1),
     labels = F)
text(x = seq(2002, 2022, by = 1), y = par()$usr[3] - 0.03 * (par()$usr[4] - par()$usr[3]),
     labels = paste0(seq(2002, 2022, by = 1), "QI"),
     srt = 45, adj = 1, xpd = T, cex = 0.8)
for(j in 2:47)
  lines(dates, MatX[,j], ylim = y_lim, lwd = 1.3, 
        col = rgb(0.4, 0.4, 0.4, alpha = 0.3)) 
#dev.off()
```

#### Descriptive Analysis

We represent the variable "Unemployment" aggregated over different periods of 2 years (i.e. eight quarters).

```{r, fig.width = 12, fig.height= 4.5}
nb_split <- 10
step_years <- split(1:80, 
           sort(rep_len(1:nb_split, length.out = length(dates))))
#pdf(paste0("figures/Spain_evol.pdf"), width = 10, height = 7)
par(mfrow = c(3, 4), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, colMeans(MatX[step_years[[j]], ]))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 7, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlOrRd"
    my_pal <- rev(alpha(colorspace::sequential_hcl(7, palette = nom_pal), 1))

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      my_mean <- colMeans(MatX[step_years[[j]], ])
      my_col <- alpha(my_pal[findInterval(my_mean, my_interval, all.inside = T)],
                  1)
    
      plot_tiles(nc_osm)
      mf_shadow(spain, add = T, cex = 0.8)
      plot(st_geometry(region_spain), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- dates[chosen_years_5]
      title(paste0(my_years[1], "-", round(my_years[length(my_years)])), line = -.75)
     plot(st_geometry(region_spain), border = rgb(0.9, 0.9, 0.9), 
          lwd = 0.00000001, add = T)
     plot(st_geometry(spain), border = rgb(0.5, 0.5, 0.5), lwd = 0.4, add = T)
     if(j == nb_split)
       maplegend::leg(type = "choro", val = my_interval, pos = "bottomright", 
                 pal = my_pal, val_rnd = 3, title = "Unemp")
    }
#dev.off()
```

Average across all the years :

```{r, fig.width = 12, fig.height= 4.5}
nb_split <- 1
step_years <- split(1:80, 
           sort(rep_len(1:nb_split, length.out = length(dates))))
#pdf(paste0("figures/Spain_mean.pdf"), width = 8, height = 7)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, colMeans(MatX[step_years[[j]], ]))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 7, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlOrRd"
    my_pal <- rev(alpha(colorspace::sequential_hcl(7, palette = nom_pal), 1))

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      my_mean <- colMeans(MatX[step_years[[j]], ])
      my_col <- alpha(my_pal[findInterval(my_mean, my_interval, all.inside = T)],
                  1)
    
      plot_tiles(nc_osm)
      mf_shadow(spain, add = T, cex = 0.8)
      plot(st_geometry(region_spain), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- dates[chosen_years_5]
      # title(paste0(my_years[1], "-", round(my_years[length(my_years)])), line = -.75)
     plot(st_geometry(region_spain), border = rgb(0.9, 0.9, 0.9), 
          lwd = 0.00000001, add = T)
     plot(st_geometry(spain), border = rgb(0.5, 0.5, 0.5), lwd = 0.4, add = T)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "bottomright", 
                 pal = my_pal, val_rnd = 3, title = "Unemp")
    }
#dev.off()
```

We compute all the potential clusters :

```{r}
my_pairs_sp <- find_all_cluster(Matcoordalpha)
```

We use the different methods to detect clusters.

### NPFSS method

**Most likely cluster**

```{r}
res_np <- compute_np(my_pairs_sp[[1]], my_pairs_sp[[2]], MatX)
res_np
```

**Significance**

```{r, eval = F}
p_value_np <- 0
B <- 999
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_np(my_pairs_sp[[1]], my_pairs_sp[[2]], MatXsim)
  p_value_np <- p_value_np + (res_np$stat < temp$stat)
}
cat("p-value: ", (1 + p_value_np) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

We drop from the list of potential cluster all the observations belonging to the most likely cluster. For instance, if the most likely cluster contains the observations 1, 3, 5, the potential cluster with observations 8, 1, 3, 5, 10,  is replaced by 8, 10. We then compute the NPFSS method on the new possible combinations

```{r}
cluster_g1_temp <- sapply(my_pairs_sp[[1]], function(x) setdiff(x, res_np$vec))
cluster_g2_temp <- sapply(my_pairs_sp[[2]], function(x) setdiff(x, res_np$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_np_2 <- compute_np(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], MatX)
res_np_2
```

**Significance**

```{r, eval = F}
p_value_np_2 <- 0
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_np(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], MatXsim)
  p_value_np_2 <- p_value_np_2 + (res_np_2$stat < temp$stat)
}
cat("p-value: ", (1 + p_value_np_2) / (1 + B))
```

```{r, echo = F}
cat("p-value:  0.023")
```


### PFSS method

**Most likely cluster**

```{r}
res_p <- compute_p(my_pairs_sp[[1]], my_pairs_sp[[2]], MatX)
res_p
```

**Significance**

```{r, eval = F}
p_value_p <- 0
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_np(my_pairs_sp[[1]], my_pairs_sp[[2]], MatXsim)
  p_value_p <- p_value_p + (res_p$stat < temp$stat)
}
cat("p-value: ", (1 + p_value_p) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```

**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(my_pairs_sp[[1]], function(x) setdiff(x, res_p$vec))
cluster_g2_temp <- sapply(my_pairs_sp[[2]], function(x) setdiff(x, res_p$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```

```{r}
res_p_2 <- compute_p(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], MatX)
res_p_2
```
**Significance of the secondary cluster 1**

```{r, eval = F}
p_value_p_2 <- 0
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_p(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], MatXsim)
  p_value_p_2 <- p_value_p_2 + (res_p_2$stat < temp$stat)
}
cat("p-value: ", (1 + p_value_p_2) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.015")
```



### DFFSS method

**Most likely cluster**

```{r}
res_dffss <- compute_dffss(my_pairs_sp[[1]], my_pairs_sp[[2]], MatX)
res_dffss
```

**Significance**

```{r, eval = F}
p_value_dffss <- 0
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_dffss(my_pairs_sp[[1]], my_pairs_sp[[2]], MatX)
  p_value_dffss <- p_value_dffss + (res_dffss$stat < temp$stat)
}
(1 + p_value_dffss) /  (1 + B)
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(my_pairs_sp[[1]], function(x) setdiff(x, res_dffss$vec))
cluster_g2_temp <- sapply(my_pairs_sp[[2]], function(x) setdiff(x, res_dffss$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```

```{r}
res_dffss_2 <- compute_dffss(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], MatX)
res_dffss_2
```

**Significance of the secondary cluster 1**

```{r, eval = F}
p_value_dffss_2 <- 0
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_dffss(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], MatXsim)
  p_value_dffss_2 <- p_value_dffss_2 + (res_dffss_2$stat < temp$stat)
}
cat("p-value: ", (1 + p_value_dffss_2) /  (1 + B))
```

```{r, echo = F}
cat("p-value:  0.004")
```



### HFSS method

**Most likely cluster**

We first determine the value of $d$:

```{r, fig.width = 6, fig.height = 3}
#pdf("figures/spain_h_CPV.pdf", width = 6, height = 4)
temp <- compute_h(my_pairs_sp[[1]], my_pairs_sp[[2]], MatX, 
                           d = nrow(MatX), plot_eigen = T)
#dev.off()
```

We choose $d=2$ (which corresponds to around $90\%$ of the variance explained; note that using $d=10$, which corresponds to $95\%$ of the variance leads to the same detected cluster)

```{r}
res_h <- compute_h(my_pairs_sp[[1]], my_pairs_sp[[2]], MatX, d = 2)
res_h
```

```{r, eval = F}
p_value_h <- 0
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_h(my_pairs_sp[[1]], my_pairs_sp[[2]], MatXsim,
                               d = 2)
  p_value_h <- p_value_h + (res_h$stat < temp$stat)
}
(1 + p_value_h) /  (1 + B)
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(my_pairs_sp[[1]], function(x) setdiff(x, res_h$vec))
cluster_g2_temp <- sapply(my_pairs_sp[[2]], function(x) setdiff(x, res_h$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


We look for an optimal value of $d$:

```{r, fig.width = 6, fig.height = 3}
#pdf("figures/spain_h_CPV_2.pdf", width = 6, height = 4)
temp <- compute_h(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], MatX, 
                           d = nrow(MatX), plot_eigen = T)
#dev.off()
```

We choose $d=2$.

```{r}
res_h_2 <- compute_h(cluster_g1_temp[-id_pos], 
                     cluster_g2_temp[-id_pos], MatX, d = 2)
res_h_2
```

**Significance of the secondary cluster 1**

```{r, eval = F}
p_value_h_2 <- 0
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_h(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], 
                    MatXsim, d = 2)
  p_value_h_2 <- p_value_h_2 + (res_h_2$stat < temp$stat)
}
cat("p-value: ", (1 + p_value_h_2) /  (1 + B))
```



```{r, echo = F}
cat("p-value:  0.003")
```


### Summary of the results

**Visualization of the result**

```{r}
res <- vector("list", 4)
res[[1]][[1]] <- res_h
res[[1]][[2]] <- res_h_2
res[[2]][[1]] <- res_dffss
res[[2]][[2]] <- res_dffss_2
res[[3]][[1]] <- res_np
res[[3]][[2]] <- res_np_2
res[[4]][[1]] <- res_p
res[[4]][[2]] <- res_p_2
names_method <- c("HFSS", "DFFSS", "NPFSS", "PFSS")
cols = c("#D35C79", "#009593")
```


```{r, fig.width = 12, fig.height = 6, message = F, warning = F}
my_var <- 'Unemployment rate (in %)'
my_country <- "ESP" 

y_lim <- range(MatX)

for(k in 1:4) {
  my_cluster_1 <- res[[k]][[1]]$vec
  my_cluster_2 <- res[[k]][[2]]$vec

#pdf(file = paste0("figures/", my_country, "_", names_method[k], ".pdf"), width = 11.5, height = 3.9) 
sf_use_s2(F)
nf <- layout( matrix(c(1,1,2,3), nrow=2, byrow=F) )
  par(mar = c(1.5, 0, 0, 0.2), 
      oma = c(0.5, 0, 2.4, 0), mgp = c(2.4, 0.6, 0), las = 1)
  ##### Map #########
    # map
  col_geo <- rep(rgb(0.9, 0.9, 0.9, alpha = 0.1), nrow(Matcoordalpha))
  cex_geo <- rep(0.7, nrow(Matcoordalpha))
  col_geo[my_cluster_1] <- alpha(cols[1], 0.8)
  cex_geo[my_cluster_1] <- 1

  col_geo[my_cluster_2] <- alpha(cols[2], 0.5)
  cex_geo[my_cluster_2] <- 1
  
  
   plot_tiles(nc_osm)
  mf_shadow(spain, add = T, cex = 0.8)
  mf_shadow(st_union(region_spain[my_cluster_1, ]), 
                add = T, cex = 0.8, col = cols[1])
  mf_shadow(st_union(region_spain[my_cluster_2, ]), 
         add = T, col= cols[2], cex = 0.8)
      
  plot(st_geometry(spain), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

  
  plot(st_geometry(region_spain), 
        border = "white",
        col = col_geo, 
        cex = cex_geo,
        pch = 16, asp = 1, add = T, lwd = 0.1)

  plot(st_geometry(st_union(region_spain[my_cluster_1, ])), 
         add = T, border= cols[1], col = NULL)
  plot(st_geometry(st_union(region_spain[my_cluster_2, ])), 
         add = T, border= cols[2], col = NULL)

      
    temp_1 <- draw.circle(Matcoordalpha[my_cluster_1[1], 1], 
                           Matcoordalpha[my_cluster_1[1], 2], 
                  as.numeric(dist_proj[my_cluster_1[1], 
                                       my_cluster_1[length(my_cluster_1)]]))

  temp_2 <- draw.circle(Matcoordalpha[my_cluster_2[1], 1], 
                           Matcoordalpha[my_cluster_2[1], 2], 
                  as.numeric(dist_proj[my_cluster_2[1], 
                                       my_cluster_2[length(my_cluster_2)]]))
    ###############
  polygon(temp_1$x, temp_1$y, border= cols[1],
             col = alpha(cols[1], 0.4), lty=1, lwd=1)

  polygon(temp_2$x, temp_2$y, border= cols[2],
             col = alpha(cols[2], 0.4), lty=1, lwd=1)
  
    mtext(my_var,
       side = 4, line = -2.8, las = 0)
    
  legend("topleft", legend = c("Most likely cluster", "Secondary cluster 1"),
         fill = c(cols[1], cols[2]), cex = 0.9, box.lty = 0)
      
  ##### Functional data
  plot(dates, MatX[, 1], ylim = y_lim, xlab = '',
       ylab = '', col = rgb(0.6, 0.6, 0.6, alpha = 0.5), xaxt = 'n', 
        type = "l")
  legend("topleft", legend = c("Most likely cluster"),
         lty = 1, col = c(cols[1]), cex = 0.9)
  abline(v = seq(2002, 2022, by = 4), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 40, by = 10),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))

  for (j in 2:ncol(MatX))
        lines(dates, MatX[, j], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
    
  for(i in my_cluster_1)
        lines(dates, MatX[, i], col = alpha(cols[1], 0.3),
          lty = 1, lwd = 1.3)
  
  lines(dates, rowMeans(MatX), lwd = 1.3, lty = 2)
  

plot(dates, MatX[, 1], ylim = y_lim, xlab = 'Years',
       ylab = '', xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  legend("topleft", legend = c("Secondary cluster 1"),
         lty = 1, col = c(cols[2]), cex = 0.9)
  abline(v = seq(2002, 2022, by = 4), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 40, by = 10),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(2002, 2022, by = 4), xlab = "years",
           labels = as.character(seq(2002, 2022, by = 4)))
  for (j in 2:ncol(MatX))
        lines(dates, MatX[, j], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
      
  for(i in my_cluster_2)
        lines(dates, MatX[, i], col = alpha(cols[2], 0.3),
          lty = 1, lwd = 1.3)
  
    
  lines(dates, rowMeans(MatX), lwd = 1.3, lty = 2)
  
  mtext(paste0("Clusters for the ", names_method[k]), side = 3, line = 0.8, outer = TRUE)
#dev.off()
}
```

We present in the following table the results obtained by the different methods. 

```{r}
res_spain <- data.frame(nb_cluster_1 = c(length(res_np$vec), 
                            length(res_p$vec),
                            length(res_dffss$vec),
                            length(res_h$vec)),
           sign_cluster_1 = c(0.001, 0.001, 0.001, 0.001),
           nb_cluster_2 = c(length(res_np_2$vec), 
                            length(res_p_2$vec),
                            length(res_dffss_2$vec),
                            length(res_h_2$vec)),
           sign_cluster_2 = c(0.023, 0.015, 0.004, 0.003))
row.names(res_spain) <- c("NPFSS", "PFSS", "DFFSS", "HFSS")
knitr::kable(res_spain)
```


## Climate Data

### Preparation of the data

Import the data:
```{r, message = F}
load("data/my_grid.RData")
coord_region <- st_as_sf(my_grid, coords = c("long", "lat"),
                         crs = 4326)
countries_regions <- st_read("data/world-administrative-boundaries.geojson")
```

Preparation of the data:
```{r}
# transforn the points into grid
long_lat_temp <- my_grid[, c("long", "lat")] %>%
  filter(lat > -90, lat < 90, long > -180, long < 180) %>%
  arrange(lat, long)   %>%
  select(long, lat)
df_sf_temp <- st_as_sf(long_lat_temp, coords = c("long", "lat"))
st_crs(df_sf_temp) <- 4326
all_cells <- df_sf_temp %>% 
  st_make_grid(cellsize = c(0.625, 0.5), 
               offset = c(-179.375 - 0.3125, -89.5 -0.25)) %>% # c(-180 - 0.3125, -90 - 0.25))
  st_as_sf() %>% 
  st_join(df_sf_temp) 
all_cells$long <- long_lat_temp$long
all_cells$lat <- long_lat_temp$lat
# Initialisation
unique_year <- 1981:2023
chosen_years <- 20:43
```

### Great-Britain

We select the ISO3 corresponding to Great-Britain. 

```{r}
my_country <- "GBR"
my_proj <- 3035 
```

```{r, warning = F, message = F}
select_countries <- countries_regions[countries_regions$color_code %in% my_country, ]
  
# drop the islands
sf_obj <- select_countries %>%
    filter(iso3 == my_country[1]) %>%
    mutate(area = st_area(geometry)) %>%
    top_n(1, area) %>%
    rowid_to_column() %>%
    st_cast("POLYGON")  %>% 
    mutate(area = st_area(geometry)) %>%
    group_by(rowid) %>%
    top_n(1, area)
  
if(length(my_country) > 1) {
  for(i in 2:length(my_country)) {
    sf_obj <- rbind(sf_obj, select_countries %>%
      filter(iso3 == my_country[i]) %>%
      mutate(area = st_area(geometry)) %>%
      top_n(1, area) %>%
      rowid_to_column() %>%
      st_cast("POLYGON")  %>% 
      mutate(area = st_area(geometry)) %>%
      group_by(rowid) %>%
      top_n(1, area)
    )
  }
  }
is_intersect <- st_intersects(all_cells, sf_obj)
is_intersect <- which(sapply(is_intersect, function(x) length(x) != 0))
lldata_poly <- all_cells[is_intersect, ]
  
my_contours <- st_intersection(countries_regions, 
                               st_union(lldata_poly, is_coverage = T))

# dowload tiles and compose raster (SpatRaster)
nc_osm <- get_tiles(my_contours, 
                      provider = "Esri.WorldShadedRelief", 
                      zoom = 7, crop = F)
poly_cell <- merge(lldata_poly, my_grid, by = c("long", "lat"))
# coordinates of the centroid
coord_temp <- st_transform(poly_cell, my_proj)
# simplify the geometry
poly_cell <- st_intersection(poly_cell, my_contours)
poly_cell <- poly_cell[!duplicated(cbind(poly_cell$long, poly_cell$lat)), ]
```

We compute all the potential clusters:

```{r, warning = F}
  coord_proj <- st_coordinates(st_centroid(coord_temp))
  pairs_geo <- find_all_cluster(coord_proj)
  dist_proj <- as(dist(cbind(coord_proj[, 1], coord_proj[, 2])), "matrix")

  dist_4326 <- as(dist(cbind(poly_cell$long, poly_cell$lat)), "matrix")
  coord_4326 <- st_transform(poly_cell, 4326) 
```

```{r}
  my_var <- "t2m_diff_" 
  temp_var <- poly_cell[, paste0(my_var, unique_year)[chosen_years]]
  MatX <- as.matrix(st_drop_geometry(temp_var))
```

#### Descriptive Analysis

We first plot the data:

```{r, fig.width = 10, height = 4}
y_lim <- range(MatX)
dates <- unique_year[chosen_years]
#pdf(file = "figures/GBR_data.pdf", width = 10, height = 4.5)
par(mfrow = c(1, 2), mar = c(3.7, 3, 1, 1), oma = c(0, 0, 0, 0),
    las = 1, mgp = c(2.2, 0.8, 0))
# map
plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

# data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = 'Difference from average temperatures (in C)', xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(-4, 4, by = 0.5),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
#dev.off()
```

We map the variable "Difference from average temperatures" aggregated over different periods of four years. 

```{r, fig.width = 12, fig.height= 3.5}
nb_split <- 8
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/GB_evol.pdf"), width = 12, height = 8)
par(mfrow = c(2, 4), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 8, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlOrRd"
    my_pal <- rev(alpha(colorspace::sequential_hcl(8, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
      title(paste0(my_years[1], "-", my_years[length(my_years)]), line = -1.25)
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Diff Temp")
    }
#dev.off()
```

Average mean of difference temperatures:
```{r, fig.width = 12, fig.height= 3.5}
nb_split <- 1
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/GBR_mean.pdf"), width = 7, height = 8)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 8, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlOrRd"
    my_pal <- rev(alpha(colorspace::sequential_hcl(8, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
     # title(paste0(my_years[1], "-", my_years[length(my_years)]), line = -1.25)
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Diff Temp")
    }
#dev.off()
```

#### NPFSS method

**Most likely cluster**

```{r}
res_np <- compute_np(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_np
```

**Significance (using parallel computing)**

```{r, eval = F}
B <- 999
compute_fun_par <- function(b, fun, ...) {
  set.seed(b)
  perm <- sample(nrow(MatX))
  MatXsim <- MatX[perm, ]
  temp <- fun(c1, c2, t(MatXsim))
  temp$stat
}

c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

require(parallel)
cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_np$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_np$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```

```{r}
res_np_2 <- compute_np(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_np_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np_2$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```



#### PFSS method

**Most likely cluster**

```{r}
res_p <- compute_p(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_p
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p$stat <x))) /  (1 + B))
```



```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_p$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_p$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_p_2 <- compute_p(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_p_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p_2$stat <x))) /  (1 + B))
```

```{r, echo = F}
cat("p-value:  0.001")
```



#### DFFSS method

**Most likely cluster**

```{r}
res_dffss <- compute_dffss(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_dffss
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_p$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_dffss$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_dffss$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_dffss_2 <- compute_dffss(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_dffss_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_dffss_2$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```



#### HFSS method

**Most likely cluster**

We first determine the value of $d$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV.pdf"), width = 6, height = 4)
temp <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), 
                           d = ncol(MatX), plot_eigen = T)
#dev.off()
```

We choose $d=6$:

```{r}
res_h <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), d = 6)
res_h
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, d = 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_h$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_h$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```

We look for an optimal value of $d$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV_2.pdf"), width = 6, height = 4)
temp <- compute_h(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], t(MatX), 
                           d = ncol(MatX), plot_eigen = T)
#dev.off()
```

We choose $d=6$.

```{r}
res_h_2 <- compute_h(cluster_g1_temp[-id_pos], 
                     cluster_g2_temp[-id_pos], t(MatX), d = 6)
res_h_2
```


**Significance of the secondary cluster 1**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, d = 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h_2$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```



#### Summary of the results

**Visualization of the result**

```{r}
res <- vector("list", 4)
res[[1]][[1]] <- res_h
res[[1]][[2]] <- res_h_2
res[[2]][[1]] <- res_dffss
res[[2]][[2]] <- res_dffss_2
res[[3]][[1]] <- res_np
res[[3]][[2]] <- res_np_2
res[[4]][[1]] <- res_p
res[[4]][[2]] <- res_p_2
```

```{r, fig.width = 12, fig.height = 6, message = F, warning = F}
my_var <- 'Difference with the normal temperature (in C)'
dates <- unique_year[chosen_years]
y_lim <- range(MatX)

for(k in 1:4) {
  my_cluster_1 <- res[[k]][[1]]$vec
  my_cluster_2 <- res[[k]][[2]]$vec

#pdf(file = paste0("figures/", my_country, "_", names_method[k], ".pdf"), width = 11.5, height = 3.8) 
sf_use_s2(F)
nf <- layout( matrix(c(1,1,2,3), nrow=2, byrow=F) )
  par(mar = c(1.5, 0, 0, 0.2), 
      oma = c(0.5, 0, 2.4, 0), mgp = c(2.4, 0.6, 0), las = 1)
  ##### Map #########
    # map
  col_geo <- rep(rgb(0.9, 0.9, 0.9, alpha = 0.1), nrow(poly_cell))
  cex_geo <- rep(0.7, nrow(poly_cell))
  col_geo[my_cluster_1] <- alpha(cols[1], 0.8)
  cex_geo[my_cluster_1] <- 1

  col_geo[my_cluster_2] <- alpha(cols[2], 0.5)
  cex_geo[my_cluster_2] <- 1
  
  plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  mf_shadow(st_union(poly_cell[my_cluster_1, ]), 
                add = T, cex = 0.8, col = cols[1])
  mf_shadow(st_union(poly_cell[my_cluster_2, ]), 
         add = T, col= cols[2], cex = 0.8)
      
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

  
  plot(st_geometry(poly_cell), 
        border = "white",
        col = col_geo, 
        cex = cex_geo,
        pch = 16, asp = 1, add = T, lwd = 0.1)

  plot(st_geometry(st_union(poly_cell[my_cluster_1, ])), 
         add = T, border= cols[1], col = NULL)
  plot(st_geometry(st_union(poly_cell[my_cluster_2, ])), 
         add = T, border= cols[2], col = NULL)
      
  temp_1 <- draw.circle(coord_proj[my_cluster_1[1], 1], 
                           coord_proj[my_cluster_1[1], 2], 
                  as.numeric(dist_proj[my_cluster_1[1], 
                                       my_cluster_1[length(my_cluster_1)]]))

    my_circle_1 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_1$x, temp_1$x[1]), 
             c(temp_1$y, temp_1$y[1]))
         )), crs = my_proj
       ), 4326)
    
  temp_2 <- draw.circle(coord_proj[my_cluster_2[1], 1], 
                           coord_proj[my_cluster_2[1], 2], 
                  as.numeric(dist_proj[my_cluster_2[1], 
                                       my_cluster_2[length(my_cluster_2)]]))
    
  my_circle_2 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_2$x, temp_2$x[1]), 
             c(temp_2$y, temp_2$y[1]))
         )), crs = my_proj
       ), 4326)
  
  
  ###############
  plot(st_geometry(my_circle_2), add = T, border= cols[2],
             col = alpha(cols[2], 0.4), lty=1, lwd=1)
  plot(st_geometry(my_circle_1), add = T, border= cols[1],
             col = alpha(cols[1], 0.4), lty=1, lwd=1)
  
    mtext(my_var,
       side = 4, line = -3.5, las = 0)
    
  legend("topleft", legend = c("Most likely cluster", "Secondary cluster 1"),
         fill = c(cols[1], cols[2]), cex = 0.9, box.lty = 0)
      
  ##### Functional data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = '',
       ylab = '', col = rgb(0.6, 0.6, 0.6, alpha = 0.5), xaxt = 'n', 
        type = "l")
  legend("topleft", legend = c("Most likely cluster"),
         lty = 1, col = c(cols[1]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(-4, 4, by = 0.5),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))

  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
    
  for(i in my_cluster_1)
        lines(dates, MatX[i, ], col = alpha(cols[1], 0.3),
          lty = 1, lwd = 1.3)
  
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  

plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = 'Difference from average temperatures (in C)', xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  legend("topleft", legend = c("Secondary cluster 1"),
         lty = 1, col = c(cols[2]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(-4, 4, by = 0.5),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
      
  for(i in my_cluster_2)
        lines(dates, MatX[i, ], col = alpha(cols[2], 0.3),
          lty = 1, lwd = 1.3)
  
    
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  
  mtext(paste0("Clusters for the ", names_method[k]), side = 3, line = 0.8, outer = TRUE)
#dev.off()
}
```

```{r}
res_GB <- data.frame(nb_cluster_1 = c(length(res_np$vec), 
                            length(res_p$vec),
                            length(res_dffss$vec),
                            length(res_h$vec)),
           sign_cluster_1 = c(0.001, 0.001, 0.001, 0.001),
           nb_cluster_2 = c(length(res_np_2$vec), 
                            length(res_p_2$vec),
                            length(res_dffss_2$vec),
                            length(res_h_2$vec)),
           sign_cluster_2 = c(0.001, 0.001, 0.001, 0.001))
row.names(res_GB) <- c("NPFSS", "PFSS", "DFFSS", "HFSS")
knitr::kable(res_GB)
```



### Nigeria

We select the ISO3 of Nigeria:

```{r}
my_country <- "NGA"
my_proj <- 32629 
```

```{r, warning = F, message = F}
select_countries <- countries_regions[countries_regions$color_code %in% my_country, ]
  
# drop the islands
sf_obj <- select_countries %>%
    filter(iso3 == my_country[1]) %>%
    mutate(area = st_area(geometry)) %>%
    top_n(1, area) %>%
    rowid_to_column() %>%
    st_cast("POLYGON")  %>% 
    mutate(area = st_area(geometry)) %>%
    group_by(rowid) %>%
    top_n(1, area)
  
if(length(my_country) > 1) {
  for(i in 2:length(my_country)) {
    sf_obj <- rbind(sf_obj, select_countries %>%
      filter(iso3 == my_country[i]) %>%
      mutate(area = st_area(geometry)) %>%
      top_n(1, area) %>%
      rowid_to_column() %>%
      st_cast("POLYGON")  %>% 
      mutate(area = st_area(geometry)) %>%
      group_by(rowid) %>%
      top_n(1, area)
    )
  }
  }
is_intersect <- st_intersects(all_cells, sf_obj)
is_intersect <- which(sapply(is_intersect, function(x) length(x) != 0))
lldata_poly <- all_cells[is_intersect, ]
  
my_contours <- st_intersection(select_countries, 
                               st_union(lldata_poly, is_coverage = T))

# dowload tiles and compose raster (SpatRaster)
nc_osm <- get_tiles(my_contours, 
                      provider = "Esri.WorldShadedRelief", 
                      zoom = 7, crop = F)
poly_cell <- merge(lldata_poly, my_grid, by = c("long", "lat"))
# coordinates of the centroid
coord_temp <- st_transform(poly_cell, my_proj)
# simplify the geometry
poly_cell <- st_intersection(poly_cell, my_contours)
```

We compute all the potential clusters:

```{r, warning = F}
  coord_proj <- st_coordinates(st_centroid(coord_temp))
  pairs_geo <- find_all_cluster(coord_proj)
  dist_proj <- as(dist(cbind(coord_proj[, 1], coord_proj[, 2])), "matrix")

  dist_4326 <- as(dist(cbind(poly_cell$long, poly_cell$lat)), "matrix")
  coord_4326 <- st_transform(poly_cell, 4326) 
```

```{r}
  my_var <- "prec_5days_" 
  temp_var <- poly_cell[, paste0(my_var, unique_year)[chosen_years]]
  MatX <- as.matrix(st_drop_geometry(temp_var))
```

#### Descriptive Analysis

We first plot the data:

```{r, fig.width = 10, height = 4}
y_lim <- range(MatX)
title_var <- 'Maximum consecutive 5-days precipitation (in mm)'
#pdf(file = "figures/NGA_data.pdf", width = 12, height = 4.5)
par(mfrow = c(1, 2), mar = c(3.7, 3, 1, 1), oma = c(0, 0, 0, 0),
    las = 1, mgp = c(2.2, 0.5, 0))
# map
plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

# data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = title_var, xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
abline(h = seq(0, 2500, by = 500),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
#dev.off()
```

We map the average of the variable over a 3-year window. 

```{r, fig.width = 12, fig.height= 6}
nb_split <- 8
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/NGA_evol.pdf"), width = 12, height = 5)
par(mfrow = c(2, 4), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 8, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlGnBu"
    my_pal <- rev(alpha(colorspace::sequential_hcl(8, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
      title(paste0(my_years[1], "-", my_years[length(my_years)]), line = -1.25)
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Prec 5-days")
    }
#dev.off()
```

Average mean:
```{r, fig.width = 12, fig.height= 6}
nb_split <- 1
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/NGA_mean.pdf"), width = 7, height = 5)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 8, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlGnBu"
    my_pal <- rev(alpha(colorspace::sequential_hcl(8, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
      #title(paste0(my_years[1], "-", my_years[length(my_years)]), line = -1.25)
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Prec 5-days")
    }
#dev.off()
```

#### NPFSS method

**Most likely cluster**

```{r}
res_np <- compute_np(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_np
```

**Significance (using parallel computing)**

```{r, eval = F}
B <- 999

c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_np$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_np$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_np_2 <- compute_np(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_np_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np_2$stat < x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```




#### PFSS method

**Most likely cluster**

```{r}
res_p <- compute_p(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_p
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p$stat <x))) /  (1 + B))
```



```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_p$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_p$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_p_2 <- compute_p(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_p_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p_2$stat <x))) /  (1 + B))
```

```{r, echo = F}
cat("p-value:  0.001")
```



#### DFFSS method

**Most likely cluster**

```{r}
res_dffss <- compute_dffss(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_dffss
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_p$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_dffss$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_dffss$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```



```{r}
res_dffss_2 <- compute_dffss(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_dffss_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_dffss_2$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```





#### HFSS Method 

**Most likely cluster**

We first determine the value of $d$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV.pdf"), width = 6, height = 4)
temp <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), 
                           d = ncol(MatX), plot_eigen = T)
#dev.off()
```

We choose $d=6$:
```{r}
res_h <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), d = 6)
res_h
```

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, d = 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_h$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_h$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


We look for an optimal value of $d$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV_2.pdf"), width = 6, height = 4)
temp <- compute_h(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], t(MatX), 
                           d = ncol(MatX), plot_eigen = T)
#dev.off()
```

We choose $d=6$.

```{r}
res_h_2 <- compute_h(cluster_g1_temp[-id_pos], 
                     cluster_g2_temp[-id_pos], t(MatX), d = 6)
res_h_2
```


**Significance of the secondary cluster 1**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, d = 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h_2$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```



#### Summary of the results

**Visualization of the result**

```{r}
res <- vector("list", 4)
res[[1]][[1]] <- res_h
res[[1]][[2]] <- res_h_2
res[[2]][[1]] <- res_dffss
res[[2]][[2]] <- res_dffss_2
res[[3]][[1]] <- res_np
res[[3]][[2]] <- res_np_2
res[[4]][[1]] <- res_p
res[[4]][[2]] <- res_p_2
```

```{r, fig.width = 12, fig.height = 6, message = F, warning = F}
title_var <- 'Maximum consecutive 5-days precipitation (in mm)'
dates <- unique_year[chosen_years]
y_lim <- range(MatX)

for(k in 1:4) {
  my_cluster_1 <- res[[k]][[1]]$vec
  my_cluster_2 <- res[[k]][[2]]$vec

#pdf(file = paste0("figures/", my_country, "_", names_method[k], ".pdf"), width = 13, height = 4.) 
sf_use_s2(F)
nf <- layout( matrix(c(1,1,2,3), nrow=2, byrow=F) )
  par(mar = c(1.5, 0, 0, 0.2), 
      oma = c(0.5, 0, 2.4, 0), mgp = c(2.4, 0.6, 0), las = 1)
  ##### Map #########
    # map
  col_geo <- rep(rgb(0.9, 0.9, 0.9, alpha = 0.1), nrow(poly_cell))
  cex_geo <- rep(0.7, nrow(poly_cell))
  col_geo[my_cluster_1] <- alpha(cols[1], 0.8)
  cex_geo[my_cluster_1] <- 1

  col_geo[my_cluster_2] <- alpha(cols[2], 0.5)
  cex_geo[my_cluster_2] <- 1
  
  plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  mf_shadow(st_union(poly_cell[my_cluster_1, ]), 
                add = T, cex = 0.8, col = cols[1])
  mf_shadow(st_union(poly_cell[my_cluster_2, ]), 
         add = T, col= cols[2], cex = 0.8)
      
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

  
  plot(st_geometry(poly_cell), 
        border = "white",
        col = col_geo, 
        cex = cex_geo,
        pch = 16, asp = 1, add = T, lwd = 0.1)

  plot(st_geometry(st_union(poly_cell[my_cluster_1, ])), 
         add = T, border= cols[1], col = NULL)
  plot(st_geometry(st_union(poly_cell[my_cluster_2, ])), 
         add = T, border= cols[2], col = NULL)
      
  temp_1 <- draw.circle(coord_proj[my_cluster_1[1], 1], 
                           coord_proj[my_cluster_1[1], 2], 
                  as.numeric(dist_proj[my_cluster_1[1], 
                                       my_cluster_1[length(my_cluster_1)]]))

    my_circle_1 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_1$x, temp_1$x[1]), 
             c(temp_1$y, temp_1$y[1]))
         )), crs = my_proj
       ), 4326)
    
  temp_2 <- draw.circle(coord_proj[my_cluster_2[1], 1], 
                           coord_proj[my_cluster_2[1], 2], 
                  as.numeric(dist_proj[my_cluster_2[1], 
                                       my_cluster_2[length(my_cluster_2)]]))
    
  my_circle_2 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_2$x, temp_2$x[1]), 
             c(temp_2$y, temp_2$y[1]))
         )), crs = my_proj
       ), 4326)
  
  
  ###############
  plot(st_geometry(my_circle_2), add = T, border= cols[2],
             col = alpha(cols[2], 0.4), lty=1, lwd=1)
  plot(st_geometry(my_circle_1), add = T, border= cols[1],
             col = alpha(cols[1], 0.4), lty=1, lwd=1)
  
    mtext(my_var, side = 4, line = -3.5, las = 0, cex = 0.8)
    
  legend("topleft", legend = c("Most likely cluster", "Secondary cluster 1"),
         fill = c(cols[1], cols[2]), cex = 0.9, box.lty = 0)
      
  ##### Functional data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = '',
       ylab = '', col = rgb(0.6, 0.6, 0.6, alpha = 0.5), xaxt = 'n', 
        type = "l")
  legend("topleft", legend = c("Most likely cluster"),
         lty = 1, col = c(cols[1]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 2500, by = 500),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))

  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
    
  for(i in my_cluster_1)
        lines(dates, MatX[i, ], col = alpha(cols[1], 0.3),
          lty = 1, lwd = 1.3)
  
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  

plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = title_var, xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  legend("topleft", legend = c("Secondary cluster 1"),
         lty = 1, col = c(cols[2]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 2500, by = 500),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
      
  for(i in my_cluster_2)
        lines(dates, MatX[i, ], col = alpha(cols[2], 0.3),
          lty = 1, lwd = 1.3)
  
    
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  
  mtext(paste0("Clusters for the ", names_method[k]), side = 3, line = 0.8, outer = TRUE)
#dev.off()
}
```


We present in the following table the results obtained by the different methods. 

```{r}
res_NGA <- data.frame(nb_cluster_1 = c(length(res_np$vec), 
                            length(res_p$vec),
                            length(res_dffss$vec),
                            length(res_h$vec)),
           sign_cluster_1 = c(0.001, 0.001, 0.001, 0.001),
           nb_cluster_2 = c(length(res_np_2$vec), 
                            length(res_p_2$vec),
                            length(res_dffss_2$vec),
                            length(res_h_2$vec)),
           sign_cluster_2 = c(0.001, 0.001, 0.001, 0.001))
row.names(res_NGA) <- c("NPFSS", "PFSS", "DFFSS", "HFSS")
knitr::kable(res_NGA)
```



### Pakistan

We select the ISO3 for Pakistan:

```{r}
my_country <- "PAK"
my_proj <- 9678 
```

```{r, warning = F, message = F}
select_countries <- countries_regions[countries_regions$color_code %in% my_country, ]
  
# drop the islands
sf_obj <- select_countries %>%
    filter(iso3 == my_country[1]) %>%
    mutate(area = st_area(geometry)) %>%
    top_n(1, area) %>%
    rowid_to_column() %>%
    st_cast("POLYGON")  %>% 
    mutate(area = st_area(geometry)) %>%
    group_by(rowid) %>%
    top_n(1, area)
  
if(length(my_country) > 1) {
  for(i in 2:length(my_country)) {
    sf_obj <- rbind(sf_obj, select_countries %>%
      filter(iso3 == my_country[i]) %>%
      mutate(area = st_area(geometry)) %>%
      top_n(1, area) %>%
      rowid_to_column() %>%
      st_cast("POLYGON")  %>% 
      mutate(area = st_area(geometry)) %>%
      group_by(rowid) %>%
      top_n(1, area)
    )
  }
  }
is_intersect <- st_intersects(all_cells, sf_obj)
is_intersect <- which(sapply(is_intersect, function(x) length(x) != 0))
lldata_poly <- all_cells[is_intersect, ]
  
my_contours <- st_intersection(select_countries, 
                               st_union(lldata_poly, is_coverage = T))

# dowload tiles and compose raster (SpatRaster)
nc_osm <- get_tiles(my_contours, 
                      provider = "Esri.WorldShadedRelief", 
                      zoom = 7, crop = F)
poly_cell <- merge(lldata_poly, my_grid, by = c("long", "lat"))
# coordinates of the centroid
coord_temp <- st_transform(poly_cell, my_proj)
# simplify the geometry
poly_cell <- st_intersection(poly_cell, my_contours)
```

We compute all the potential clusters:

```{r, warning = F}
  coord_proj <- st_coordinates(st_centroid(coord_temp))
  pairs_geo <- find_all_cluster(coord_proj)
  dist_proj <- as(dist(cbind(coord_proj[, 1], coord_proj[, 2])), "matrix")

  dist_4326 <- as(dist(cbind(poly_cell$long, poly_cell$lat)), "matrix")
  coord_4326 <- st_transform(poly_cell, 4326) 
```

```{r}
  my_var <- "prec_5days_" 
  temp_var <- poly_cell[, paste0(my_var, unique_year)[chosen_years]]
  MatX <- as.matrix(st_drop_geometry(temp_var))
```

#### Descriptive Analysis

We first plot the data:

```{r, fig.width = 10, height = 4}
y_lim <- range(MatX)
title_var <- 'Maximum consecutive 5-days precipitation (in mm)'
#pdf(file = "figures/PAK_data.pdf", width = 10, height = 4.5)
par(mfrow = c(1, 2), mar = c(3.7, 3, 1, 1), oma = c(0, 0, 0, 0),
    las = 1, mgp = c(2.2, 0.8, 0))
# map
plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

# data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = title_var, xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
abline(h = seq(0, 800, by = 200),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
#dev.off()
```

We map the average of the variable over a 3-year window Before the last period, the Southeaster and the North were a little bit more impacted by heavy precipitations. However, during the last period, heavy rainfall affected the whole country.

```{r, fig.width = 12, fig.height= 6}
nb_split <- 8
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/PAK_evol.pdf"), width = 12, height = 8)
par(mfrow = c(2, 4), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 8, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlGnBu"
    my_pal <- rev(alpha(colorspace::sequential_hcl(8, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
      title(paste0(my_years[1], "-", my_years[length(my_years)]), line = -1.25)
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Prec 5-days")
    }
#dev.off()
```

Average mean:
```{r, fig.width = 12, fig.height= 6}
nb_split <- 1
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/PAK_mean.pdf"), width = 8, height = 8)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 8, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlGnBu"
    my_pal <- rev(alpha(colorspace::sequential_hcl(8, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
      #title(paste0(my_years[1], "-", my_years[length(my_years)]), line = -1.25)
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Prec 5-days")
    }
#dev.off()
```
#### NPFSS method

**Most likely cluster**

```{r}
res_np <- compute_np(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_np
```

**Significance (using parallel computing)**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

require(parallel)
cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np$stat < x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_np$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_np$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```

```{r}
res_np_2 <- compute_np(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_np_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np_2$stat < x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```



#### PFSS method

**Most likely cluster**

```{r}
res_p <- compute_p(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_p
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p$stat <x))) /  (1 + B))
```



```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_p$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_p$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_p_2 <- compute_p(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_p_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p_2$stat <x))) /  (1 + B))
```

```{r, echo = F}
cat("p-value:  0.001")
```




#### DFFSS method

**Most likely cluster**

```{r}
res_dffss <- compute_dffss(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_dffss
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_p$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_dffss$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_dffss$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```



```{r}
res_dffss_2 <- compute_dffss(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_dffss_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_dffss_2$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```




### HFSS method

**Most likely cluster**

We first determine the value of $d$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV.pdf"), width = 6, height = 4)
temp <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), 
                           d = ncol(MatX), plot_eigen = T)
#dev.off()
```

We choose $d=6$:
```{r}
res_h <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), d = 6)
res_h
```

To compute the significance, we make $B$ permutations on the data and compute the number of times the scan statistic is lower than the observed one: 

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, d = 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_h$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_h$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


We look for an optimal value of $d$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV_2.pdf"), width = 6, height = 4)
temp <- compute_h(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], t(MatX), 
                           d = ncol(MatX), plot_eigen = T)
#dev.off()
```

We choose $d=6$.

```{r}
res_h_2 <- compute_h(cluster_g1_temp[-id_pos], 
                     cluster_g2_temp[-id_pos], t(MatX), d = 6)
res_h_2
```


**Significance of the secondary cluster 1**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, d = 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h_2$stat <x))) /  (1 + B))
```



```{r, echo = F}
cat("p-value:  0.001")
```


### Summary of the results

**Visualization of the result**

```{r}
res <- vector("list", 4)
res[[1]][[1]] <- res_h
res[[1]][[2]] <- res_h_2
res[[2]][[1]] <- res_dffss
res[[2]][[2]] <- res_dffss_2
res[[3]][[1]] <- res_np
res[[3]][[2]] <- res_np_2
res[[4]][[1]] <- res_p
res[[4]][[2]] <- res_p_2
```

```{r, fig.width = 12, fig.height = 6, message = F, warning = F}
title_var <- 'Maximum consecutive 5-days precipitation (in mm)'
dates <- unique_year[chosen_years]
y_lim <- range(MatX)

for(k in 1:4) {
  my_cluster_1 <- res[[k]][[1]]$vec
  my_cluster_2 <- res[[k]][[2]]$vec

#pdf(file = paste0("figures/", my_country, "_", names_method[k], ".pdf"), width = 13, height = 4.2) 
sf_use_s2(F)
nf <- layout( matrix(c(1,1,2,3), nrow=2, byrow=F) )
  par(mar = c(1.5, 0, 0, 0.2), 
      oma = c(0.5, 0, 2.4, 0), mgp = c(2.4, 0.6, 0), las = 1)
  ##### Map #########
    # map
  col_geo <- rep(rgb(0.9, 0.9, 0.9, alpha = 0.1), nrow(poly_cell))
  cex_geo <- rep(0.7, nrow(poly_cell))
  col_geo[my_cluster_1] <- alpha(cols[1], 0.8)
  cex_geo[my_cluster_1] <- 1

  col_geo[my_cluster_2] <- alpha(cols[2], 0.5)
  cex_geo[my_cluster_2] <- 1
  
  plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  mf_shadow(st_union(poly_cell[my_cluster_1, ]), 
                add = T, cex = 0.8, col = cols[1])
  mf_shadow(st_union(poly_cell[my_cluster_2, ]), 
         add = T, col= cols[2], cex = 0.8)
      
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

  
  plot(st_geometry(poly_cell), 
        border = "white",
        col = col_geo, 
        cex = cex_geo,
        pch = 16, asp = 1, add = T, lwd = 0.1)

  plot(st_geometry(st_union(poly_cell[my_cluster_1, ])), 
         add = T, border= cols[1], col = NULL)
  plot(st_geometry(st_union(poly_cell[my_cluster_2, ])), 
         add = T, border= cols[2], col = NULL)
      
  temp_1 <- draw.circle(coord_proj[my_cluster_1[1], 1], 
                           coord_proj[my_cluster_1[1], 2], 
                  as.numeric(dist_proj[my_cluster_1[1], 
                                       my_cluster_1[length(my_cluster_1)]]))

    my_circle_1 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_1$x, temp_1$x[1]), 
             c(temp_1$y, temp_1$y[1]))
         )), crs = my_proj
       ), 4326)
    
  temp_2 <- draw.circle(coord_proj[my_cluster_2[1], 1], 
                           coord_proj[my_cluster_2[1], 2], 
                  as.numeric(dist_proj[my_cluster_2[1], 
                                       my_cluster_2[length(my_cluster_2)]]))
    
  my_circle_2 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_2$x, temp_2$x[1]), 
             c(temp_2$y, temp_2$y[1]))
         )), crs = my_proj
       ), 4326)
  
  
  ###############
  plot(st_geometry(my_circle_2), add = T, border= cols[2],
             col = alpha(cols[2], 0.4), lty=1, lwd=1)
  plot(st_geometry(my_circle_1), add = T, border= cols[1],
             col = alpha(cols[1], 0.4), lty=1, lwd=1)
  
    mtext(my_var, side = 4, line = -3.5, las = 0, cex = 0.8)
    
  legend("topleft", legend = c("Most likely cluster", "Secondary cluster 1"),
         fill = c(cols[1], cols[2]), cex = 0.9, box.lty = 0)
      
  ##### Functional data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = '',
       ylab = '', col = rgb(0.6, 0.6, 0.6, alpha = 0.5), xaxt = 'n', 
        type = "l")
  legend("topleft", legend = c("Most likely cluster"),
         lty = 1, col = c(cols[1]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 2500, by = 200),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))

  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
    
  for(i in my_cluster_1)
        lines(dates, MatX[i, ], col = alpha(cols[1], 0.3),
          lty = 1, lwd = 1.3)
  
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  

plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = title_var, xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  legend("topleft", legend = c("Secondary cluster 1"),
         lty = 1, col = c(cols[2]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 2500, by = 200),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
      
  for(i in my_cluster_2)
        lines(dates, MatX[i, ], col = alpha(cols[2], 0.3),
          lty = 1, lwd = 1.3)
  
    
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  
  mtext(paste0("Clusters for the ", names_method[k]), side = 3, line = 0.8, outer = TRUE)
#dev.off()
}
```

We present in the following table the results obtained by the different methods. 

```{r}
res_PAK <- data.frame(nb_cluster_1 = c(length(res_np$vec), 
                            length(res_p$vec),
                            length(res_dffss$vec),
                            length(res_h$vec)),
           sign_cluster_1 = c(0.001, 0.001, 0.001, 0.001),
           nb_cluster_2 = c(length(res_np_2$vec), 
                            length(res_p_2$vec),
                            length(res_dffss_2$vec),
                            length(res_h_2$vec)),
           sign_cluster_2 = c(0.001, 0.001, 0.001, 0.001))
row.names(res_PAK) <- c("NPFSS", "PFSS", "DFFSS", "HFSS")
knitr::kable(res_PAK)
```




### Venezuela

We select the ISO3 for Venezuela:

```{r}
my_country <- "VEN"
my_proj <- 4247 
```

```{r, warning = F, message = F}
select_countries <- countries_regions[countries_regions$color_code %in% my_country, ]
  
# drop the islands
sf_obj <- select_countries %>%
    filter(iso3 == my_country[1]) %>%
    mutate(area = st_area(geometry)) %>%
    top_n(1, area) %>%
    rowid_to_column() %>%
    st_cast("POLYGON")  %>% 
    mutate(area = st_area(geometry)) %>%
    group_by(rowid) %>%
    top_n(1, area)
  
if(length(my_country) > 1) {
  for(i in 2:length(my_country)) {
    sf_obj <- rbind(sf_obj, select_countries %>%
      filter(iso3 == my_country[i]) %>%
      mutate(area = st_area(geometry)) %>%
      top_n(1, area) %>%
      rowid_to_column() %>%
      st_cast("POLYGON")  %>% 
      mutate(area = st_area(geometry)) %>%
      group_by(rowid) %>%
      top_n(1, area)
    )
  }
  }
is_intersect <- st_intersects(all_cells, sf_obj)
is_intersect <- which(sapply(is_intersect, function(x) length(x) != 0))
lldata_poly <- all_cells[is_intersect, ]
  
my_contours <- st_intersection(select_countries, 
                               st_union(lldata_poly, is_coverage = T))

# dowload tiles and compose raster (SpatRaster)
nc_osm <- get_tiles(my_contours, 
                      provider = "Esri.WorldShadedRelief", 
                      zoom = 7, crop = F)
poly_cell <- merge(lldata_poly, my_grid, by = c("long", "lat"))
# coordinates of the centroid
coord_temp <- st_transform(poly_cell, my_proj)
# simplify the geometry
poly_cell <- st_intersection(poly_cell, my_contours)
```

We compute all the potential clusters:

```{r, warning = F}
  coord_proj <- st_coordinates(st_centroid(coord_temp))
  pairs_geo <- find_all_cluster(coord_proj)
  dist_proj <- as(dist(cbind(coord_proj[, 1], coord_proj[, 2])), "matrix")

  dist_4326 <- as(dist(cbind(poly_cell$long, poly_cell$lat)), "matrix")
  coord_4326 <- st_transform(poly_cell, 4326) 
```

```{r}
  my_var <- "hwd_upp_" 
  temp_var <- poly_cell[, paste0(my_var, unique_year)[chosen_years]]
  MatX <- as.matrix(st_drop_geometry(temp_var))
```

#### Descriptive Analysis

We first plot the data:

```{r, fig.width = 10, height = 4}
y_lim <- range(MatX)
title_var <- 'Heat wave duration (in days)'
#pdf(file = "figures/VEN_data.pdf", width = 11, height = 4.5)
par(mfrow = c(1, 2), mar = c(3.7, 3, 1, 1), oma = c(0, 0, 0, 0),
    las = 1, mgp = c(2.2, 0.8, 0))
# map
plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

# data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = title_var, xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
abline(h = seq(0, 250, by = 50),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
#dev.off()
```

We map the average of the variable over a 3-year window. 

```{r, fig.width = 12, fig.height= 6}
nb_split <- 8
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/VEN_evol.pdf"), width = 12, height = 5)
par(mfrow = c(2, 4), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 7, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlGnBu"
    my_pal <- rev(alpha(colorspace::sequential_hcl(7, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
      title(paste0(my_years[1], "-", my_years[length(my_years)]), line = -1.25)
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Heatwave days")
    }
#dev.off()
```

Average:

```{r, fig.width = 12, fig.height= 6}
nb_split <- 1
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/VEN_mean.pdf"), width = 7, height = 5)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 7, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlGnBu"
    my_pal <- rev(alpha(colorspace::sequential_hcl(7, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
   #   title(paste0(my_years[1], "-", my_years[length(my_years)]), line = -1.25)
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Heatwave days")
    }
#dev.off()
```

#### NPFSS method

**Most likely cluster**

```{r}
res_np <- compute_np(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_np
```

**Significance (using parallel computing)**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np$stat < x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_np$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_np$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_np_2 <- compute_np(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_np_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np_2$stat < x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


#### PFSS method

**Most likely cluster**

```{r}
res_p <- compute_p(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_p
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p$stat <x))) /  (1 + B))
```



```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_p$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_p$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_p_2 <- compute_p(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_p_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p_2$stat <x))) /  (1 + B))
```

```{r, echo = F}
cat("p-value:  0.001")
```




#### DFFSS method

**Most likely cluster**

```{r}
res_dffss <- compute_dffss(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_dffss
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_p$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_dffss$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_dffss$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```

```{r}
res_dffss_2 <- compute_dffss(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_dffss_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_dffss_2$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```



### HFSS method 

**Most likely cluster**

We first determine the value of $d$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV.pdf"), width = 6, height = 4)
temp <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), 
                           d = ncol(MatX), plot_eigen = T)
#dev.off()
```

We choose $d=6$:
```{r}
res_h <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), d = 6)
res_h
```

To compute the significance, we make $B$ permutations on the data and compute the number of times the scan statistic is lower than the observed one: 

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, d = 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster 1**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_h$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_h$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


We look for an optimal value of $d$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV_2.pdf"), width = 6, height = 4)
temp <- compute_h(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], t(MatX), 
                           d = ncol(MatX), plot_eigen = T)
#dev.off()
```

We choose $d=6$.

```{r}
res_h_2 <- compute_h(cluster_g1_temp[-id_pos], 
                     cluster_g2_temp[-id_pos], t(MatX), d = 6)
res_h_2
```


**Significance of the secondary cluster 1**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, d = 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h_2$stat <x))) /  (1 + B))
```



```{r, echo = F}
cat("p-value:  0.001")
```




### Summary of the results

**Visualization of the result**

```{r}
res <- vector("list", 4)
res[[1]][[1]] <- res_h
res[[1]][[2]] <- res_h_2
res[[2]][[1]] <- res_dffss
res[[2]][[2]] <- res_dffss_2
res[[3]][[1]] <- res_np
res[[3]][[2]] <- res_np_2
res[[4]][[1]] <- res_p
res[[4]][[2]] <- res_p_2
```

```{r, fig.width = 12, fig.height = 6, message = F, warning = F}
my_var <- 'Heat wave duration (in days)'
dates <- unique_year[chosen_years]
y_lim <- range(MatX)

for(k in 1:4) {
  my_cluster_1 <- res[[k]][[1]]$vec
  my_cluster_2 <- res[[k]][[2]]$vec

#pdf(file = paste0("figures/", my_country, "_", names_method[k], ".pdf"), width = 13, height = 4.2) 
sf_use_s2(F)
nf <- layout( matrix(c(1,1,2,3), nrow=2, byrow=F) )
  par(mar = c(1.5, 0, 0, 0.2), 
      oma = c(0.5, 0, 2.4, 0), mgp = c(2.4, 0.6, 0), las = 1)
  ##### Map #########
    # map
  col_geo <- rep(rgb(0.9, 0.9, 0.9, alpha = 0.1), nrow(poly_cell))
  cex_geo <- rep(0.7, nrow(poly_cell))
  col_geo[my_cluster_1] <- alpha(cols[1], 0.8)
  cex_geo[my_cluster_1] <- 1

  col_geo[my_cluster_2] <- alpha(cols[2], 0.5)
  cex_geo[my_cluster_2] <- 1
  
  plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  mf_shadow(st_union(poly_cell[my_cluster_1, ]), 
                add = T, cex = 0.8, col = cols[1])
  mf_shadow(st_union(poly_cell[my_cluster_2, ]), 
         add = T, col= cols[2], cex = 0.8)
      
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

  
  plot(st_geometry(poly_cell), 
        border = "white",
        col = col_geo, 
        cex = cex_geo,
        pch = 16, asp = 1, add = T, lwd = 0.1)

  plot(st_geometry(st_union(poly_cell[my_cluster_1, ])), 
         add = T, border= cols[1], col = NULL)
  plot(st_geometry(st_union(poly_cell[my_cluster_2, ])), 
         add = T, border= cols[2], col = NULL)
      
  temp_1 <- draw.circle(coord_proj[my_cluster_1[1], 1], 
                           coord_proj[my_cluster_1[1], 2], 
                  as.numeric(dist_proj[my_cluster_1[1], 
                                       my_cluster_1[length(my_cluster_1)]]))

    my_circle_1 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_1$x, temp_1$x[1]), 
             c(temp_1$y, temp_1$y[1]))
         )), crs = my_proj
       ), 4326)
    
  temp_2 <- draw.circle(coord_proj[my_cluster_2[1], 1], 
                           coord_proj[my_cluster_2[1], 2], 
                  as.numeric(dist_proj[my_cluster_2[1], 
                                       my_cluster_2[length(my_cluster_2)]]))
    
  my_circle_2 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_2$x, temp_2$x[1]), 
             c(temp_2$y, temp_2$y[1]))
         )), crs = my_proj
       ), 4326)
  
  
  ###############
  plot(st_geometry(my_circle_2), add = T, border= cols[2],
             col = alpha(cols[2], 0.4), lty=1, lwd=1)
  plot(st_geometry(my_circle_1), add = T, border= cols[1],
             col = alpha(cols[1], 0.4), lty=1, lwd=1)
  
    mtext(my_var, side = 4, line = -3.5, las = 0, cex = 0.8)
    
  legend("topleft", legend = c("Most likely cluster", "Secondary cluster 1"),
         fill = c(cols[1], cols[2]), cex = 0.9, box.lty = 0)
      
  ##### Functional data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = '',
       ylab = '', col = rgb(0.6, 0.6, 0.6, alpha = 0.5), xaxt = 'n', 
        type = "l")
  legend("topleft", legend = c("Most likely cluster"),
         lty = 1, col = c(cols[1]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 2500, by = 200),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))

  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
    
  for(i in my_cluster_1)
        lines(dates, MatX[i, ], col = alpha(cols[1], 0.3),
          lty = 1, lwd = 1.3)
  
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  

plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = my_var, xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  legend("topleft", legend = c("Secondary cluster 1"),
         lty = 1, col = c(cols[2]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 2500, by = 200),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
      
  for(i in my_cluster_2)
        lines(dates, MatX[i, ], col = alpha(cols[2], 0.3),
          lty = 1, lwd = 1.3)
  
    
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  
  mtext(paste0("Clusters for the ", names_method[k]), side = 3, line = 0.8, outer = TRUE)
#dev.off()
}
```

We present in the following table the results obtained by the different methods. 


```{r}
res_VEN <- data.frame(nb_cluster_1 = c(length(res_np$vec), 
                            length(res_p$vec),
                            length(res_dffss$vec),
                            length(res_h$vec)),
           sign_cluster_1 = c(0.001, 0.001, 0.001, 0.001),
           nb_cluster_2 = c(length(res_np_2$vec), 
                            length(res_p_2$vec),
                            length(res_dffss_2$vec),
                            length(res_h_2$vec)),
           sign_cluster_2 = c(0.001, 0.001, 0.001, 0.001))
row.names(res_VEN) <- c("NPFSS", "PFSS", "DFFSS", "HFSS")
knitr::kable(res_VEN)
```

