---
title: "Supplementary material: Codes and Data for the paper ``A new parametric spatial scan statistic for functional data: application to climate change data''"
author: "Zaineb Smida, Thibault Laurent, Lionel Cucala"
date: "Last update: `r Sys.Date()`"
header-includes:
- \usepackage{booktabs}
- \usepackage{makecell}
output:
  github_document:   
    number_sections: yes
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(fig.align = "center",
               cache = F,
               cache.lazy = FALSE
)
options(kableExtra.latex.load_packages = FALSE)
```


This document provides the R code and data used to generate the computational results featured in the paper "A New Parametric Spatial Scan Statistic for Functional Data: Application to Climate Change Data." To reference this work, please use the following citation:

Zaineb Smida, Thibault Laurent and Lionel Cucala (2024). [A Hotelling spatial scan statistic for functional data: application to economic and climate data](), *TSE WP*.

Packages needed:

```{r, message = F, warning = F}
library(mapsf) # cartography
library(maptiles) # import spatial contours
library(sf) # spatial data analysis
library(tidyverse) # tidyverse
library(latex2exp) # add LaTeX
library(ggh4x) # customize ggplot graphic
library(progress) # progress bar
library(rARPACK) # compute only the d first eigen values/eigen vectors
library(parallel) # parallel computing
```

The document is organized into three parts:

* Part 1: Details the functions developed specifically for this work.
* Part 2: Provides the code necessary to reproduce the results presented in the Simulation Study section.
* Part 3: Contains the code to replicate the results from the Application to Real Data section.


# Developed Functions

## Functional Data Simulation

The function `simulvec()` is used to simulate functional data as described in the Simulation Study section of the article. It accepts the following arguments:

* `npoints`: The number of measurements.
* `shape`: The distribution of the random variable $Z$, which can be one of the following: `"gauss"`, `"student"`, `"chisq"`, or `"exp"`.

```{r class.source = 'fold-hide'}
simulvec <- function(npoints, shape = "gauss") {
  # initi
  vect <- (0:npoints)/npoints
  k <- 1
  sigma <- 1/((k - 0.5) * pi)
  
  vecyphi <- sqrt(2) * sin(vect/sigma)
  
  if(shape == "gauss") {
    y <- rnorm(1, mean = 0, sd = sigma)
  } else {
    if(shape == "student") { 
      y <- sigma * rt(1, 4)
    } else  {
      if(shape == "chisq") { 
        y <- sigma * rchisq(1, 4)
      } else {
        y <- sigma * rexp(1, 0.5)
      }
    }
  }
  vecY <- y * vecyphi
  exvecY <- vecY
  
  flag <- TRUE
  k <- 2
  while (flag) {
    
    sigma <- 1/((k - 0.5) * pi)
    if(shape == "gauss") {
      y <- rnorm(1, mean = 0, sd = sigma)
    } else {
      if(shape == "student") { 
        y <- sigma * rt(1, 4)
      } else  {
        if(shape == "chisq") { 
        y <- sigma * rchisq(1, 4)
        } else {
          y <- sigma * rexp(1, 4)  
        }
      }
    }
    
    vecyphi <- sqrt(2) * sin(vect/sigma)
    
    exvecY <- vecY
    vecY <- vecY + y * vecyphi
    flag <- (sum((vecY - exvecY) ^ 2) / sum((vecY) ^ 2) > 0.001)
    k <- k + 1
  }
  vecY
}
```

**Example**: Simulating a sample of 50 functions, each measured at 100 equidistant points.

```{r}
nobs <- 50
npoints <- 75
X <- matrix(0, npoints + 25, nobs)
set.seed(777)
for (k in 1:nobs) {
  X[, k] <- simulvec(npoints + 24, shape = "gauss") 
}
```

**Note:** It is common to discard the initial simulated data. In this example, we plot the full set of functional data (figure on the left) and retain only the last 75 values (figure on the right).

```{r, fig.width = 10, fig.height = 4.5}
par(oma = c(0, 0, 0, 0), mar = c(3, 3, 1, 1), las = 1, mfrow = c(1, 2))
matplot(X, type = "l", lty = 1, col ="grey")
abline(v = 25, lty = 2)
X <- X[26:100, ]
matplot(X, type = "l", lty = 1, col ="grey")
```



## Identify All Potential Clusters

The `find_all_cluster()` function accepts a matrix of Cartesian coordinates as input and returns all potential clusters. The output consists of two elements: the first element contains a list of all potential clusters, and the second element contains a list of their corresponding complements.

If the geographical coordinates are provided in Longitude/Latitude format, it is recommended to convert them to an appropriate Coordinate Reference System (CRS). For assistance, you can refer to https://epsg.io and use the **sf** package for the transformation.

```{r}
find_all_cluster <- function (Matcoord) {
  n <- nrow(Matcoord)
  Matdist <- as.matrix(dist(Matcoord, upper = TRUE))
  vecord_list <- vector("list", n)
  for (k in 1:n) {
    vecord_list[[k]] <- order(Matdist[, k])
  }
  res_cluster_g1 <- vector("list", 0)
  res_cluster_g2 <- vector("list", 0)
  
  matrix_g1 <- vector("list", n)

  for(k in 1:n) {
    matrix_g1[[k]] <- rep(0, n)
  }
  
  nb_combi <- 0
  for (k in 1:(n-1)) {
    for (j in 1:n) {
      temp_1 <- vecord_list[[j]][1:k]
      temp_2 <- vecord_list[[j]][(k+1):n]

      my_vec <- my_vec_2 <-numeric(n)
      my_vec[temp_1] <- 1
      my_vec_2[temp_2] <- 1
      # my_length <- k
      cond_1 <-  any(matrix_g1[[k]] %*% my_vec == k)
      cond_2 <-  any(matrix_g1[[n-k]] %*% my_vec_2 == n-k)
        
      if (!(any(cond_1) | any(cond_2))) {
        nb_combi <- nb_combi + 1
        res_cluster_g1[[nb_combi]] <- temp_1
        res_cluster_g2[[nb_combi]] <- temp_2
        matrix_g1[[k]] <- rbind(matrix_g1[[k]], my_vec)
    #    matrix_g2[[n-k]] <- rbind(matrix_g2[[n-k]], rep(1, n) - my_vec)
      }
    }
  }
  cat("Number of unique combination: ", nb_combi, "\n")
  return(list(vec_g1 = res_cluster_g1,
              vec_g2 = res_cluster_g2))
}
```

**Example:** We consider a random spatial point process with 50 observations.

```{r, fig.width = 5, fig.height = 5}
set.seed(1)
matCoord <- cbind(runif(nobs), runif(nobs))
par(las = 1, mar = c(4, 4, 0.5, 0.5), mgp = c(2.2, 0.7, 0))
plot(matCoord, xlab = "Longitude", ylab = "Latitude", asp = 1)
```

We identify all potential spatial clusters.

```{r}
my_pairs_ex <- find_all_cluster(matCoord)
```

**Note**: Once the potential clusters are identified, this reduces the number of combinations to test from 2450 ($50 \times 49$) to 1644.


## Representation of a circle in a map

The `draw.circle()` function returns the coordinates of a circle of radius `radius` and centered at the coordinates `x` and `y`:

```{r}
draw.circle <- function (x, y, radius, nv = 100) {
  ymult <- 1
  angle.inc <- 2 * pi/nv
  angles <- seq(0, 2 * pi - angle.inc, by = angle.inc)
  for (circle in 1:length(radius)) {
    xv <- cos(angles) * radius[circle] + x
    yv <- sin(angles) * radius[circle] * ymult + y
  }
  invisible(list(x = xv, y = yv))
}
```

**Example**: We consider a hypothetical cluster, $C$, consisting of eight observations from the previously generated data. The cluster is centered around observation 50 and includes the eight closest observations to it. The radius of the cluster is defined as the distance between observation 50 and the farthest observation within the cluster (observation 13).

```{r}
my_cluster <- c(50, 9, 15, 8, 43, 17, 32, 13)
my_dist <- dist(matCoord[c(50, 13), ])
```


```{r, fig.width = 5, fig.height = 5}
plot(matCoord, xlab = "x", ylab = "y", asp = 1)
points(matCoord[my_cluster, ], pch = 16, col = "red")
temp_plot <- draw.circle(matCoord[50, 1], matCoord[50, 2], 
                         my_dist, nv = 100)
lines(temp_plot, col = "red")
```

For the remainder of this section, we modify the functional data of the hypothetical cluster. We apply a shift $\Delta_2(t) = \alpha t(1 - t)$, with $\alpha = 2$. Our goal is to detect the cluster using various methods.

```{r}
t.disc <- (1:75) / (75)
for(k in 1:8)
  X[, my_cluster[k]] <- X[, k] + 2 * t.disc
```

```{r, fig.width = 10, fig.height = 4.5}
par(oma = c(0, 0, 0, 0), mar = c(3, 3, 1, 1), las = 1, mfrow = c(1, 2))
# The function
matplot(X, type = "l", lty = 1, col ="grey")
matplot(X[, my_cluster], type = "l", lty = 1, col ="red", add = T)
# The map
plot(matCoord, xlab = "x", ylab = "y", asp = 1)
points(matCoord[my_cluster, ], pch = 16, col = "red")
lines(temp_plot, col = "red")
```


## Package HDSpatialScan

The `SpatialScan()` function from the `HDSpatialScan` package (Frévent et al., 2021) can be used to compute various methods. Its main arguments include a vector of method names, the spatial coordinates, and the functional data of the observations. Additional parameters can also be specified, such as the minimum and maximum cluster sizes to be detected, the number of replications for significance testing, and more.

For example, to compute the methods ("DFFSS", "PFSS", "NPFSS"), the `SpatialScan()` function can be used as follows:
 
```{r, eval = F}
fss_result <- HDSpatialScan::SpatialScan(c("NPFSS", "PFSS", "DFFSS"),
        t(X), sites_coord = matCoord, mini = 1, maxi = 49,
        system = "Euclidean", MC=99, typeI = 0.25)
```

For practical reasons, we implement our own functions and use them in our simulation framework.

## Functions for Scan Statistic 

We implement the functions `compute_np()`, `compute_p()`, `compute_dffss()` and 
`compute_h()`, which correspond to the four methods "NPFSS", "PFSS", "DFFSS", and "HFSS", respectively. Each function takes as arguments the list of potential clusters `c1` (and, correspondingly, the complement of the potential clusters `c2`) and the functional data `my_mat`. They return the highest test statistic value found across all possible combinations, along with the associated cluster. These functions are available in the file [functions_to_cluster.R](functions_to_cluster.R)  

```{r}
source("codes/functions_to_cluster.R")
```


### Non Parametric NPFSS method

To compute the result for the NPFSS method, we use the `compute_np()` function. This yields the following test statistic and associated cluster.

```{r}
res_np <- compute_np(my_pairs_ex[[1]], my_pairs_ex[[2]], X)
res_np
```

The detected cluster contains `r length(res_np$vec)` observations. To assess significance, we perform $B$ permutations on the data and count how many times the scan statistic is less than the observed value. nhh

```{r, eval = F}
p_value_np <- 0
B <- 99
pb <- progress_bar$new(total = B)

set.seed(123)
for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(X))
  MatXsim <- X[, perm]
  temp <- compute_np(my_pairs_ex[[1]], my_pairs_ex[[2]], MatXsim)
  p_value_np <- p_value_np + (res_np$stat < temp$stat)
}
cat("p-value: ", p_value_np / 100)
```

```{r, echo = F}
cat("p-value:  0.31")
```


In this example, the cluster detected by the NPFSS method is not significant.


### Parametric PFSS

To obtain the result for the PFSS method, we use the `compute_p()` function.

```{r}
res_p <- compute_p(my_pairs_ex[[1]], my_pairs_ex[[2]], X)
res_p
```
The detected cluster contains `r length(res_p$vec)` observations. To assess significance, we perform $B$ permutations on the data and count the number of times the scan statistic is less than the observed value.

```{r, eval = F}
p_value_p <- 0
B <- 99
pb <- progress_bar$new(total = B)
set.seed(123)
for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(X))
  MatXsim <- X[, perm]
  temp <- compute_p(my_pairs_ex[[1]], my_pairs_ex[[2]], MatXsim)
  p_value_p <- p_value_p + (res_p$stat < temp$stat)
}
p_value_p / 100
```

```{r, echo = F}
cat("p-value:  0.04")
```

The detected cluster is significant. Among the `r length(res_p$vec)` observations, `r length(intersect(res_p$vec, my_cluster))` belong to the real cluster. 

### Method DFFSS

To obtain the result for the DFFSS method, we use the `compute_dffss()` function.

```{r}
res_dffss <- compute_dffss(my_pairs_ex[[1]], my_pairs_ex[[2]], X)
res_dffss
```

The detected cluster contains `r length(res_dffss$vec)` observations. To assess significance, we perform $B$ permutations on the data and count how many times the scan statistic is less than the observed value.

```{r, eval = F}
p_value_dffss <- 0
B <- 99
pb <- progress_bar$new(total = B)
set.seed(123)
for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(X))
  MatXsim <- X[, perm]
  temp <- compute_dffss(my_pairs_ex[[1]], my_pairs_ex[[2]], MatXsim)
  p_value_dffss <- p_value_dffss + (res_dffss$stat < temp$stat)
}
p_value_dffss / 100
```


```{r, echo = F}
cat("p-value:  0.03")
```

The detected cluster is significant. Among the `r length(res_dffss$vec)` observations, `r length(intersect(res_dffss$vec, my_cluster))` are part of the actual cluster.


### Method HFSS

To obtain the result for the HFSS method, we use the `compute_h()` function. An additional argument allows for the selection of the optimal $K$.

The `plot_eigen` argument enables the plotting of the CPV function. In this example, we recommend choosing 
$K=5$.


**Example**:
```{r, fig.width = 6, fig.height = 3}
res_h <- compute_h(my_pairs_ex[[1]], my_pairs_ex[[2]], X, npoints,
                   plot_eigen = T)
```

**Example** when $K=5$:
```{r}
res_h <- compute_h(my_pairs_ex[[1]], my_pairs_ex[[2]], X, 5)
res_h
```

To assess significance, we perform $B$ permutations on the data and count how many times the scan statistic is less than the observed value.

```{r, eval = F}
p_value_h <- 0
B <- 99
pb <- progress_bar$new(total = B)
set.seed(123)
for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(X))
  MatXsim <- X[, perm]
  temp <- compute_h(my_pairs_ex[[1]], my_pairs_ex[[2]], MatXsim, 5)
  p_value_h <- p_value_h + (res_h$stat < temp$stat)
}
p_value_h / 100
```

```{r, echo = F}
cat("p-value:  0.02")
```

The detected cluster is significant. Among the `r length(res_h$vec)` observations, `r length(intersect(res_h$vec, my_cluster))` are part of the actual cluster.

**Improving Computational Time**: If `plot_eigen = TRUE`, the spectral decomposition of the covariance matrix is performed using the `eigen()` function, which computes all eigenvalues and eigenvectors. Conversely, if `plot_eigen = FALSE`, we utilize the `eigs_sym()` function from the `rARPACK` package, which computes only the first $K$ eigenvalues and eigenvectors. This approach improves computation time by a factor of 3.


# Simulation part

## The spatial data

First, we import the contours of the French departments:

```{r, warning = F, message = F}
dep <- read_sf("data/departements.geojson")
dep <- dep[!dep$code %in% c("2A", "2B"), ]
dep <- dep[order(dep$code), ]
my_region <- st_union(dep)
nc_osm <- get_tiles(dep, provider = "Esri.WorldShadedRelief", 
                      zoom = 7, crop = T)
```

Next, we compute the Cartesian coordinates of the centroids of the departments in the official French Coordinate Reference System (CRS 2154), ensuring that distances between locations are measured in meters.

```{r, warning = F, message = F}
my_proj <- 2154
dep_proj <- st_transform(dep, 2154)
Matcoord <- st_coordinates(st_centroid(dep_proj))
dist_proj <- as(dist(Matcoord), "matrix")
```

We highlight the departments surrounding the city of Paris (74, 92, 91, 93, 77, 90, 94, 76), which will be simulated differently from the other departments to represent the cluster we aim to detect.

```{r, warning = F, message = F}
cols = c("#D35C79", "#009593")
# id of the cluster
vecclus <- c(74, 92, 91, 93, 77, 90, 94, 76)
# graphical parameters
col_geo <- rep(rgb(0.9, 0.9, 0.9, alpha = 0.1), nrow(Matcoord))
col_geo[vecclus] <- alpha(cols[1], 0.8)
```

We represent Figure S1 on the left:

```{r, warning = F, message = F}
#pdf("figures/french_cluster.pdf", width = 7, height = 7)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
plot_tiles(nc_osm, adjust = T)
mf_shadow(st_geometry(st_union(dep[vecclus, ])), add = T, cex = 0.5)
plot(st_geometry(dep), border = "white", col = col_geo,  
     add = T, lwd = 0.1)
plot(st_geometry(my_region), add = T, lwd = 0.5)
#text(par()$usr[1] + 0.03 * (par()$usr[2] - par()$usr[1]), 
#     par()$usr[4] - 0.07 * (par()$usr[4] - par()$usr[3]), 
#     labels = "A)", pos = 4, cex = 2)
#dev.off()
#R.utils::compressPDF("figures/french_cluster.pdf")
```

## The different shifts/probabilistic models

We consider three types of shift:

* $\Delta_1(t)=\alpha t$,
* $\Delta_2(t)=\alpha  t(1-t)$,
* $\Delta_3(t)=\alpha  \exp(-100(t-0.5)^2)/3$,

and four different probabilistic models $Z_{i,k}/\sigma_k$:

* a Gaussian $N(0,1)$,
* a Student $t(4)$
* a Chi $\chi^2(4)$
* an Exponential $e(4)$

with $i=1,\ldots,94$ and $t=1,\ldots,200$. The first 100 simulations are discarded to allow the process to converge. For each combination of shift and probabilistic method, we consider different values of $\alpha$.

```{r}
nobs <- nrow(Matcoord)
npoints <- 100
ndrop <- 100
t.disc <- (1:(npoints)) / (npoints) 
veccluster <- rep(0, nobs)
veccluster[vecclus] <- 1 
```

  
For $\alpha=3$, we plot an example of simulations for each combination of shift and probabilistic model:

```{r}
alpha <- 10
X_aggregated <- data.frame(
  x = integer(),
  y = numeric(), 
  shift = character(),
  proba = character()
) 
for(type_shift in 2:4) {
  for(shape in c("gauss", "student", "chisq", "exp")) {
    X <- matrix(0, npoints+ndrop, nobs)
    for (k in 1:nobs) {
      X[, k] <- simulvec(npoints+(ndrop-1), shape = shape) 
      if(type_shift == 1) {
        X[(ndrop+1):nrow(X), k] <- X[(ndrop+1):nrow(X), k] + alpha * (veccluster[k] == 1)
        } else {
          if (type_shift == 2) {
            alpha <- 3
            X[(ndrop+1):nrow(X), k] <- X[(ndrop+1):nrow(X), k] + alpha * t.disc * (veccluster[k] == 1)
            } else {
              if (type_shift == 3) {
                alpha <- 10
                X[(ndrop+1):nrow(X), k] <- X[(ndrop+1):nrow(X), k] + alpha * t.disc * (1 - t.disc) * (veccluster[k] == 1)
                } else {
                  alpha <- 10
                  X[(ndrop+1):nrow(X), k] <- X[(ndrop+1):nrow(X), k] + alpha * exp(-100 * (t.disc - 0.5) ^ 2) / 3 * (veccluster[k] == 1)
                }
            }
        }
    }
    # drop first observation 
    X <- X[-(1:ndrop), ]
    # aggregate data 
    X_aggregated <- rbind(X_aggregated,
                          data.frame(
                            x = seq(0, 1, length.out = 100),
                            y = as.vector(X), 
                            shift = type_shift,
                            proba = shape,
                            id = rep(1:94, each = 100)
                          ))
  }
}
X_aggregated$shift <- factor(X_aggregated$shift) 
levels(X_aggregated$shift) = c(`2` = TeX("$\\Delta_1(t)$"), 
                               `3` = TeX("$\\Delta_2(t)$"), 
                               `4` = TeX("$\\Delta_3(t)$"))
X_aggregated$Cluster <- factor(ifelse(X_aggregated$id %in% vecclus, "Yes", "No"),
                               levels = c("No", "Yes"))
X_aggregated$proba <- factor(X_aggregated$proba, levels = c("gauss", "student", "chisq", "exp"))
levels(X_aggregated$proba) <- c(`gauss` = TeX("$N(0,1)$"), 
                                `student` = TeX("$t(4)$"), 
                                `exp` = TeX("$Exp(4)$"),
                                `chisq` = TeX("$\\chi^2(4)$"))
```

We represent Figure S2:

```{r}
temp <- X_aggregated[order(as.numeric(X_aggregated$Cluster), X_aggregated$id),]
X_aggregated %>%
  ggplot(aes(x = x, y = y, color = Cluster)) +
  geom_line(aes(group = id)) +
  theme_bw() +
  theme(strip.background = element_rect(color = "black", fill = alpha("#EE9E94", 0.1))) +
  facet_grid(rows=vars(proba),
  cols=vars(shift),
  labeller=label_parsed,
  scales = "free") +
      scale_colour_manual(values = c("grey", "tomato2")) +
  xlab(TeX("$t$")) +
  ylab(TeX("$X_i(t)$"))
# ggsave("figures/simu.pdf", width = 7, height = 6)
#R.utils::compressPDF("figures/simu.pdf")
```

## Results

We utilized a server with 94 cores, launching a unique parameter set for each core, comprising the shift, probabilistic model, and value of $\alpha$. The computation time was approximately 7 days. We repeated the following procedure on each core:

```{r}
parms_df <- rbind(
  # line 1 
  data.frame(
    shape = "gauss", type_shift = 2, alpha = seq(0, 3, 0.5), sizeclust = 8),
  data.frame(
    shape = "gauss", type_shift = 3, alpha = seq(0, 10.5, 1.5), sizeclust = 8),
  data.frame(
    shape = "gauss", type_shift = 4, alpha = seq(0, 12, 2), sizeclust = 8),  
  # line 2
  data.frame(
    shape = "student", type_shift = 2, alpha = seq(0, 7, 1), sizeclust = 8),
  data.frame(
    shape = "student", type_shift = 3, alpha = seq(0, 14, 2), sizeclust = 8),
  data.frame(
    shape = "student", type_shift = 4, alpha = seq(0, 14, 2), sizeclust = 8),  
  # line 3
  data.frame(
    shape = "chisq", type_shift = 2, alpha = seq(0, 14, 2), sizeclust = 8),
  data.frame(
    shape = "chisq", type_shift = 3, alpha = seq(0, 14, 2), sizeclust = 8),
  data.frame(
    shape = "chisq", type_shift = 4, alpha = seq(0, 14, 2), sizeclust = 8),  
  # line 4
  data.frame(
    shape = "exp", type_shift = 2, alpha = seq(0, 7, 1), sizeclust = 8),
  data.frame(
    shape = "exp", type_shift = 3, alpha = seq(0, 14, 2), sizeclust = 8),
  data.frame(
    shape = "exp", type_shift = 4, alpha = seq(0, 14, 2), sizeclust = 8)
)
```

Repeat the following procedure 200 times:

* Simulate a set of functional data (with a shift in the Paris region).
* Compute the scan statistics for the methods: "HFSS," "PFSS," "NPFSS," and "DFFSS."
* Generate 199 permutation samples and compute the scan statistics for each to assess significance.

Based on the significance results and the detected clusters, we calculate:

* The power,
* The percentage of true positives,
* The percentage of false negatives.

### Choice of the optimal $K$

For the HDFS method, we use the Cumulative Proportion of Variance (CPV) criterion to determine the optimal $K$.

Empirically, we generate CPV curves for several simulations across different values of $\alpha$, the three shifts, and the four probabilistic models.

We represent Figure S3:

```{r, eval = F}
# parameters of simulation 
nobs <- nrow(Matcoord)
npoints <- 100
ndrop <- 100
t.disc <- (1:(npoints)) / (npoints) # (1:(npoints+ndrop)) / (npoints+ndrop)

vecclus <- c(74, 92, 91, 93, 77, 90, 94, 76)
nclus <- length(vecclus)
veccluster <- numeric(nobs)
veccluster[vecclus] <- 1

# the possible clusters
my_pairs <- find_all_cluster(Matcoord)

# restriction on the size of the cluster 
mini <- 2 
maxi <- trunc(nobs / 2)
cluster_g1 <- my_pairs[[1]] 
cluster_g2 <- my_pairs[[2]] 
size_group <- sapply(cluster_g1, length)
ind_restriction <- size_group >= mini & size_group <= maxi
cluster_g1 <- cluster_g1[ind_restriction]
cluster_g2 <- cluster_g2[ind_restriction] 
# number of combinaison
nb_combi <- length(cluster_g1)
K_choice <- data.frame(
  x = numeric(),
  y = numeric(),
  shift = character(),
  proba = character(),
  alpha = integer()
)
shift_levels <- levels(X_aggregated$shift)
proba_levels <- levels(X_aggregated$proba)
nsimu <- 10

for(alpha in c(1, 3, 5, 10)) {
  print(alpha)
  for (shape in c("gauss", "student", "chisq", "exp")) {
    for (type_shift in 2:4) {
      cpv <- numeric(npoints)
      # simulations 
      i_simu <- 1
      while (i_simu <= nsimu) {
        X <- matrix(0, npoints+ndrop, nobs)
        for (k in 1:nobs) {
          X[, k] <- simulvec(npoints+(ndrop-1), shape = shape) 
          if(type_shift == 1) {
            X[(ndrop+1):nrow(X), k] <- X[(ndrop+1):nrow(X), k] + alpha * (veccluster[k] == 1)
            } else {
              if (type_shift == 2) {
                X[(ndrop+1):nrow(X), k] <- X[(ndrop+1):nrow(X), k] + alpha * t.disc * (veccluster[k] == 1)
                } else {
                  if (type_shift == 3) {
                    X[(ndrop+1):nrow(X), k] <- X[(ndrop+1):nrow(X), k] + alpha * t.disc * (1 - t.disc) * (veccluster[k] == 1)
                    } else {
                      X[(ndrop+1):nrow(X), k] <- X[(ndrop+1):nrow(X), k] + alpha * exp(-100 * (t.disc - 0.5) ^ 2) / 3 * (veccluster[k] == 1)
                    }
                }
            }
        }
        # drop first observation 
        X <- X[-(1:ndrop), ]
  
        # initialization
        npoints <- nrow(X)
        my_dist <- as(dist(t(X)), "matrix")
  
        # observed stat
        for (i in (1:nb_combi)) {
          vecindin <- cluster_g1[[i]]
          vecindout <- cluster_g2[[i]]
    
          nx <- length(vecindin)
          ny <- length(vecindout)
    
          if(nx == 1) {
            myX <- matrix(X[,vecindin], nrow(X), 1)
            cov_1 <- matrix(0, npoints, npoints)
            } else {
              myX <- X[, vecindin]
              cov_1 <- cov(t(myX)) 
              }
          if(ny == 1) {    
            myY <- matrix(X[,vecindout], nrow(X), 1)
            cov_2 <- matrix(0, npoints, npoints)
            } else {
              myY <- X[, vecindout]
              cov_2 <- cov(t(myY)) # 
            }
          ##### Initi
          D <- 1 / (nx + ny - 2) * ((nx - 1) * cov_1  +  (ny - 1) * cov_2)
          D_hotelling <- (nx + ny) / (nx * ny) * D
          
          temp_svd <- eigen(D_hotelling)
          tau_k <- temp_svd$values
          cpv <- cpv + cumsum(tau_k) / sum(tau_k)
        }
        i_simu <- i_simu + 1
      } 
      K_choice <- rbind(K_choice,
                        data.frame(
                          x = 1:15,
                          y = cpv[1:15] / (nsimu * nb_combi),
                          shift = type_shift,
                          proba = shape,
                          alpha = alpha)
                        )
    }
  }
}
K_choice$shift <- factor(K_choice$shift) 
levels(K_choice$shift) = c(`2` = TeX("$\\Delta_1(t)$"), 
                               `3` = TeX("$\\Delta_2(t)$"), 
                               `4` = TeX("$\\Delta_3(t)$"))

K_choice$proba <- factor(K_choice$proba, levels = c("gauss", "student", "exp", "chisq"))
levels(K_choice$proba) <- c(`gauss` = TeX("$N(0,1)$"),
                            `student` = TeX("$t(4)$"),
                            `exp` = TeX("$Exp(4)$"),
                            `chisq` = TeX("$\\chi^2(4)$"))
shift <- rep(rep(c("delta_1", "delta_2", "delta_3"), each = 15), times = 16)
K_choice$shift <- shift
proba <- rep(rep(c("gauss", "student", "chisq", "exp"), each = 45), times = 4)
K_choice$proba <- proba

K_choice$shift <- factor(K_choice$shift) 
levels(K_choice$shift) = c(`delta_1` = TeX("$\\Delta_1(t)$"), 
                               `delta_2` = TeX("$\\Delta_2(t)$"), 
                               `delta_3` = TeX("$\\Delta_3(t)$"))
K_choice$proba <- factor(K_choice$proba, levels = c("gauss", "student", "exp", "chisq"))
levels(K_choice$proba) <- c(`gauss` = TeX("$N(0,1)$"),
                            `student` = TeX("$t(4)$"),
                            `exp` = TeX("$Exp(4)$"),
                            `chisq` = TeX("$\\chi^2(4)$"))
# save(K_choice, file = "results/K_choice.RData")
```


```{r}
load("results/K_choice.RData")
my_plot <- vector("list", 4)
my_alpha <- c(1, 3, 5, 10)
for(k in 1:4) {
  my_plot[[k]] <- K_choice %>%
    filter(alpha == my_alpha[k]) %>%
    ggplot(aes(x = x, y = y)) +
    geom_line() +
    geom_point(size = 0.5) +
    geom_vline(xintercept = 5, linetype="dotted", 
                color = "blue", size=.6) +
    theme_bw() +
    theme(strip.background = element_rect(color = "black", fill = alpha("#EE9E94", 0.1))) +
    facet_grid(rows=vars(proba),
               cols=vars(shift),
               labeller=label_parsed,
               scales = "free") +
    xlab("k") +
    ylab(TeX("Mean function of $\\bar{CPV}(k)$ curves for 10 simulations"))
}
cowplot::plot_grid(my_plot[[1]], my_plot[[2]], my_plot[[3]], my_plot[[4]], nrow = 2,
          labels =   c("a)", "b)", "c)", "d)")) # c(TeX("$\\alpha$"), TeX("$\\alpha$"), TeX("$\\alpha$")))
# ggsave("figures/choice_K.pdf", width = 10, height = 10)
# R.utils::compressPDF("figures/choice_K.pdf")
```

* We plot the theoretical CPV for the Gaussian case:

```{r, fig.width = 10, fig.height = 8}
my_sigma_k <- function(k)
  1 / (pi * (k - 0.5)) ^ 2

my_val <- my_sigma_k(1:100)
plot(cumsum(my_val) / sum(my_val), type = "b",
     xlab = "k", ylab = "CPV")
```

### Batch code

The codes used are given in the files [batch_cluster_size_8.R](codes/batch_cluster_size_8.R), by using the functions in [functions_to_cluster.R](codes/functions_to_cluster.R).


```{r, echo = F}
# we load the results obtained from the servor 
load("results/my_res_new_1.RData")
load("results/my_res_new_2.RData")
load("results/my_res_new_3.RData")
load("results/my_res_new_4.RData")
```

We present the results in a format that can be easily visualized in a graph:

```{r}
power_to_plot <- data.frame(
  shape = character(0),
  type_shift = integer(0),
  alpha = numeric(0)
)
nb_est <- 10
for(k in 1:nrow(parms_df)) {
  power_to_plot <- rbind(
    power_to_plot,
    parms_df[rep(k, nb_est), 1:3])
}
power_to_plot$method <- c("DFFSS", "PFSS", "NPFSS", "hotelling_1", "hotelling_2", 
                          "hotelling_3", "hotelling_4", "HFSS", "hotelling_10", 
                          "hotelling_15")

FP_to_plot <- TP_to_plot <- power_to_plot

for(k in 1:length(res_par_1)) {
  power_to_plot$value[(1:nb_est)+(k-1)*nb_est] <- (res_par_1[[k]]$power + res_par_2[[k]]$power + 
                                                     res_par_3[[k]]$power + res_par_4[[k]]$power) / 200 # res.final$power / 100 # 
  TP_to_plot$value[(1:nb_est)+(k-1)*nb_est] <- (res_par_1[[k]]$nTP + res_par_2[[k]]$nTP + 
                                                  res_par_3[[k]]$nTP + res_par_4[[k]]$nTP) /  
    (res_par_1[[k]]$power + res_par_2[[k]]$power + res_par_3[[k]]$power + res_par_4[[k]]$power) / 8 # res.final$nTP / res.final$power / 8 # 
  FP_to_plot$value[(1:nb_est)+(k-1)*nb_est] <- (res_par_1[[k]]$nFP + res_par_2[[k]]$nFP + 
                                                  res_par_3[[k]]$nFP + res_par_4[[k]]$nFP) /  
    (res_par_1[[k]]$power + res_par_2[[k]]$power + res_par_3[[k]]$power + res_par_4[[k]]$power) / 86 #res.final$nFP / res.final$power / 8 # res_par[[k]]$nFP / res_par[[k]]$power / 8
}
power_to_plot$criteria <- "power"
TP_to_plot$criteria <- "TP"
FP_to_plot$criteria <- "FP"
to_plot <- rbind(power_to_plot, TP_to_plot, FP_to_plot)
# we select the most interseting points
parms_df_select <- rbind(
  # line 1 
  data.frame(
    shape = "gauss", type_shift = 2, alpha = seq(0, 3, 0.5)[-2], sizeclust = 8),
  data.frame(
    shape = "gauss", type_shift = 3, alpha = seq(0, 10.5, 1.5)[-c(2, 8)], sizeclust = 8),
  data.frame(
    shape = "gauss", type_shift = 4, alpha = seq(0, 12, 2)[-2], sizeclust = 8),  
  # line 2
  data.frame(
    shape = "student", type_shift = 2, alpha = seq(0, 7, 1)[-c(2, 8)], sizeclust = 8),
  data.frame(
    shape = "student", type_shift = 3, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 8),
  data.frame(
    shape = "student", type_shift = 4, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 8),  
  # line 3
  data.frame(
    shape = "chisq", type_shift = 2, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 8),
  data.frame(
    shape = "chisq", type_shift = 3, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 8),
  data.frame(
    shape = "chisq", type_shift = 4, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 8),  
  # line 4
  data.frame(
    shape = "exp", type_shift = 2, alpha = seq(0, 7, 1)[-c(2, 8)], sizeclust = 8),
  data.frame(
    shape = "exp", type_shift = 3, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 8),
  data.frame(
    shape = "exp", type_shift = 4, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 8)
)
to_plot <- merge(parms_df_select, to_plot, by = c("shape", "type_shift", "alpha"))
# we give labels
to_plot$type_shift <- factor(to_plot$type_shift) 
levels(to_plot$type_shift) = c(`2` = TeX("$\\Delta_1(t)$"), # TeX("$\\Delta_1(t)=ct$"), 
                         `3` = TeX("$\\Delta_2(t)$"),  # TeX("$\\Delta_2(t)=ct(1-t)$"), 
                         `4` = TeX("$\\Delta_3(t)$")) #  TeX("$\\Delta_3(t)=\\frac{c}{3}\\exp(-100(t-0.5)^2)$"))
to_plot$shape <- factor(to_plot$shape, levels = c("gauss", "student", "exp", "chisq"))
levels(to_plot$shape) <- c(`gauss` = TeX("$N(0,1)$"), 
                                `student` = TeX("$t(4)$"), 
                                `exp` = TeX("$Exp(4)$"),
                                `chisq` = TeX("$\\chi^2(4)$"))
to_plot$method <- factor(to_plot$method, c("HFSS", "DFFSS", "PFSS", "NPFSS",
                                           "hotelling_1", "hotelling_2", "hotelling_3", "hotelling_4",
                                           "hotelling_5", "hotelling_10", "hotelling_15"))
```



### Delta 1 

```{r}
threshold <- data.frame(
  yintercept = c(0.05, NA, NA), 
  criteria = factor(c("power", "TP", "FP"), levels = c("power", "TP", "FP")))
to_plot$criteria <- factor(to_plot$criteria, levels = c("power", "TP", "FP"))
to_plot2 <- to_plot[-which(to_plot$alpha == 0 & to_plot$criteria == "TP"), ]
to_plot2 <- to_plot2[-which(to_plot2$alpha == 0 & to_plot2$criteria == "FP"), ]
delta_1 <- to_plot2 %>%
  filter(type_shift == "Delta[1](t)") %>%
  filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot2 %>% 
              filter(type_shift == "Delta[1](t)") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")), 
            aes(group = method)) +
  geom_point(size = 0.8, pch = 15) +
  geom_hline(data = to_plot2 %>% filter(criteria == "power"), 
               aes(yintercept = 0.05, 
                 linetype = "0.05")) +
  ggh4x::facet_grid2(rows=vars(criteria),
                     cols=vars(shape),
                     labeller=label_parsed,
                     scales = "free", 
                     axes = "margins")  +
  theme_bw() +
  xlab(TeX("$\\alpha$")) +
  ylab("") +
  ggtitle(TeX("$\\Delta_1$")) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_linetype_manual(name = "threshold", values = 2) +
  theme(strip.background = element_rect(colour = "black", 
        fill = alpha("#EE9E94", 0.1)))
delta_1
# ggsave("figures/simu_8_delta_1.pdf", width = 8, height = 6)
# R.utils::compressPDF("figures/simu_8_delta_1.pdf")
```


### Delta 2 

```{r}
threshold <- data.frame(
  yintercept = c(0.05, NA, NA), 
  criteria = factor(c("power", "TP", "FP"), levels = c("power", "TP", "FP")))
to_plot$criteria <- factor(to_plot$criteria, levels = c("power", "TP", "FP"))
to_plot2 <- to_plot[-which(to_plot$alpha == 0 & to_plot$criteria == "TP"), ]
to_plot2 <- to_plot2[-which(to_plot2$alpha == 0 & to_plot2$criteria == "FP"), ]
delta_2 <- to_plot2 %>%
  filter(type_shift == "Delta[2](t)") %>%
   # filter(alpha != 0 & criteria == "TP") %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot2 %>% 
              filter(type_shift == "Delta[2](t)") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")), 
            aes(group = method)) +
  geom_point(size = 0.8, pch = 15) +
#  geom_hline(aes(yintercept = yintercept, 
#                 linetype = criteria), data = threshold,
#             linetype = 2) +
    geom_hline(data = to_plot2 %>% filter(criteria == "power"), 
               aes(yintercept = 0.05, 
                 linetype = "0.05")) +
  ggh4x::facet_grid2(rows=vars(criteria),
                     cols=vars(shape),
                     labeller=label_parsed,
                     scales = "free", 
                     axes = "margins")  +
  theme_bw() +
  xlab(TeX("$\\alpha$")) +
  ylab("") +
  ggtitle(TeX("$\\Delta_2$")) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none") +
  scale_linetype_manual(name = "threshold", values = 2) +
  theme(strip.background = element_rect(colour = "black", 
        fill = alpha("#EE9E94", 0.1)))
delta_2
#ggsave("figures/simu_8_delta_2.pdf", width = 8, height = 6.5)
```


### Delta 3

```{r}
threshold <- data.frame(
  yintercept = c(0.05, NA, NA), 
  criteria = factor(c("power", "TP", "FP"), levels = c("power", "TP", "FP")))
to_plot$criteria <- factor(to_plot$criteria, levels = c("power", "TP", "FP"))
to_plot2 <- to_plot[-which(to_plot$alpha == 0 & to_plot$criteria == "TP"), ]
to_plot2 <- to_plot2[-which(to_plot2$alpha == 0 & to_plot2$criteria == "FP"), ]
delta_3 <- to_plot2 %>%
  filter(type_shift == "Delta[3](t)") %>%
   # filter(alpha != 0 & criteria == "TP") %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot2 %>% 
              filter(type_shift == "Delta[3](t)") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")), 
            aes(group = method)) +
  geom_point(size = 0.8, pch = 15) +
#  geom_hline(aes(yintercept = yintercept, 
#                 linetype = criteria), data = threshold,
#             linetype = 2) +
    geom_hline(data = to_plot2 %>% filter(criteria == "power"), 
               aes(yintercept = 0.05, 
                 linetype = "0.05")) +
  ggh4x::facet_grid2(rows=vars(criteria),
                     cols=vars(shape),
                     labeller=label_parsed,
                     scales = "free", 
                     axes = "margins")  +
  theme_bw() +
  xlab(TeX("$\\alpha$")) +
  ylab("") +
  ggtitle(TeX("$\\Delta_3$")) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_linetype_manual(name = "threshold", values = 2) +
  theme(strip.background = element_rect(colour = "black", 
        fill = alpha("#EE9E94", 0.1)))
delta_3
#ggsave("figures/simu_8_delta_3.pdf", width = 8, height = 6.5)
```


```{r}
cowplot::plot_grid(delta_2, delta_3, nrow = 2, rel_heights = c(1,1.1))
#ggsave("figures/simu_8_delta_23.pdf", width = 8, height = 11)
#R.utils::compressPDF("figures/simu_8_delta_23.pdf")
```


## Checking Robustess

We apply the same procedure while varying the size of the cluster. This time, we consider a cluster of size 10 and add two more departments to the previous list: 59 and 27.


```{r, warning = F, message = F}
# id of the cluster
vecclus <- c(74, 92, 91, 93, 77, 90, 94, 76, 59, 27)
# graphical parameters
col_geo <- rep(rgb(0.9, 0.9, 0.9, alpha = 0.1), nrow(Matcoord))
col_geo[vecclus] <- alpha("#D35C79", 0.8)
```

```{r, warning = F, message = F}
#pdf("figures/french_cluster_10.pdf", width = 7, height = 7)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
plot_tiles(nc_osm, adjust = T)
mf_shadow(st_geometry(st_union(dep[vecclus, ])), add = T, cex = 0.5)
plot(st_geometry(dep), border = "white", col = col_geo,  
     add = T, lwd = 0.1)
plot(st_geometry(my_region), add = T, lwd = 0.5)
#text(par()$usr[1] + 0.03 * (par()$usr[2] - par()$usr[1]), 
#     par()$usr[4] - 0.07 * (par()$usr[4] - par()$usr[3]), 
#     labels = "B)", pos = 4, cex = 2)
#dev.off()
#R.utils::compressPDF("figures/french_cluster_10.pdf")
```

The interpretations are the same as in the previous section.

```{r, echo = F}
# we load the results obtained from the servor 
load("results/my_res_new_10_1.RData")
load("results/my_res_new_10_2.RData")
load("results/my_res_new_10_3.RData")
load("results/my_res_new_10_4.RData")
```

```{r}
power_to_plot <- data.frame(
  shape = character(0),
  type_shift = integer(0),
  alpha = numeric(0)
)
nb_est <- 10
for(k in 1:nrow(parms_df)) {
  power_to_plot <- rbind(
    power_to_plot,
    parms_df[rep(k, nb_est), 1:3])
}
power_to_plot$method <- c("DFFSS", "PFSS", "NPFSS", "hotelling_1", "hotelling_2", 
                          "hotelling_3", "hotelling_4", "HFSS", "hotelling_10", 
                          "hotelling_15")
FP_to_plot <- TP_to_plot <- power_to_plot

for(k in 1:length(res_par_1)) {
  power_to_plot$value[(1:nb_est)+(k-1)*nb_est] <- (
    res_par_1[[k]]$power + res_par_2[[k]]$power +  
    res_par_3[[k]]$power  + res_par_4[[k]]$power) / 200 
  TP_to_plot$value[(1:nb_est)+(k-1)*nb_est] <- (
    res_par_1[[k]]$nTP + res_par_2[[k]]$nTP + 
      res_par_3[[k]]$nTP + res_par_4[[k]]$nTP) /  
    (res_par_1[[k]]$power + res_par_2[[k]]$power + 
       res_par_3[[k]]$power + res_par_4[[k]]$power) / 8 
  FP_to_plot$value[(1:nb_est)+(k-1)*nb_est] <- (
    res_par_1[[k]]$nFP + res_par_2[[k]]$nFP + 
      res_par_3[[k]]$nFP + res_par_4[[k]]$nFP) /  
    (res_par_1[[k]]$power + res_par_2[[k]]$power + 
       res_par_3[[k]]$power + res_par_4[[k]]$power) / 86 
}
power_to_plot$criteria <- "power"
TP_to_plot$criteria <- "TP"
FP_to_plot$criteria <- "FP"
to_plot <- rbind(power_to_plot, TP_to_plot, FP_to_plot)
# we select the most interseting points
parms_df_select <- rbind(
  # line 1
  data.frame(
    shape = "gauss", type_shift = 2, alpha = seq(0, 3, 0.5)[-c(2)], sizeclust = 10),
  data.frame(
    shape = "gauss", type_shift = 3, alpha = seq(0, 10.5, 1.5)[-c(2, 8)], sizeclust = 10),
  data.frame(
    shape = "gauss", type_shift = 4, alpha = seq(0, 12, 2)[-c(2, 8)], sizeclust = 10),
  # line 2
  data.frame(
    shape = "student", type_shift = 2, alpha = seq(0, 7, 1)[-c(2, 8)], sizeclust = 10),
  data.frame(
    shape = "student", type_shift = 3, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 10),
  data.frame(
    shape = "student", type_shift = 4, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 10),
  # line 3
  data.frame(
    shape = "chisq", type_shift = 2, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 10),
  data.frame(
    shape = "chisq", type_shift = 3, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 10),
  data.frame(
    shape = "chisq", type_shift = 4, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 10),
  # line 4
  data.frame(
    shape = "exp", type_shift = 2, alpha = seq(0, 7, 1)[-c(2, 8)], sizeclust = 10),
  data.frame(
    shape = "exp", type_shift = 3, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 10),
  data.frame(
    shape = "exp", type_shift = 4, alpha = seq(0, 14, 2)[-c(2, 8)], sizeclust = 10)
)
to_plot <- merge(parms_df_select, to_plot, by = c("shape", "type_shift", "alpha"))
# we give labels
to_plot$type_shift <- factor(to_plot$type_shift) 
levels(to_plot$type_shift) = c(`2` = TeX("$\\Delta_1(t)$"), # TeX("$\\Delta_1(t)=ct$"), 
                         `3` = TeX("$\\Delta_2(t)$"),  # TeX("$\\Delta_2(t)=ct(1-t)$"), 
                         `4` = TeX("$\\Delta_3(t)$")) #  TeX("$\\Delta_3(t)=\\frac{c}{3}\\exp(-100(t-0.5)^2)$"))
to_plot$shape <- factor(to_plot$shape, levels = c("gauss", "student", "exp", "chisq"))
levels(to_plot$shape) <- c(`gauss` = TeX("$N(0,1)$"), 
                                `student` = TeX("$t(4)$"), 
                                `exp` = TeX("$Exp(4)$"),
                                `chisq` = TeX("$\\chi^2(4)$"))
to_plot$method <- factor(to_plot$method, c("HFSS", "DFFSS", "PFSS", "NPFSS"))
```


### Delta 1 

```{r}
threshold <- data.frame(
  yintercept = c(0.05, NA, NA), 
  criteria = factor(c("power", "TP", "FP"), levels = c("power", "TP", "FP")))
to_plot$criteria <- factor(to_plot$criteria, levels = c("power", "TP", "FP"))
to_plot2 <- to_plot[-which(to_plot$alpha == 0 & to_plot$criteria == "TP"), ]
to_plot2 <- to_plot2[-which(to_plot2$alpha == 0 & to_plot2$criteria == "FP"), ]
delta_1 <- to_plot2 %>%
  filter(type_shift == "Delta[1](t)") %>%
   # filter(alpha != 0 & criteria == "TP") %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot2 %>% 
              filter(type_shift == "Delta[1](t)") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")), 
            aes(group = method)) +
  geom_point(size = 0.8, pch = 15) +
#  geom_hline(aes(yintercept = yintercept, 
#                 linetype = criteria), data = threshold,
#             linetype = 2) +
    geom_hline(data = to_plot2 %>% filter(criteria == "power"), 
               aes(yintercept = 0.05, 
                 linetype = "0.05")) +
  ggh4x::facet_grid2(rows=vars(criteria),
                     cols=vars(shape),
                     labeller=label_parsed,
                     scales = "free", 
                     axes = "margins")  +
  theme_bw() +
  xlab(TeX("$\\alpha$")) +
  ylab("") +
  ggtitle(TeX("$\\Delta_1$")) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none") +
  scale_linetype_manual(name = "threshold", values = 2) +
  theme(strip.background = element_rect(colour = "black", 
        fill = alpha("#EE9E94", 0.1)))
#ggsave("figures/simu_10_delta_1.pdf", width = 8, height = 6.5)
#R.utils::compressPDF("figures/simu_8_delta_23.pdf")
```


### Delta 2 

```{r}
threshold <- data.frame(
  yintercept = c(0.05, NA, NA), 
  criteria = factor(c("power", "TP", "FP"), levels = c("power", "TP", "FP")))
to_plot$criteria <- factor(to_plot$criteria, levels = c("power", "TP", "FP"))
to_plot2 <- to_plot[-which(to_plot$alpha == 0 & to_plot$criteria == "TP"), ]
to_plot2 <- to_plot2[-which(to_plot2$alpha == 0 & to_plot2$criteria == "FP"), ]
delta_2 <- to_plot2 %>%
  filter(type_shift == "Delta[2](t)") %>%
   # filter(alpha != 0 & criteria == "TP") %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot2 %>% 
              filter(type_shift == "Delta[2](t)") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")), 
            aes(group = method)) +
  geom_point(size = 0.8, pch = 15) +
#  geom_hline(aes(yintercept = yintercept, 
#                 linetype = criteria), data = threshold,
#             linetype = 2) +
    geom_hline(data = to_plot2 %>% filter(criteria == "power"), 
               aes(yintercept = 0.05, 
                 linetype = "0.05")) +
  ggh4x::facet_grid2(rows=vars(criteria),
                     cols=vars(shape),
                     labeller=label_parsed,
                     scales = "free", 
                     axes = "margins")  +
  theme_bw() +
  xlab(TeX("$\\alpha$")) +
  ylab("") +
  ggtitle(TeX("$\\Delta_2$")) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_linetype_manual(name = "threshold", values = 2) +
  theme(strip.background = element_rect(colour = "black", 
        fill = alpha("#EE9E94", 0.1)))
#ggsave("figures/simu_10_delta_2.pdf", width = 8, height = 6.5)
#R.utils::compressPDF("figures/simu_10_delta_2.pdf")
```


```{r}
cowplot::plot_grid(delta_1, delta_2, nrow = 2, rel_heights = c(1,1.1))
ggsave("figures/simu_10_delta_12.pdf", width = 8, height = 11)
#R.utils::compressPDF("figures/simu_8_delta_23.pdf")
```

### Delta 3

```{r}
threshold <- data.frame(
  yintercept = c(0.05, NA, NA), 
  criteria = factor(c("power", "TP", "FP"), levels = c("power", "TP", "FP")))
to_plot$criteria <- factor(to_plot$criteria, levels = c("power", "TP", "FP"))
to_plot2 <- to_plot[-which(to_plot$alpha == 0 & to_plot$criteria == "TP"), ]
to_plot2 <- to_plot2[-which(to_plot2$alpha == 0 & to_plot2$criteria == "FP"), ]
to_plot2 %>%
  filter(type_shift == "Delta[3](t)") %>%
   # filter(alpha != 0 & criteria == "TP") %>%
   filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")) %>%
  ggplot(aes(x = alpha, y = value, color = method)) +
  geom_line(data = to_plot2 %>% 
              filter(type_shift == "Delta[3](t)") %>%
              filter(method %in% c("DFFSS", "PFSS", "NPFSS", "HFSS")), 
            aes(group = method)) +
  geom_point(size = 0.8, pch = 15) +
#  geom_hline(aes(yintercept = yintercept, 
#                 linetype = criteria), data = threshold,
#             linetype = 2) +
    geom_hline(data = to_plot2 %>% filter(criteria == "power"), 
               aes(yintercept = 0.05, 
                 linetype = "0.05")) +
  ggh4x::facet_grid2(rows=vars(criteria),
                     cols=vars(shape),
                     labeller=label_parsed,
                     scales = "free", 
                     axes = "margins")  +
  theme_bw() +
  xlab(TeX("$\\alpha$")) +
  ylab("") +
  ggtitle(TeX("$\\Delta_3$")) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_linetype_manual(name = "threshold", values = 2) +
  theme(strip.background = element_rect(colour = "black", 
        fill = alpha("#EE9E94", 0.1)))
#ggsave("figures/simu_10_delta_3.pdf", width = 8, height = 6.5)
#R.utils::compressPDF("figures/simu_10_delta_3.pdf")
```


# Empirical part

## Spanish region 

The [spain_unemp.RData](spain_unemp.RData) file contains three objects:

* `Matcoordalpha`, the coordinates of the centroid of the Spanish regions (expressed in Cartesian coordinates),
* `MatX`,  the unemployment trends in Spanish regions over 80 quarters from 2002 to 2022,
* `region_spain`, the spatial contours of the Spanish regions.

```{r}
load("data/spain_unemp.RData")
dates <- seq(2002, 2021.75, by = 0.25)
y_lim <- range(MatX)
# import the OSM map
nc_osm <- get_tiles(region_spain, 
                      provider = "Esri.WorldShadedRelief", 
                      zoom = 7, crop = T)
# compute the distance between points
dist_proj <- as(dist(Matcoordalpha), "matrix")
# cartography
my_proj <- st_crs(region_spain)
spain <- st_union(region_spain)
```

First, we plot the data:

```{r, fig.width = 10, height = 4}
#pdf(file = "figures/spain_data.pdf", width = 10, height = 4.5)
par(mfrow = c(1, 2), mar = c(3.7, 3, 1, 1), oma = c(0, 0, 0, 0),
    las = 1, mgp = c(2.15, 0.75, 0))
# map
plot_tiles(nc_osm)
mf_shadow(spain, add = T, cex = 0.8)
plot(st_geometry(spain), border = rgb(0.5, 0.5, 0.5), lwd = 0.4, add = T,
     col = rgb(0.82, 0.82, 0.82))
plot(st_geometry(region_spain), border = rgb(1, 1, 1), lwd = 0.4, add = T,
     col = rgb(0.82, 0.82, 0.82))
# data
plot(dates, MatX[, 1], ylim = y_lim, xlab = 'Quarters of the period 2002-2022',
       ylab = 'Unemployment rate', xaxt = "n", 
     col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
     type = "l")
# abline(v = seq(2002, 2022, by = 4), lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
abline(h = seq(0, 40, by = 10), lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))

axis(1, at = seq(2002, 2022, by = 1),
     labels = F)
text(x = seq(2002, 2022, by = 1), y = par()$usr[3] - 0.03 * (par()$usr[4] - par()$usr[3]),
     labels = paste0(seq(2002, 2022, by = 1), "QI"),
     srt = 45, adj = 1, xpd = T, cex = 0.8)
for(j in 2:47)
  lines(dates, MatX[,j], ylim = y_lim, lwd = 1.3, 
        col = rgb(0.4, 0.4, 0.4, alpha = 0.3)) 
#dev.off()
#R.utils::compressPDF("figures/spain_data.pdf")
qpdf::pdf_compress("compressedPDFs/spain_data.pdf")
```

#### Descriptive Analysis

We represent the variable "Unemployment" aggregated over different 2-year periods (i.e., eight quarters).

```{r, fig.width = 12, fig.height= 4.5}
nb_split <- 10
step_years <- split(1:80, 
           sort(rep_len(1:nb_split, length.out = length(dates))))
#pdf(paste0("figures/Spain_evol.pdf"), width = 10, height = 7)
par(mfrow = c(3, 4), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, colMeans(MatX[step_years[[j]], ]))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 7, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlOrRd"
    my_pal <- rev(alpha(colorspace::sequential_hcl(7, palette = nom_pal), 1))

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      my_mean <- colMeans(MatX[step_years[[j]], ])
      my_col <- alpha(my_pal[findInterval(my_mean, my_interval, all.inside = T)],
                  1)
    
      plot_tiles(nc_osm)
      mf_shadow(spain, add = T, cex = 0.8)
      plot(st_geometry(region_spain), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- dates[chosen_years_5]
      title(paste0(my_years[1], "-", round(my_years[length(my_years)])), line = -.75)
     plot(st_geometry(region_spain), border = rgb(0.9, 0.9, 0.9), 
          lwd = 0.00000001, add = T)
     plot(st_geometry(spain), border = rgb(0.5, 0.5, 0.5), lwd = 0.4, add = T)
     if(j == nb_split)
       maplegend::leg(type = "choro", val = my_interval, pos = "bottomright", 
                 pal = my_pal, val_rnd = 3, title = "Unemp")
    }
#dev.off()
#R.utils::compressPDF("figures/Spain_evol.pdf")
```

Average over all the years:

```{r, fig.width = 12, fig.height= 4.5}
nb_split <- 1
step_years <- split(1:80, 
           sort(rep_len(1:nb_split, length.out = length(dates))))
#pdf(paste0("figures/Spain_mean.pdf"), width = 8, height = 7)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, colMeans(MatX[step_years[[j]], ]))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 7, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlOrRd"
    my_pal <- rev(alpha(colorspace::sequential_hcl(7, palette = nom_pal), 1))

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      my_mean <- colMeans(MatX[step_years[[j]], ])
      my_col <- alpha(my_pal[findInterval(my_mean, my_interval, all.inside = T)],
                  1)
    
      plot_tiles(nc_osm)
      mf_shadow(spain, add = T, cex = 0.8)
      plot(st_geometry(region_spain), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- dates[chosen_years_5]
      # title(paste0(my_years[1], "-", round(my_years[length(my_years)])), line = -.75)
     plot(st_geometry(region_spain), border = rgb(0.9, 0.9, 0.9), 
          lwd = 0.00000001, add = T)
     plot(st_geometry(spain), border = rgb(0.5, 0.5, 0.5), lwd = 0.4, add = T)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "bottomright", 
                 pal = my_pal, val_rnd = 3, title = "Unemp")
    }
#dev.off()
# R.utils::compressPDF("figures/Spain_mean.pdf")
```

We calculate all possible clusters:

```{r}
my_pairs_sp <- find_all_cluster(Matcoordalpha)
```

We use the four methods to detect clusters.

### NPFSS method

**Most Likely Cluster (MLC)**

```{r}
res_np <- compute_np(my_pairs_sp[[1]], my_pairs_sp[[2]], MatX)
res_np
```

**Significance**

```{r, eval = F}
p_value_np <- 0
B <- 999
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_np(my_pairs_sp[[1]], my_pairs_sp[[2]], MatXsim)
  p_value_np <- p_value_np + (res_np$stat < temp$stat)
}
cat("p-value: ", (1 + p_value_np) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

We remove all observations belonging to the most likely cluster from the list of potential clusters. For example, if the most likely cluster contains observations 1, 3, and 5, then the potential cluster with observations 8, 1, 3, 5, and 10 is replaced by 8 and 10. We then compute the NPFSS method on the new combinations.

```{r}
cluster_g1_temp <- sapply(my_pairs_sp[[1]], function(x) setdiff(x, res_np$vec))
cluster_g2_temp <- sapply(my_pairs_sp[[2]], function(x) setdiff(x, res_np$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_np_2 <- compute_np(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], MatX)
res_np_2
```

**Significance**

```{r, eval = F}
p_value_np_2 <- 0
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_np(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], MatXsim)
  p_value_np_2 <- p_value_np_2 + (res_np_2$stat < temp$stat)
}
cat("p-value: ", (1 + p_value_np_2) / (1 + B))
```

```{r, echo = F}
cat("p-value:  0.023")
```


### PFSS method

**Most Likely Cluster (MLC)**

```{r}
res_p <- compute_p(my_pairs_sp[[1]], my_pairs_sp[[2]], MatX)
res_p
```

**Significance**

```{r, eval = F}
p_value_p <- 0
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_np(my_pairs_sp[[1]], my_pairs_sp[[2]], MatXsim)
  p_value_p <- p_value_p + (res_p$stat < temp$stat)
}
cat("p-value: ", (1 + p_value_p) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```

**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(my_pairs_sp[[1]], function(x) setdiff(x, res_p$vec))
cluster_g2_temp <- sapply(my_pairs_sp[[2]], function(x) setdiff(x, res_p$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```

```{r}
res_p_2 <- compute_p(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], MatX)
res_p_2
```

**Significance of the secondary cluster**

```{r, eval = F}
p_value_p_2 <- 0
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_p(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], MatXsim)
  p_value_p_2 <- p_value_p_2 + (res_p_2$stat < temp$stat)
}
cat("p-value: ", (1 + p_value_p_2) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.015")
```



### DFFSS method

**Most Likely Cluster (MLC)**

```{r}
res_dffss <- compute_dffss(my_pairs_sp[[1]], my_pairs_sp[[2]], MatX)
res_dffss
```

**Significance**

```{r, eval = F}
p_value_dffss <- 0
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_dffss(my_pairs_sp[[1]], my_pairs_sp[[2]], MatX)
  p_value_dffss <- p_value_dffss + (res_dffss$stat < temp$stat)
}
(1 + p_value_dffss) /  (1 + B)
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(my_pairs_sp[[1]], function(x) setdiff(x, res_dffss$vec))
cluster_g2_temp <- sapply(my_pairs_sp[[2]], function(x) setdiff(x, res_dffss$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```

```{r}
res_dffss_2 <- compute_dffss(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], MatX)
res_dffss_2
```

**Significance of the 2MLC**

```{r, eval = F}
p_value_dffss_2 <- 0
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_dffss(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], MatXsim)
  p_value_dffss_2 <- p_value_dffss_2 + (res_dffss_2$stat < temp$stat)
}
cat("p-value: ", (1 + p_value_dffss_2) /  (1 + B))
```

```{r, echo = F}
cat("p-value:  0.004")
```



### HFSS method

**Most Likely Cluster (MLC)**

We first determine the optimal $K$:

```{r, fig.width = 6, fig.height = 3}
#pdf("figures/spain_h_CPV.pdf", width = 6, height = 4)
temp <- compute_h(my_pairs_sp[[1]], my_pairs_sp[[2]], MatX, 
                           nrow(MatX), plot_eigen = T)
#dev.off()
# R.utils::compressPDF("figures/spain_h_CPV.pdf")
```

We select $K=2$, which explains approximately $90\%$ of the variance. Note that using $K=10$, which accounts for $95\%$ of the variance, leads to the same identified clusters.

```{r}
res_h <- compute_h(my_pairs_sp[[1]], my_pairs_sp[[2]], MatX, 2)
res_h
```

```{r, eval = F}
p_value_h <- 0
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_h(my_pairs_sp[[1]], my_pairs_sp[[2]], MatXsim, 2)
  p_value_h <- p_value_h + (res_h$stat < temp$stat)
}
(1 + p_value_h) /  (1 + B)
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(my_pairs_sp[[1]], function(x) setdiff(x, res_h$vec))
cluster_g2_temp <- sapply(my_pairs_sp[[2]], function(x) setdiff(x, res_h$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


We aim to find the optimal value of $K$:

```{r, fig.width = 6, fig.height = 3}
#pdf("figures/spain_h_CPV_2.pdf", width = 6, height = 4)
temp <- compute_h(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], MatX, 
                           nrow(MatX), plot_eigen = T)
#dev.off()
# R.utils::compressPDF("figures/spain_h_CPV_2.pdf")
```

We choose $K=2$.

```{r}
res_h_2 <- compute_h(cluster_g1_temp[-id_pos], 
                     cluster_g2_temp[-id_pos], MatX, 2)
res_h_2
```

**Significance of the 2MLC**

```{r, eval = F}
p_value_h_2 <- 0
pb <- progress_bar$new(total = B)

for(b in 1:B) {
  pb$tick()
  perm <- sample(ncol(MatX))
  MatXsim <- MatX[, perm]
  temp <- compute_h(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], 
                    MatXsim, 2)
  p_value_h_2 <- p_value_h_2 + (res_h_2$stat < temp$stat)
}
cat("p-value: ", (1 + p_value_h_2) /  (1 + B))
```



```{r, echo = F}
cat("p-value:  0.003")
```


### Summary of the results

**Visualization of the result**

```{r}
res <- vector("list", 4)
res[[1]][[1]] <- res_h
res[[1]][[2]] <- res_h_2
res[[2]][[1]] <- res_dffss
res[[2]][[2]] <- res_dffss_2
res[[3]][[1]] <- res_np
res[[3]][[2]] <- res_np_2
res[[4]][[1]] <- res_p
res[[4]][[2]] <- res_p_2
names_method <- c("HFSS", "DFFSS", "NPFSS", "PFSS")
cols = c("#D35C79", "#009593")
```


```{r, fig.width = 12, fig.height = 6, message = F, warning = F}
my_var <- 'Unemployment rate (in %)'
my_country <- "ESP" 

y_lim <- range(MatX)

for(k in 1:4) {
  my_cluster_1 <- res[[k]][[1]]$vec
  my_cluster_2 <- res[[k]][[2]]$vec

#pdf(file = paste0("figures/", my_country, "_", names_method[k], ".pdf"), width = 11.5, height = 3.9) 
sf_use_s2(F)
nf <- layout( matrix(c(1,1,2,3), nrow=2, byrow=F) )
  par(mar = c(1.5, 0, 0, 0.2), 
      oma = c(0.5, 0, 2.4, 0), mgp = c(2.4, 0.6, 0), las = 1)
  ##### Map #########
    # map
  col_geo <- rep(rgb(0.9, 0.9, 0.9, alpha = 0.1), nrow(Matcoordalpha))
  cex_geo <- rep(0.7, nrow(Matcoordalpha))
  col_geo[my_cluster_1] <- alpha(cols[1], 0.8)
  cex_geo[my_cluster_1] <- 1

  col_geo[my_cluster_2] <- alpha(cols[2], 0.5)
  cex_geo[my_cluster_2] <- 1
  
  
   plot_tiles(nc_osm)
  mf_shadow(spain, add = T, cex = 0.8)
  mf_shadow(st_union(region_spain[my_cluster_1, ]), 
                add = T, cex = 0.8, col = cols[1])
  mf_shadow(st_union(region_spain[my_cluster_2, ]), 
         add = T, col= cols[2], cex = 0.8)
      
  plot(st_geometry(spain), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

  
  plot(st_geometry(region_spain), 
        border = "white",
        col = col_geo, 
        cex = cex_geo,
        pch = 16, asp = 1, add = T, lwd = 0.1)

  plot(st_geometry(st_union(region_spain[my_cluster_1, ])), 
         add = T, border= cols[1], col = NULL)
  plot(st_geometry(st_union(region_spain[my_cluster_2, ])), 
         add = T, border= cols[2], col = NULL)

      
    temp_1 <- draw.circle(Matcoordalpha[my_cluster_1[1], 1], 
                           Matcoordalpha[my_cluster_1[1], 2], 
                  as.numeric(dist_proj[my_cluster_1[1], 
                                       my_cluster_1[length(my_cluster_1)]]))

  temp_2 <- draw.circle(Matcoordalpha[my_cluster_2[1], 1], 
                           Matcoordalpha[my_cluster_2[1], 2], 
                  as.numeric(dist_proj[my_cluster_2[1], 
                                       my_cluster_2[length(my_cluster_2)]]))
    ###############
  polygon(temp_1$x, temp_1$y, border= cols[1],
             col = alpha(cols[1], 0.4), lty=1, lwd=1)

  polygon(temp_2$x, temp_2$y, border= cols[2],
             col = alpha(cols[2], 0.4), lty=1, lwd=1)
  
    mtext(my_var,
       side = 4, line = -2.8, las = 0)
    
  legend("topleft", legend = c("Most likely cluster", "Secondary cluster"),
         fill = c(cols[1], cols[2]), cex = 0.9, box.lty = 0)
      
  ##### Functional data
  plot(dates, MatX[, 1], ylim = y_lim, xlab = '',
       ylab = '', col = rgb(0.6, 0.6, 0.6, alpha = 0.5), xaxt = 'n', 
        type = "l")
  legend("topleft", legend = c("Most likely cluster"),
         lty = 1, col = c(cols[1]), cex = 0.9)
  abline(v = seq(2002, 2022, by = 4), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 40, by = 10),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))

  for (j in 2:ncol(MatX))
        lines(dates, MatX[, j], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
    
  for(i in my_cluster_1)
        lines(dates, MatX[, i], col = alpha(cols[1], 0.3),
          lty = 1, lwd = 1.3)
  
  lines(dates, rowMeans(MatX), lwd = 1.3, lty = 2)
  

plot(dates, MatX[, 1], ylim = y_lim, xlab = 'Years',
       ylab = '', xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  legend("topleft", legend = c("Secondary cluster"),
         lty = 1, col = c(cols[2]), cex = 0.9)
  abline(v = seq(2002, 2022, by = 4), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 40, by = 10),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(2002, 2022, by = 4), xlab = "years",
           labels = as.character(seq(2002, 2022, by = 4)))
  for (j in 2:ncol(MatX))
        lines(dates, MatX[, j], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
      
  for(i in my_cluster_2)
        lines(dates, MatX[, i], col = alpha(cols[2], 0.3),
          lty = 1, lwd = 1.3)
  
    
  lines(dates, rowMeans(MatX), lwd = 1.3, lty = 2)
  
  mtext(paste0("Clusters for the ", names_method[k]), side = 3, line = 0.8, outer = TRUE)
#dev.off()
#R.utils::compressPDF(paste0("figures/", my_country, "_", names_method[k], ".pdf"))
}
```

The following table presents the results obtained from the different methods.

```{r}
res_spain <- data.frame(nb_cluster_1 = c(length(res_np$vec), 
                            length(res_p$vec),
                            length(res_dffss$vec),
                            length(res_h$vec)),
           sign_cluster_1 = c(0.001, 0.001, 0.001, 0.001),
           nb_cluster_2 = c(length(res_np_2$vec), 
                            length(res_p_2$vec),
                            length(res_dffss_2$vec),
                            length(res_h_2$vec)),
           sign_cluster_2 = c(0.023, 0.015, 0.004, 0.003))
row.names(res_spain) <- c("NPFSS", "PFSS", "DFFSS", "HFSS")
knitr::kable(res_spain)
```


## Climate Data

### Preparation of the data

Import the data:
```{r, message = F}
load("data/my_grid.RData")
coord_region <- st_as_sf(my_grid, coords = c("long", "lat"),
                         crs = 4326)
countries_regions <- st_read("data/world-administrative-boundaries.geojson")
```

Preparation of the data:
```{r}
# transforn the points into grid
long_lat_temp <- my_grid[, c("long", "lat")] %>%
  filter(lat > -90, lat < 90, long > -180, long < 180) %>%
  arrange(lat, long)   %>%
  dplyr::select(long, lat)
df_sf_temp <- st_as_sf(long_lat_temp, coords = c("long", "lat"))
st_crs(df_sf_temp) <- 4326
all_cells <- df_sf_temp %>% 
  st_make_grid(cellsize = c(0.625, 0.5), 
               offset = c(-179.375 - 0.3125, -89.5 -0.25)) %>% # c(-180 - 0.3125, -90 - 0.25))
  st_as_sf() %>% 
  st_join(df_sf_temp) 
all_cells$long <- long_lat_temp$long
all_cells$lat <- long_lat_temp$lat
# Initialisation
unique_year <- 1981:2023
chosen_years <- 20:43
```

### Great-Britain

We select the ISO3 code corresponding to Great-Britain. 

```{r}
my_country <- "GBR"
my_proj <- 3035 
```

```{r, warning = F, message = F}
select_countries <- countries_regions[countries_regions$color_code %in% my_country, ]
# drop the islands
sf_obj <- select_countries %>%
    filter(iso3 == my_country[1]) %>%
    mutate(area = st_area(geometry)) %>%
    top_n(1, area) %>%
    rowid_to_column() %>%
    st_cast("POLYGON")  %>% 
    mutate(area = st_area(geometry)) %>%
    group_by(rowid) %>%
    top_n(1, area)
```


```{r, warning = F, message = F}
if(length(my_country) > 1) {
  for(i in 2:length(my_country)) {
    sf_obj <- rbind(sf_obj, select_countries %>%
      filter(iso3 == my_country[i]) %>%
      mutate(area = st_area(geometry)) %>%
      top_n(1, area) %>%
      rowid_to_column() %>%
      st_cast("POLYGON")  %>% 
      mutate(area = st_area(geometry)) %>%
      group_by(rowid) %>%
      top_n(1, area)
    )
  }
  }
is_intersect <- st_intersects(all_cells, sf_obj)
is_intersect <- which(sapply(is_intersect, function(x) length(x) != 0))
lldata_poly <- all_cells[is_intersect, ]
  
my_contours <- st_intersection(countries_regions, 
                               st_union(lldata_poly, is_coverage = T))

# dowload tiles and compose raster (SpatRaster)
nc_osm <- get_tiles(my_contours, 
                      provider = "Esri.WorldShadedRelief", 
                      zoom = 7, crop = F)
poly_cell <- merge(lldata_poly, my_grid, by = c("long", "lat"))
# coordinates of the centroid
coord_temp <- st_transform(poly_cell, my_proj)
# simplify the geometry
poly_cell <- st_intersection(poly_cell, my_contours)
poly_cell <- poly_cell[!duplicated(cbind(poly_cell$long, poly_cell$lat)), ]
```

We compute all possible clusters:

```{r, warning = F}
  coord_proj <- st_coordinates(st_centroid(coord_temp))
  pairs_geo <- find_all_cluster(coord_proj)
  dist_proj <- as(dist(cbind(coord_proj[, 1], coord_proj[, 2])), "matrix")

  dist_4326 <- as(dist(cbind(poly_cell$long, poly_cell$lat)), "matrix")
  coord_4326 <- st_transform(poly_cell, 4326) 
```

```{r}
  my_var <- "t2m_diff_" 
  temp_var <- poly_cell[, paste0(my_var, unique_year)[chosen_years]]
  MatX <- as.matrix(st_drop_geometry(temp_var))
```

#### Descriptive Analysis

First, we plot the data:

```{r, fig.width = 10, height = 4}
y_lim <- range(MatX)
dates <- unique_year[chosen_years]
#pdf(file = "figures/GBR_data.pdf", width = 10, height = 4.5)
par(mfrow = c(1, 2), mar = c(3.7, 3, 1, 1), oma = c(0, 0, 0, 0),
    las = 1, mgp = c(2.2, 0.8, 0))
# map
plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

# data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = 'Difference from average temperatures (in C)', xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(-4, 4, by = 0.5),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
#dev.off()
#  R.utils::compressPDF("figures/GBR_data.pdf")
```

We map the variable "Difference from average temperatures" aggregated over different four-year periods. 

```{r, fig.width = 12, fig.height= 3.5}
nb_split <- 8
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/GB_evol.pdf"), width = 12, height = 8)
par(mfrow = c(2, 4), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 8, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlOrRd"
    my_pal <- rev(alpha(colorspace::sequential_hcl(8, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
      title(paste0(my_years[1], "-", my_years[length(my_years)]), line = -1.25)
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Diff Temp")
    }
#dev.off()
 #    R.utils::compressPDF("figures/GB_evol.pdf")
```

Average mean of difference temperatures:

```{r, fig.width = 12, fig.height= 3.5}
nb_split <- 1
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/GBR_mean.pdf"), width = 7, height = 8)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 8, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlOrRd"
    my_pal <- rev(alpha(colorspace::sequential_hcl(8, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Diff Temp")
    }
#dev.off()
#     R.utils::compressPDF("figures/GBR_mean.pdf")
```

#### NPFSS method

**Most Likely Cluster (MLC)**

```{r}
res_np <- compute_np(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_np
```

**Significance (using parallel computing)**

```{r, eval = F}
B <- 999
compute_fun_par <- function(b, fun, ...) {
  set.seed(b)
  perm <- sample(nrow(MatX))
  MatXsim <- MatX[perm, ]
  temp <- fun(c1, c2, t(MatXsim))
  temp$stat
}

c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

require(parallel)
cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_np$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_np$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```

```{r}
res_np_2 <- compute_np(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_np_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np_2$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```



#### PFSS method

**Most Likely Cluster (MLC)**

```{r}
res_p <- compute_p(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_p
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p$stat <x))) /  (1 + B))
```



```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_p$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_p$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_p_2 <- compute_p(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_p_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p_2$stat <x))) /  (1 + B))
```

```{r, echo = F}
cat("p-value:  0.001")
```



#### DFFSS method

**Most Likely Cluster (MLC)**

```{r}
res_dffss <- compute_dffss(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_dffss
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_p$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_dffss$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_dffss$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_dffss_2 <- compute_dffss(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_dffss_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_dffss_2$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```



#### HFSS method

**Most Likely Cluster (MLC)**

First, we determine the value of $K$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV.pdf"), width = 6, height = 4)
temp <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), 
                           ncol(MatX), plot_eigen = T)
#dev.off()
```

We select $K=6$:

```{r}
res_h <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), 6)
res_h
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_h$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_h$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```

We seek the optimal value of $K$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV_2.pdf"), width = 6, height = 4)
temp <- compute_h(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], t(MatX), 
                           ncol(MatX), plot_eigen = T)
#dev.off()
```

We choose $K=6$.

```{r}
res_h_2 <- compute_h(cluster_g1_temp[-id_pos], 
                     cluster_g2_temp[-id_pos], t(MatX), 6)
res_h_2
```


**Significance of 2MLC**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h_2$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```



#### Summary of the results

**Visualization of the result**

```{r}
res <- vector("list", 4)
res[[1]][[1]] <- res_h
res[[1]][[2]] <- res_h_2
res[[2]][[1]] <- res_dffss
res[[2]][[2]] <- res_dffss_2
res[[3]][[1]] <- res_np
res[[3]][[2]] <- res_np_2
res[[4]][[1]] <- res_p
res[[4]][[2]] <- res_p_2
```

```{r, fig.width = 12, fig.height = 6, message = F, warning = F}
my_var <- 'Difference with the normal temperature (in C)'
dates <- unique_year[chosen_years]
y_lim <- range(MatX)

for(k in 1:4) {
  my_cluster_1 <- res[[k]][[1]]$vec
  my_cluster_2 <- res[[k]][[2]]$vec

#pdf(file = paste0("figures/", my_country, "_", names_method[k], ".pdf"), width = 11.5, height = 3.8) 
sf_use_s2(F)
nf <- layout( matrix(c(1,1,2,3), nrow=2, byrow=F) )
  par(mar = c(1.5, 0, 0, 0.2), 
      oma = c(0.5, 0, 2.4, 0), mgp = c(2.4, 0.6, 0), las = 1)
  ##### Map #########
    # map
  col_geo <- rep(rgb(0.9, 0.9, 0.9, alpha = 0.1), nrow(poly_cell))
  cex_geo <- rep(0.7, nrow(poly_cell))
  col_geo[my_cluster_1] <- alpha(cols[1], 0.8)
  cex_geo[my_cluster_1] <- 1

  col_geo[my_cluster_2] <- alpha(cols[2], 0.5)
  cex_geo[my_cluster_2] <- 1
  
  plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  mf_shadow(st_union(poly_cell[my_cluster_1, ]), 
                add = T, cex = 0.8, col = cols[1])
  mf_shadow(st_union(poly_cell[my_cluster_2, ]), 
         add = T, col= cols[2], cex = 0.8)
      
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

  
  plot(st_geometry(poly_cell), 
        border = "white",
        col = col_geo, 
        cex = cex_geo,
        pch = 16, asp = 1, add = T, lwd = 0.1)

  plot(st_geometry(st_union(poly_cell[my_cluster_1, ])), 
         add = T, border= cols[1], col = NULL)
  plot(st_geometry(st_union(poly_cell[my_cluster_2, ])), 
         add = T, border= cols[2], col = NULL)
      
  temp_1 <- draw.circle(coord_proj[my_cluster_1[1], 1], 
                           coord_proj[my_cluster_1[1], 2], 
                  as.numeric(dist_proj[my_cluster_1[1], 
                                       my_cluster_1[length(my_cluster_1)]]))

    my_circle_1 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_1$x, temp_1$x[1]), 
             c(temp_1$y, temp_1$y[1]))
         )), crs = my_proj
       ), 4326)
    
  temp_2 <- draw.circle(coord_proj[my_cluster_2[1], 1], 
                           coord_proj[my_cluster_2[1], 2], 
                  as.numeric(dist_proj[my_cluster_2[1], 
                                       my_cluster_2[length(my_cluster_2)]]))
    
  my_circle_2 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_2$x, temp_2$x[1]), 
             c(temp_2$y, temp_2$y[1]))
         )), crs = my_proj
       ), 4326)
  
  
  ###############
  plot(st_geometry(my_circle_2), add = T, border= cols[2],
             col = alpha(cols[2], 0.4), lty=1, lwd=1)
  plot(st_geometry(my_circle_1), add = T, border= cols[1],
             col = alpha(cols[1], 0.4), lty=1, lwd=1)
  
    mtext(my_var,
       side = 4, line = -3.5, las = 0)
    
  legend("topleft", legend = c("Most likely cluster", "Secondary cluster"),
         fill = c(cols[1], cols[2]), cex = 0.9, box.lty = 0)
      
  ##### Functional data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = '',
       ylab = '', col = rgb(0.6, 0.6, 0.6, alpha = 0.5), xaxt = 'n', 
        type = "l")
  legend("topleft", legend = c("Most likely cluster"),
         lty = 1, col = c(cols[1]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(-4, 4, by = 0.5),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))

  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
    
  for(i in my_cluster_1)
        lines(dates, MatX[i, ], col = alpha(cols[1], 0.3),
          lty = 1, lwd = 1.3)
  
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  

plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = 'Difference from average temperatures (in C)', xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  legend("topleft", legend = c("Secondary cluster"),
         lty = 1, col = c(cols[2]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(-4, 4, by = 0.5),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
      
  for(i in my_cluster_2)
        lines(dates, MatX[i, ], col = alpha(cols[2], 0.3),
          lty = 1, lwd = 1.3)
  
    
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  
  mtext(paste0("Clusters for the ", names_method[k]), side = 3, line = 0.8, outer = TRUE)
#dev.off()
#  R.utils::compressPDF(paste0("figures/", my_country, "_", names_method[k], ".pdf"))
}
```

```{r}
res_GB <- data.frame(nb_cluster_1 = c(length(res_np$vec), 
                            length(res_p$vec),
                            length(res_dffss$vec),
                            length(res_h$vec)),
           sign_cluster_1 = c(0.001, 0.001, 0.001, 0.001),
           nb_cluster_2 = c(length(res_np_2$vec), 
                            length(res_p_2$vec),
                            length(res_dffss_2$vec),
                            length(res_h_2$vec)),
           sign_cluster_2 = c(0.001, 0.001, 0.001, 0.001))
row.names(res_GB) <- c("NPFSS", "PFSS", "DFFSS", "HFSS")
knitr::kable(res_GB)
```



### Nigeria

We select the ISO3 code of Nigeria:

```{r}
my_country <- "NGA"
my_proj <- 32629 
```

```{r, warning = F, message = F}
select_countries <- countries_regions[countries_regions$color_code %in% my_country, ]
  
# drop the islands
sf_obj <- select_countries %>%
    filter(iso3 == my_country[1]) %>%
    mutate(area = st_area(geometry)) %>%
    top_n(1, area) %>%
    rowid_to_column() %>%
    st_cast("POLYGON")  %>% 
    mutate(area = st_area(geometry)) %>%
    group_by(rowid) %>%
    top_n(1, area)
  
if(length(my_country) > 1) {
  for(i in 2:length(my_country)) {
    sf_obj <- rbind(sf_obj, select_countries %>%
      filter(iso3 == my_country[i]) %>%
      mutate(area = st_area(geometry)) %>%
      top_n(1, area) %>%
      rowid_to_column() %>%
      st_cast("POLYGON")  %>% 
      mutate(area = st_area(geometry)) %>%
      group_by(rowid) %>%
      top_n(1, area)
    )
  }
  }
is_intersect <- st_intersects(all_cells, sf_obj)
is_intersect <- which(sapply(is_intersect, function(x) length(x) != 0))
lldata_poly <- all_cells[is_intersect, ]
  
my_contours <- st_intersection(select_countries, 
                               st_union(lldata_poly, is_coverage = T))

# dowload tiles and compose raster (SpatRaster)
nc_osm <- get_tiles(my_contours, 
                      provider = "Esri.WorldShadedRelief", 
                      zoom = 7, crop = F)
poly_cell <- merge(lldata_poly, my_grid, by = c("long", "lat"))
# coordinates of the centroid
coord_temp <- st_transform(poly_cell, my_proj)
# simplify the geometry
poly_cell <- st_intersection(poly_cell, my_contours)
```

We compute all possible clusters:

```{r, warning = F}
  coord_proj <- st_coordinates(st_centroid(coord_temp))
  pairs_geo <- find_all_cluster(coord_proj)
  dist_proj <- as(dist(cbind(coord_proj[, 1], coord_proj[, 2])), "matrix")

  dist_4326 <- as(dist(cbind(poly_cell$long, poly_cell$lat)), "matrix")
  coord_4326 <- st_transform(poly_cell, 4326) 
```

```{r}
  my_var <- "prec_5days_" 
  temp_var <- poly_cell[, paste0(my_var, unique_year)[chosen_years]]
  MatX <- as.matrix(st_drop_geometry(temp_var))
```

#### Descriptive Analysis

First, we plot the data:

```{r, fig.width = 10, height = 4}
y_lim <- range(MatX)
title_var <- 'Maximum consecutive 5-days precipitation (in mm)'
#pdf(file = "figures/NGA_data.pdf", width = 12, height = 4.5)
par(mfrow = c(1, 2), mar = c(3.7, 3, 1, 1), oma = c(0, 0, 0, 0),
    las = 1, mgp = c(2.2, 0.5, 0))
# map
plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

# data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = title_var, xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
abline(h = seq(0, 2500, by = 500),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
#dev.off()
#  R.utils::compressPDF("figures/NGA_data.pdf")
```

We map the average of the variable over a 3-year window. 

```{r, fig.width = 12, fig.height= 6}
nb_split <- 8
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/NGA_evol.pdf"), width = 12, height = 5)
par(mfrow = c(2, 4), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 8, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlGnBu"
    my_pal <- rev(alpha(colorspace::sequential_hcl(8, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
      title(paste0(my_years[1], "-", my_years[length(my_years)]), line = -1.25)
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Prec 5-days")
    }
#dev.off()
#      R.utils::compressPDF("figures/NGA_evol.pdf")
```

Average mean:

```{r, fig.width = 12, fig.height= 6}
nb_split <- 1
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/NGA_mean.pdf"), width = 7, height = 5)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 8, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlGnBu"
    my_pal <- rev(alpha(colorspace::sequential_hcl(8, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
      #title(paste0(my_years[1], "-", my_years[length(my_years)]), line = -1.25)
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Prec 5-days")
    }
#dev.off()
 #   R.utils::compressPDF("figures/NGA_mean.pdf")
```

#### NPFSS method

**Most Likely Cluster (MLC)**

```{r}
res_np <- compute_np(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_np
```

**Significance (using parallel computing)**

```{r, eval = F}
B <- 999

c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_np$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_np$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_np_2 <- compute_np(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_np_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np_2$stat < x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```




#### PFSS method

**Most Likely Cluster (MLC)**

```{r}
res_p <- compute_p(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_p
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p$stat <x))) /  (1 + B))
```



```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_p$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_p$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_p_2 <- compute_p(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_p_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p_2$stat <x))) /  (1 + B))
```

```{r, echo = F}
cat("p-value:  0.001")
```



#### DFFSS method

**Most Likely Cluster (MLC)**

```{r}
res_dffss <- compute_dffss(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_dffss
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_p$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_dffss$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_dffss$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```



```{r}
res_dffss_2 <- compute_dffss(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_dffss_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_dffss_2$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```





#### HFSS Method 

**Most Likely Cluster (MLC)**

First, we determine the value of $K$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV.pdf"), width = 6, height = 4)
temp <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), 
                           ncol(MatX), plot_eigen = T)
#dev.off()
```

We select $K=6$:
```{r}
res_h <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), 6)
res_h
```

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_h$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_h$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


We seek the optimal value of $K$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV_2.pdf"), width = 6, height = 4)
temp <- compute_h(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], t(MatX), 
                           ncol(MatX), plot_eigen = T)
#dev.off()
```

We choose $K=6$.

```{r}
res_h_2 <- compute_h(cluster_g1_temp[-id_pos], 
                     cluster_g2_temp[-id_pos], t(MatX), 6)
res_h_2
```


**Significance of the 2MLC**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h_2$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```



#### Summary of the results

**Visualization of the result**

```{r}
res <- vector("list", 4)
res[[1]][[1]] <- res_h
res[[1]][[2]] <- res_h_2
res[[2]][[1]] <- res_dffss
res[[2]][[2]] <- res_dffss_2
res[[3]][[1]] <- res_np
res[[3]][[2]] <- res_np_2
res[[4]][[1]] <- res_p
res[[4]][[2]] <- res_p_2
```

```{r, fig.width = 12, fig.height = 6, message = F, warning = F}
title_var <- 'Maximum consecutive 5-days precipitation (in mm)'
dates <- unique_year[chosen_years]
y_lim <- range(MatX)

for(k in 1:4) {
  my_cluster_1 <- res[[k]][[1]]$vec
  my_cluster_2 <- res[[k]][[2]]$vec

#pdf(file = paste0("figures/", my_country, "_", names_method[k], ".pdf"), width = 13, height = 4.) 

sf_use_s2(F)
nf <- layout( matrix(c(1,1,2,3), nrow=2, byrow=F) )
  par(mar = c(1.5, 0, 0, 0.2), 
      oma = c(0.5, 0, 2.4, 0), mgp = c(2.4, 0.6, 0), las = 1)
  ##### Map #########
    # map
  col_geo <- rep(rgb(0.9, 0.9, 0.9, alpha = 0.1), nrow(poly_cell))
  cex_geo <- rep(0.7, nrow(poly_cell))
  col_geo[my_cluster_1] <- alpha(cols[1], 0.8)
  cex_geo[my_cluster_1] <- 1

  col_geo[my_cluster_2] <- alpha(cols[2], 0.5)
  cex_geo[my_cluster_2] <- 1
  
  plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  mf_shadow(st_union(poly_cell[my_cluster_1, ]), 
                add = T, cex = 0.8, col = cols[1])
  mf_shadow(st_union(poly_cell[my_cluster_2, ]), 
         add = T, col= cols[2], cex = 0.8)
      
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

  
  plot(st_geometry(poly_cell), 
        border = "white",
        col = col_geo, 
        cex = cex_geo,
        pch = 16, asp = 1, add = T, lwd = 0.1)

  plot(st_geometry(st_union(poly_cell[my_cluster_1, ])), 
         add = T, border= cols[1], col = NULL)
  plot(st_geometry(st_union(poly_cell[my_cluster_2, ])), 
         add = T, border= cols[2], col = NULL)
      
  temp_1 <- draw.circle(coord_proj[my_cluster_1[1], 1], 
                           coord_proj[my_cluster_1[1], 2], 
                  as.numeric(dist_proj[my_cluster_1[1], 
                                       my_cluster_1[length(my_cluster_1)]]))

    my_circle_1 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_1$x, temp_1$x[1]), 
             c(temp_1$y, temp_1$y[1]))
         )), crs = my_proj
       ), 4326)
    
  temp_2 <- draw.circle(coord_proj[my_cluster_2[1], 1], 
                           coord_proj[my_cluster_2[1], 2], 
                  as.numeric(dist_proj[my_cluster_2[1], 
                                       my_cluster_2[length(my_cluster_2)]]))
    
  my_circle_2 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_2$x, temp_2$x[1]), 
             c(temp_2$y, temp_2$y[1]))
         )), crs = my_proj
       ), 4326)
  
  
  ###############
  plot(st_geometry(my_circle_2), add = T, border= cols[2],
             col = alpha(cols[2], 0.4), lty=1, lwd=1)
  plot(st_geometry(my_circle_1), add = T, border= cols[1],
             col = alpha(cols[1], 0.4), lty=1, lwd=1)
  
    mtext(my_var, side = 4, line = -3.5, las = 0, cex = 0.8)
    
  legend("topleft", legend = c("Most likely cluster", "Secondary cluster"),
         fill = c(cols[1], cols[2]), cex = 0.9, box.lty = 0)
      
  ##### Functional data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = '',
       ylab = '', col = rgb(0.6, 0.6, 0.6, alpha = 0.5), xaxt = 'n', 
        type = "l")
  legend("topleft", legend = c("Most likely cluster"),
         lty = 1, col = c(cols[1]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 2500, by = 500),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))

  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
    
  for(i in my_cluster_1)
        lines(dates, MatX[i, ], col = alpha(cols[1], 0.3),
          lty = 1, lwd = 1.3)
  
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  

plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = title_var, xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  legend("topleft", legend = c("Secondary cluster"),
         lty = 1, col = c(cols[2]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 2500, by = 500),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
      
  for(i in my_cluster_2)
        lines(dates, MatX[i, ], col = alpha(cols[2], 0.3),
          lty = 1, lwd = 1.3)
  
    
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  
  mtext(paste0("Clusters for the ", names_method[k]), side = 3, line = 0.8, outer = TRUE)
#dev.off()
#  R.utils::compressPDF(paste0("figures/", my_country, "_", names_method[k], ".pdf"))
}
```


The following table presents the results obtained from the various methods.

```{r}
res_NGA <- data.frame(nb_cluster_1 = c(length(res_np$vec), 
                            length(res_p$vec),
                            length(res_dffss$vec),
                            length(res_h$vec)),
           sign_cluster_1 = c(0.001, 0.001, 0.001, 0.001),
           nb_cluster_2 = c(length(res_np_2$vec), 
                            length(res_p_2$vec),
                            length(res_dffss_2$vec),
                            length(res_h_2$vec)),
           sign_cluster_2 = c(0.001, 0.001, 0.001, 0.001))
row.names(res_NGA) <- c("NPFSS", "PFSS", "DFFSS", "HFSS")
knitr::kable(res_NGA)
```



### Pakistan

We select the ISO3 code for Pakistan:

```{r}
my_country <- "PAK"
my_proj <- 9678 
```

```{r, warning = F, message = F}
select_countries <- countries_regions[countries_regions$color_code %in% my_country, ]
  
# drop the islands
sf_obj <- select_countries %>%
    filter(iso3 == my_country[1]) %>%
    mutate(area = st_area(geometry)) %>%
    top_n(1, area) %>%
    rowid_to_column() %>%
    st_cast("POLYGON")  %>% 
    mutate(area = st_area(geometry)) %>%
    group_by(rowid) %>%
    top_n(1, area)
  
if(length(my_country) > 1) {
  for(i in 2:length(my_country)) {
    sf_obj <- rbind(sf_obj, select_countries %>%
      filter(iso3 == my_country[i]) %>%
      mutate(area = st_area(geometry)) %>%
      top_n(1, area) %>%
      rowid_to_column() %>%
      st_cast("POLYGON")  %>% 
      mutate(area = st_area(geometry)) %>%
      group_by(rowid) %>%
      top_n(1, area)
    )
  }
  }
is_intersect <- st_intersects(all_cells, sf_obj)
is_intersect <- which(sapply(is_intersect, function(x) length(x) != 0))
lldata_poly <- all_cells[is_intersect, ]
  
my_contours <- st_intersection(select_countries, 
                               st_union(lldata_poly, is_coverage = T))

# dowload tiles and compose raster (SpatRaster)
nc_osm <- get_tiles(my_contours, 
                      provider = "Esri.WorldShadedRelief", 
                      zoom = 7, crop = F)
poly_cell <- merge(lldata_poly, my_grid, by = c("long", "lat"))
# coordinates of the centroid
coord_temp <- st_transform(poly_cell, my_proj)
# simplify the geometry
poly_cell <- st_intersection(poly_cell, my_contours)
```

We compute all possible clusters:

```{r, warning = F}
  coord_proj <- st_coordinates(st_centroid(coord_temp))
  pairs_geo <- find_all_cluster(coord_proj)
  dist_proj <- as(dist(cbind(coord_proj[, 1], coord_proj[, 2])), "matrix")

  dist_4326 <- as(dist(cbind(poly_cell$long, poly_cell$lat)), "matrix")
  coord_4326 <- st_transform(poly_cell, 4326) 
```

```{r}
  my_var <- "prec_5days_" 
  temp_var <- poly_cell[, paste0(my_var, unique_year)[chosen_years]]
  MatX <- as.matrix(st_drop_geometry(temp_var))
```

#### Descriptive Analysis

First, we plot the data:

```{r, fig.width = 10, height = 4}
y_lim <- range(MatX)
title_var <- 'Maximum consecutive 5-days precipitation (in mm)'
#pdf(file = "figures/PAK_data.pdf", width = 10, height = 4.5)
par(mfrow = c(1, 2), mar = c(3.7, 3, 1, 1), oma = c(0, 0, 0, 0),
    las = 1, mgp = c(2.2, 0.8, 0))
# map
plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

# data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = title_var, xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
abline(h = seq(0, 800, by = 200),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
#dev.off()
# R.utils::compressPDF(paste0("figures/PAK_data.pdf"))
```

We map the average of the variable over a three-year window

```{r, fig.width = 12, fig.height= 6}
nb_split <- 8
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/PAK_evol.pdf"), width = 12, height = 8)
par(mfrow = c(2, 4), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 8, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlGnBu"
    my_pal <- rev(alpha(colorspace::sequential_hcl(8, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
      title(paste0(my_years[1], "-", my_years[length(my_years)]), line = -1.25)
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Prec 5-days")
    }
#dev.off()
# R.utils::compressPDF(paste0("figures/PAK_evol.pdf"))
```

Average mean:

```{r, fig.width = 12, fig.height= 6}
nb_split <- 1
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/PAK_mean.pdf"), width = 8, height = 8)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 8, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlGnBu"
    my_pal <- rev(alpha(colorspace::sequential_hcl(8, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
      #title(paste0(my_years[1], "-", my_years[length(my_years)]), line = -1.25)
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Prec 5-days")
    }
#dev.off()
 #   R.utils::compressPDF(paste0("figures/PAK_mean.pdf"))
```

#### NPFSS method

**Most Likely Cluster (MLC)**

```{r}
res_np <- compute_np(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_np
```

**Significance (using parallel computing)**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

require(parallel)
cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np$stat < x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_np$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_np$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```

```{r}
res_np_2 <- compute_np(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_np_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np_2$stat < x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```



#### PFSS method

**Most Likely Cluster (MLC)**

```{r}
res_p <- compute_p(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_p
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p$stat <x))) /  (1 + B))
```



```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_p$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_p$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_p_2 <- compute_p(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_p_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p_2$stat <x))) /  (1 + B))
```

```{r, echo = F}
cat("p-value:  0.001")
```




#### DFFSS method

**Most Likely Cluster (MLC)**

```{r}
res_dffss <- compute_dffss(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_dffss
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_p$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_dffss$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_dffss$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```



```{r}
res_dffss_2 <- compute_dffss(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_dffss_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_dffss_2$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```




### HFSS method

**Most Likely Cluster (MLC)**

First, we determine the value of $K$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV.pdf"), width = 6, height = 4)
temp <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), 
                           ncol(MatX), plot_eigen = T)
#dev.off()
```

We choose $K=6$:

```{r}
res_h <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), 6)
res_h
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_h$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_h$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


We seek the optimal value of $K$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV_2.pdf"), width = 6, height = 4)
temp <- compute_h(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], t(MatX), 
                           ncol(MatX), plot_eigen = T)
#dev.off()
```

We choose $K=6$.

```{r}
res_h_2 <- compute_h(cluster_g1_temp[-id_pos], 
                     cluster_g2_temp[-id_pos], t(MatX), 6)
res_h_2
```


**Significance of the 2MLC**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h_2$stat <x))) /  (1 + B))
```



```{r, echo = F}
cat("p-value:  0.001")
```


### Summary of the results

**Visualization of the result**

```{r}
res <- vector("list", 4)
res[[1]][[1]] <- res_h
res[[1]][[2]] <- res_h_2
res[[2]][[1]] <- res_dffss
res[[2]][[2]] <- res_dffss_2
res[[3]][[1]] <- res_np
res[[3]][[2]] <- res_np_2
res[[4]][[1]] <- res_p
res[[4]][[2]] <- res_p_2
```

```{r, fig.width = 12, fig.height = 6, message = F, warning = F}
title_var <- 'Maximum consecutive 5-days precipitation (in mm)'
dates <- unique_year[chosen_years]
y_lim <- range(MatX)

for(k in 1:4) {
  my_cluster_1 <- res[[k]][[1]]$vec
  my_cluster_2 <- res[[k]][[2]]$vec

#pdf(file = paste0("figures/", my_country, "_", names_method[k], ".pdf"), width = 13, height = 4.2) 
  sf_use_s2(F)
nf <- layout( matrix(c(1,1,2,3), nrow=2, byrow=F) )
  par(mar = c(1.5, 0, 0, 0.2), 
      oma = c(0.5, 0, 2.4, 0), mgp = c(2.4, 0.6, 0), las = 1)
  ##### Map #########
    # map
  col_geo <- rep(rgb(0.9, 0.9, 0.9, alpha = 0.1), nrow(poly_cell))
  cex_geo <- rep(0.7, nrow(poly_cell))
  col_geo[my_cluster_1] <- alpha(cols[1], 0.8)
  cex_geo[my_cluster_1] <- 1

  col_geo[my_cluster_2] <- alpha(cols[2], 0.5)
  cex_geo[my_cluster_2] <- 1
  
  plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  mf_shadow(st_union(poly_cell[my_cluster_1, ]), 
                add = T, cex = 0.8, col = cols[1])
  mf_shadow(st_union(poly_cell[my_cluster_2, ]), 
         add = T, col= cols[2], cex = 0.8)
      
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

  
  plot(st_geometry(poly_cell), 
        border = "white",
        col = col_geo, 
        cex = cex_geo,
        pch = 16, asp = 1, add = T, lwd = 0.1)

  plot(st_geometry(st_union(poly_cell[my_cluster_1, ])), 
         add = T, border= cols[1], col = NULL)
  plot(st_geometry(st_union(poly_cell[my_cluster_2, ])), 
         add = T, border= cols[2], col = NULL)
      
  temp_1 <- draw.circle(coord_proj[my_cluster_1[1], 1], 
                           coord_proj[my_cluster_1[1], 2], 
                  as.numeric(dist_proj[my_cluster_1[1], 
                                       my_cluster_1[length(my_cluster_1)]]))

    my_circle_1 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_1$x, temp_1$x[1]), 
             c(temp_1$y, temp_1$y[1]))
         )), crs = my_proj
       ), 4326)
    
  temp_2 <- draw.circle(coord_proj[my_cluster_2[1], 1], 
                           coord_proj[my_cluster_2[1], 2], 
                  as.numeric(dist_proj[my_cluster_2[1], 
                                       my_cluster_2[length(my_cluster_2)]]))
    
  my_circle_2 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_2$x, temp_2$x[1]), 
             c(temp_2$y, temp_2$y[1]))
         )), crs = my_proj
       ), 4326)
  
  
  ###############
  plot(st_geometry(my_circle_2), add = T, border= cols[2],
             col = alpha(cols[2], 0.4), lty=1, lwd=1)
  plot(st_geometry(my_circle_1), add = T, border= cols[1],
             col = alpha(cols[1], 0.4), lty=1, lwd=1)
  
    mtext(my_var, side = 4, line = -3.5, las = 0, cex = 0.8)
    
  legend("topleft", legend = c("Most likely cluster", "Secondary cluster"),
         fill = c(cols[1], cols[2]), cex = 0.9, box.lty = 0)
      
  ##### Functional data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = '',
       ylab = '', col = rgb(0.6, 0.6, 0.6, alpha = 0.5), xaxt = 'n', 
        type = "l")
  legend("topleft", legend = c("Most likely cluster"),
         lty = 1, col = c(cols[1]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 2500, by = 200),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))

  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
    
  for(i in my_cluster_1)
        lines(dates, MatX[i, ], col = alpha(cols[1], 0.3),
          lty = 1, lwd = 1.3)
  
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  

plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = title_var, xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  legend("topleft", legend = c("Secondary cluster"),
         lty = 1, col = c(cols[2]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 2500, by = 200),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
      
  for(i in my_cluster_2)
        lines(dates, MatX[i, ], col = alpha(cols[2], 0.3),
          lty = 1, lwd = 1.3)
  
    
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  
  mtext(paste0("Clusters for the ", names_method[k]), side = 3, line = 0.8, outer = TRUE)
#dev.off()
#R.utils::compressPDF(paste0("figures/", my_country, "_", names_method[k], ".pdf"))
}
```

The following table presents the results obtained from the various methods.

```{r}
res_PAK <- data.frame(nb_cluster_1 = c(length(res_np$vec), 
                            length(res_p$vec),
                            length(res_dffss$vec),
                            length(res_h$vec)),
           sign_cluster_1 = c(0.001, 0.001, 0.001, 0.001),
           nb_cluster_2 = c(length(res_np_2$vec), 
                            length(res_p_2$vec),
                            length(res_dffss_2$vec),
                            length(res_h_2$vec)),
           sign_cluster_2 = c(0.001, 0.001, 0.001, 0.001))
row.names(res_PAK) <- c("NPFSS", "PFSS", "DFFSS", "HFSS")
knitr::kable(res_PAK)
```




### Venezuela

We select the ISO3 code for Venezuela:

```{r}
my_country <- "VEN"
my_proj <- 4247 
```

```{r, warning = F, message = F}
select_countries <- countries_regions[countries_regions$color_code %in% my_country, ]
  
# drop the islands
sf_obj <- select_countries %>%
    filter(iso3 == my_country[1]) %>%
    mutate(area = st_area(geometry)) %>%
    top_n(1, area) %>%
    rowid_to_column() %>%
    st_cast("POLYGON")  %>% 
    mutate(area = st_area(geometry)) %>%
    group_by(rowid) %>%
    top_n(1, area)
  
if(length(my_country) > 1) {
  for(i in 2:length(my_country)) {
    sf_obj <- rbind(sf_obj, select_countries %>%
      filter(iso3 == my_country[i]) %>%
      mutate(area = st_area(geometry)) %>%
      top_n(1, area) %>%
      rowid_to_column() %>%
      st_cast("POLYGON")  %>% 
      mutate(area = st_area(geometry)) %>%
      group_by(rowid) %>%
      top_n(1, area)
    )
  }
  }
is_intersect <- st_intersects(all_cells, sf_obj)
is_intersect <- which(sapply(is_intersect, function(x) length(x) != 0))
lldata_poly <- all_cells[is_intersect, ]
  
my_contours <- st_intersection(select_countries, 
                               st_union(lldata_poly, is_coverage = T))

# dowload tiles and compose raster (SpatRaster)
nc_osm <- get_tiles(my_contours, 
                      provider = "Esri.WorldShadedRelief", 
                      zoom = 7, crop = F)
poly_cell <- merge(lldata_poly, my_grid, by = c("long", "lat"))
# coordinates of the centroid
coord_temp <- st_transform(poly_cell, my_proj)
# simplify the geometry
poly_cell <- st_intersection(poly_cell, my_contours)
```

We compute all possible clusters:

```{r, warning = F}
  coord_proj <- st_coordinates(st_centroid(coord_temp))
  pairs_geo <- find_all_cluster(coord_proj)
  dist_proj <- as(dist(cbind(coord_proj[, 1], coord_proj[, 2])), "matrix")

  dist_4326 <- as(dist(cbind(poly_cell$long, poly_cell$lat)), "matrix")
  coord_4326 <- st_transform(poly_cell, 4326) 
```

```{r}
  my_var <- "hwd_upp_" 
  temp_var <- poly_cell[, paste0(my_var, unique_year)[chosen_years]]
  MatX <- as.matrix(st_drop_geometry(temp_var))
```

#### Descriptive Analysis

First, we plot the data:

```{r, fig.width = 10, height = 4}
y_lim <- range(MatX)
title_var <- 'Heat wave duration (in days)'
#pdf(file = "figures/VEN_data.pdf", width = 11, height = 4.5)
par(mfrow = c(1, 2), mar = c(3.7, 3, 1, 1), oma = c(0, 0, 0, 0),
    las = 1, mgp = c(2.2, 0.8, 0))
# map
plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

# data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = title_var, xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
abline(h = seq(0, 250, by = 50),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
#dev.off()
#  R.utils::compressPDF("figures/VEN_data.pdf")
```

We map the average of the variable over a 3-year window. 

```{r, fig.width = 12, fig.height= 6}
nb_split <- 8
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/VEN_evol.pdf"), width = 12, height = 5)
par(mfrow = c(2, 4), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 7, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlGnBu"
    my_pal <- rev(alpha(colorspace::sequential_hcl(7, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
      title(paste0(my_years[1], "-", my_years[length(my_years)]), line = -1.25)
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Heatwave days")
    }
#dev.off()
# R.utils::compressPDF("figures/VEN_evol.pdf")    
```

Average:

```{r, fig.width = 12, fig.height= 6}
nb_split <- 1
step_years <- split(chosen_years, 
           sort(rep_len(1:nb_split, length.out = length(chosen_years))))
#pdf(paste0("figures/VEN_mean.pdf"), width = 7, height = 5)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
  my_vec <- NULL
    for (j in 1:nb_split) {
      my_vec <- c(my_vec, rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[step_years[[j]]]])))
    }
    my_interval <- round(classInt::classIntervals(my_vec, 7, style = "jenks")$brks, digits = 4)

    nom_pal <- "YlGnBu"
    my_pal <- rev(alpha(colorspace::sequential_hcl(7, palette = nom_pal), 1))
    my_col <- my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)]

    for (j in 1:nb_split) {
      chosen_years_5 <- step_years[[j]]
      poly_cell$my_mean <- rowMeans(st_drop_geometry(poly_cell[, paste0(my_var, unique_year)[chosen_years_5]]))
      my_col <- alpha(my_pal[findInterval(poly_cell$my_mean, my_interval, all.inside = T)],
                  0.5)
    
      plot_tiles(nc_osm)
      plot(st_geometry(poly_cell), 
        col = my_col,
        border = my_col, lwd = 0.001, add = T)
      my_years <- unique_year[chosen_years_5]
   #   title(paste0(my_years[1], "-", my_years[length(my_years)]), line = -1.25)
     plot(st_geometry(my_contours), add = T, lwd = 0.5)
     if(j == 1)
       maplegend::leg(type = "choro", val = my_interval, pos = "topleft", 
                 pal = my_pal, val_rnd = 3, title = "Heatwave days")
    }
#dev.off()
#    R.utils::compressPDF("figures/VEN_mean.pdf")
```

#### NPFSS method

**Most Likely Cluster (MLC)**

```{r}
res_np <- compute_np(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_np
```

**Significance (using parallel computing)**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np$stat < x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_np$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_np$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_np_2 <- compute_np(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_np_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_np <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_np)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_np, function(x) res_np_2$stat < x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


#### PFSS method

**Most Likely Cluster (MLC)**

```{r}
res_p <- compute_p(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_p
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p$stat <x))) /  (1 + B))
```



```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_p$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_p$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


```{r}
res_p_2 <- compute_p(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_p_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_p <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_p)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_p, function(x) res_p_2$stat <x))) /  (1 + B))
```

```{r, echo = F}
cat("p-value:  0.001")
```




#### DFFSS method

**Most Likely Cluster (MLC)**

```{r}
res_dffss <- compute_dffss(pairs_geo[[1]], pairs_geo[[2]], t(MatX))
res_dffss
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_p$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_dffss$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_dffss$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```

```{r}
res_dffss_2 <- compute_dffss(cluster_g1_temp[-id_pos], 
                       cluster_g2_temp[-id_pos], t(MatX))
res_dffss_2
```

**Significance**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_dffss <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_dffss)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_dffss, function(x) res_dffss_2$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```



### HFSS method 

**Most Likely Cluster (MLC)**

First, we determine the value of $K$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV.pdf"), width = 6, height = 4)
temp <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), 
                           ncol(MatX), plot_eigen = T)
#dev.off()
```

We choose $K=6$:
```{r}
res_h <- compute_h(pairs_geo[[1]], pairs_geo[[2]], t(MatX), 6)
res_h
```

**Significance**

```{r, eval = F}
c1 <- pairs_geo[[1]]
c2 <- pairs_geo[[2]]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h$stat <x))) /  (1 + B))
```


```{r, echo = F}
cat("p-value:  0.001")
```


**Secondary cluster (2MLC)**

```{r}
cluster_g1_temp <- sapply(pairs_geo[[1]], function(x) setdiff(x, res_h$vec))
cluster_g2_temp <- sapply(pairs_geo[[2]], function(x) setdiff(x, res_h$vec))
id_pos <- union(which(sapply(cluster_g1_temp, function(x) length(x) == 0)),
          which(sapply(cluster_g2_temp, function(x) length(x) == 0)))
```


We seek the optimal value of $K$:

```{r, fig.width = 6, fig.height = 3}
#pdf(paste0("figures/", my_country, "_h_CPV_2.pdf"), width = 6, height = 4)
temp <- compute_h(cluster_g1_temp[-id_pos], cluster_g2_temp[-id_pos], t(MatX), 
                           ncol(MatX), plot_eigen = T)
#dev.off()
```

We choose $K=6$.

```{r}
res_h_2 <- compute_h(cluster_g1_temp[-id_pos], 
                     cluster_g2_temp[-id_pos], t(MatX), 6)
res_h_2
```


**Significance of the 2MLC**

```{r, eval = F}
c1 <- cluster_g1_temp[-id_pos]
c2 <- cluster_g2_temp[-id_pos]

cl <- makeCluster(10)
clusterExport(cl, c("c1", "c2", "MatX", "norm"))
res_par_h <- clusterApplyLB(cl, 1:B, compute_fun_par, compute_h, 6)
stopCluster(cl)

cat("p-value: ", (1 + sum(sapply(res_par_h, function(x) res_h_2$stat <x))) /  (1 + B))
```



```{r, echo = F}
cat("p-value:  0.001")
```




### Summary of the results

**Visualization of the result**

```{r}
res <- vector("list", 4)
res[[1]][[1]] <- res_h
res[[1]][[2]] <- res_h_2
res[[2]][[1]] <- res_dffss
res[[2]][[2]] <- res_dffss_2
res[[3]][[1]] <- res_np
res[[3]][[2]] <- res_np_2
res[[4]][[1]] <- res_p
res[[4]][[2]] <- res_p_2
```

```{r, fig.width = 12, fig.height = 6, message = F, warning = F}
my_var <- 'Heat wave duration (in days)'
dates <- unique_year[chosen_years]
y_lim <- range(MatX)

for(k in 1:4) {
  my_cluster_1 <- res[[k]][[1]]$vec
  my_cluster_2 <- res[[k]][[2]]$vec

#pdf(file = paste0("figures/", my_country, "_", names_method[k], ".pdf"), width = 13, height = 4.2) 
sf_use_s2(F)
nf <- layout( matrix(c(1,1,2,3), nrow=2, byrow=F) )
  par(mar = c(1.5, 0, 0, 0.2), 
      oma = c(0.5, 0, 2.4, 0), mgp = c(2.4, 0.6, 0), las = 1)
  ##### Map #########
    # map
  col_geo <- rep(rgb(0.9, 0.9, 0.9, alpha = 0.1), nrow(poly_cell))
  cex_geo <- rep(0.7, nrow(poly_cell))
  col_geo[my_cluster_1] <- alpha(cols[1], 0.8)
  cex_geo[my_cluster_1] <- 1

  col_geo[my_cluster_2] <- alpha(cols[2], 0.5)
  cex_geo[my_cluster_2] <- 1
  
  plot_tiles(nc_osm)
  mf_shadow(my_contours, add = T, cex = 0.8)
  mf_shadow(st_union(poly_cell[my_cluster_1, ]), 
                add = T, cex = 0.8, col = cols[1])
  mf_shadow(st_union(poly_cell[my_cluster_2, ]), 
         add = T, col= cols[2], cex = 0.8)
      
  plot(st_geometry(poly_cell), border = rgb(0.5, 0.5, 0.5), 
           lwd = 0.4, add = T, col = rgb(0.82, 0.82, 0.82))

  
  plot(st_geometry(poly_cell), 
        border = "white",
        col = col_geo, 
        cex = cex_geo,
        pch = 16, asp = 1, add = T, lwd = 0.1)

  plot(st_geometry(st_union(poly_cell[my_cluster_1, ])), 
         add = T, border= cols[1], col = NULL)
  plot(st_geometry(st_union(poly_cell[my_cluster_2, ])), 
         add = T, border= cols[2], col = NULL)
      
  temp_1 <- draw.circle(coord_proj[my_cluster_1[1], 1], 
                           coord_proj[my_cluster_1[1], 2], 
                  as.numeric(dist_proj[my_cluster_1[1], 
                                       my_cluster_1[length(my_cluster_1)]]))

    my_circle_1 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_1$x, temp_1$x[1]), 
             c(temp_1$y, temp_1$y[1]))
         )), crs = my_proj
       ), 4326)
    
  temp_2 <- draw.circle(coord_proj[my_cluster_2[1], 1], 
                           coord_proj[my_cluster_2[1], 2], 
                  as.numeric(dist_proj[my_cluster_2[1], 
                                       my_cluster_2[length(my_cluster_2)]]))
    
  my_circle_2 <- st_transform(st_sfc(st_polygon(
         list(
           cbind(
             c(temp_2$x, temp_2$x[1]), 
             c(temp_2$y, temp_2$y[1]))
         )), crs = my_proj
       ), 4326)
  
  
  ###############
  plot(st_geometry(my_circle_2), add = T, border= cols[2],
             col = alpha(cols[2], 0.4), lty=1, lwd=1)
  plot(st_geometry(my_circle_1), add = T, border= cols[1],
             col = alpha(cols[1], 0.4), lty=1, lwd=1)
  
    mtext(my_var, side = 4, line = -3.5, las = 0, cex = 0.8)
    
  legend("topleft", legend = c("Most likely cluster", "Secondary cluster"),
         fill = c(cols[1], cols[2]), cex = 0.9, box.lty = 0)
      
  ##### Functional data
  plot(dates, MatX[1, ], ylim = y_lim, xlab = '',
       ylab = '', col = rgb(0.6, 0.6, 0.6, alpha = 0.5), xaxt = 'n', 
        type = "l")
  legend("topleft", legend = c("Most likely cluster"),
         lty = 1, col = c(cols[1]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 2500, by = 200),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))

  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
    
  for(i in my_cluster_1)
        lines(dates, MatX[i, ], col = alpha(cols[1], 0.3),
          lty = 1, lwd = 1.3)
  
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  

plot(dates, MatX[1, ], ylim = y_lim, xlab = 'Years',
       ylab = my_var, xaxt = "n", 
       col = rgb(0.6, 0.6, 0.6, alpha = 0.5),
        type = "l")
  legend("topleft", legend = c("Secondary cluster"),
         lty = 1, col = c(cols[2]), cex = 0.9)
  abline(v = seq(1980, 2025, by = 5), lty = 2, 
             col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  abline(h = seq(0, 2500, by = 200),
             lty = 2, col = rgb(0.7, 0.7, 0.7, alpha = 0.3))
  axis(1, at = seq(1980, 2025, by = 5), xlab = "years",
           labels = as.character(seq(1980, 2025, by = 5)))
  for (j in 2:nrow(MatX))
        lines(dates, MatX[j, ], lwd = 1.3, 
          col = rgb(0.4, 0.4, 0.4, alpha = 0.1)) 
      
  for(i in my_cluster_2)
        lines(dates, MatX[i, ], col = alpha(cols[2], 0.3),
          lty = 1, lwd = 1.3)
  
    
  lines(dates, colMeans(MatX), lwd = 1.3, lty = 2)
  
  mtext(paste0("Clusters for the ", names_method[k]), side = 3, line = 0.8, outer = TRUE)
#dev.off()
#R.utils::compressPDF(paste0("figures/", my_country, "_", names_method[k], ".pdf"))

}
```

The following table presents the results obtained from the various methods.


```{r}
res_VEN <- data.frame(nb_cluster_1 = c(length(res_np$vec), 
                            length(res_p$vec),
                            length(res_dffss$vec),
                            length(res_h$vec)),
           sign_cluster_1 = c(0.001, 0.001, 0.001, 0.001),
           nb_cluster_2 = c(length(res_np_2$vec), 
                            length(res_p_2$vec),
                            length(res_dffss_2$vec),
                            length(res_h_2$vec)),
           sign_cluster_2 = c(0.001, 0.001, 0.001, 0.001))
row.names(res_VEN) <- c("NPFSS", "PFSS", "DFFSS", "HFSS")
knitr::kable(res_VEN)
```

